\part{Gestation}

\chapter{Gestation 1.1}

Class ended in five minutes and all I could think was, {\em an hour is too long for lunch}.

Since the start of the semester, I had been looking forward to the part of Mr. Gladly's World Issues class where we'd start discussing capes. Now that it had finally arrived, I couldn't focus. I fidgeted, my pen moving from hand to hand, tapping, or absently drawing some figure in the corner of the page to join the other doodles. My eyes were restless too, darting from the clock above the door to Mr. Gladly and back to the clock. I wasn't picking up enough of his lesson to follow along. Twenty minutes to twelve; five minutes left before class ended.

He was animated, clearly excited about what he was talking about, and for once, the class was listening. He was the sort of teacher who tried to be friends with his students, the sort who went by \quotation{Mr. G} instead of Mr. Gladly. He liked to end class a little earlier than usual and chat with the popular kids, gave lots of group work so others could hang out with their friends in class, and had \quote{fun} assignments like mock trials.

He struck me as one of the \quote{popular} kids who had become a teacher. He probably thought he was everyone's favorite. I wondered how he'd react if he heard {\em my} opinion on the subject. Would it shatter his self image or would he shrug it off as an anomaly from the gloomy girl that never spoke up in class?

I glanced over my shoulder. Madison Clements sat two rows to my left and two seats back. She saw me looking and smirked, her eyes narrowing, and I lowered my eyes to my notebook. I tried to ignore the ugly, sour feeling that stewed in my stomach. I glanced up at the clock. Eleven-forty-three.

\quotation{Let me wrap up here,} Mr. Gladly said, Sorry, guys, but there {\em is} homework for the weekend. Think about capes and how they've impacted the world around you. Make a list if you want, but it's not mandatory. On Monday we'll break up into groups of four and see what group has the best list. I'll buy the winning group treats from the vending machine.

There were a series of cheers, followed by the classroom devolving into noisy chaos. The room was filled with sounds of binders snapping shut, textbooks and notebooks being slammed closed, chairs screeching on cheap tile and the dull roar of emerging conversation. A bunch of the more social members of the class gathered around Mr. Gladly to chat.

Me? I just put my books away and kept quiet. I'd written down almost nothing in the way of notes; there were collections of doodles spreading across the page and numbers in the margins where I'd counted down the minutes to lunch as if I was keeping track of the timer on a bomb.

Madison was talking with her friends. She was popular, but not gorgeous in the way the stereotypical popular girls on TV were. She was \quote{adorable}, instead. Petite. She played up the image with sky blue pins in her shoulder length brown hair and a cutesy attitude. Madison wore a strapless top and denim skirt, which seemed absolutely moronic to me given the fact that it was still early enough in the spring that we could see our breath in the mornings.

I wasn't exactly in a position to criticize her. Boys liked her and she had friends, while the same was hardly true for me. The only feminine feature I had going for me was my dark curly hair, which I'd grown long. The clothes I wore didn't show skin, and I didn't deck myself out in bright colors like a bird showing off its plumage.

Guys liked her, I think, because she was appealing without being intimidating.

If they only knew.

The bell rang with a lilting ding-dong, and I was the first one out the door. I didn't run, but I moved at a decent clip as I headed up the stairwell to the third floor and made my way to the girl's washroom.

There were a half dozen girls there already, which meant I had to wait for a stall to open up. I nervously watched the door of the bathroom, feeling my heart drop every time someone entered the room.

As soon as there was a free stall, I let myself in and locked the door. I leaned against the wall and exhaled slowly. It wasn't quite a sigh of relief. Relief implied you felt better. I wouldn't feel better until I got home. No, I just felt less uneasy.

It took maybe five minutes before the noise of others in the washroom stopped. A peek below the partitions showed that there was nobody else in the other stalls. I sat on the lid of the toilet and got my brown bag lunch to begin eating.

Lunch on the toilet was routine now. Every school day, I would finish off my brown bag lunch, then I'd do homework or read a book until lunch hour was over. The only book in my bag that I hadn't already read was called \quote{Triumvirate}, a biography of the leading three members of the Protectorate. I was thinking I would spend as long as I could on Mr. Gladly's assignment before reading, because I wasn't enjoying the book. Biographies weren't my thing, and they were especially not my thing when I was suspicious it was all made up.

Whatever my plan, I didn't even have a chance to finish my pita wrap. The door of the bathroom banged open. I froze. I didn't want to rustle the bag and clue anyone into what I was doing, so I kept still and listened.

I couldn't make out the voices. The noise of the conversation was obscured by giggling and the sound of water from the sinks. There was a knock on the door, making me jump. I ignored it, but the person on the other side just repeated the knock.

\quotation{Occupied,} I called out, hesitantly.

\quotation{Oh my god, it's Taylor!} one of the girls on the outside exclaimed with glee, then in response to something another girl whispered, I barely heard her add, \quotation{Yeah, do it!}

I stood up abruptly, letting the brown bag with the last mouthful of my lunch fall to the tiled floor. Rushing for the door, I popped the lock open and pushed. The door didn't budge.

There were noises from the stalls on either side of me, then a sound above me. I looked up to see what it was, only to get splashed in the face. My eyes started burning, and I was momentarily blinded by the stinging fluid in my eyes and my blurring of my glasses. I could taste it as it ran down to my nose and mouth. Cranberry juice.

They didn't stop there. I managed to pull my glasses off just in time to see Madison and Sophia leaning over the top of the stall, each of them with plastic bottles at the ready. I bent over with my hands shielding my head just before they emptied the contents over me.

It ran down the back of my neck, soaked my clothes, fizzed as it ran through my hair. I pushed against the door again, but the girl on the other side was braced against it with her body.

If the girls pouring juice and soda on me were Madison and Sophia, that meant the girl on the other side of the door was Emma, leader of the trio. Feeling a flare of anger at the realization, I shoved on the door, the full weight of my body slamming against it. I didn't accomplish anything, and my shoes lost traction on the juice-slick floor. I fell to my knees in the puddling juice.

Empty plastic bottles with labels for grape and cranberry juice fell to the ground around me. A bottle of orange soda bounced off my shoulder to splash into the puddle before rolling under the partition and into the next stall. The smell of the fruity drinks and sodas was sickly sweet.

The door swung open, and I glared up at the three girls. Madison, Sophia and Emma. Where Madison was cute, a late bloomer, Sophia and Emma were the types of girls that fit the \quote{prom queen} image. Sophia was dark skinned, with a slender, athletic build she'd developed as a runner on the school track team. Red-headed Emma, by contrast, had all the curves the guys wanted. She was good looking enough to get occasional jobs as a amateur model for the catalogs that the local department stores and malls put out. The three of them were laughing like it was the funniest thing in the world, but the sounds of their amusement barely registered with me. My attention was on the faint roar of blood pumping in my ears and an urgent, ominous crackling \quote{sound} that wouldn't get any quieter or less persistent if I covered my ears with my hands. I could feel dribblesrunning down my arms and back, still chilled from the refrigerated vending machines.

I didn't trust myself to say something that wouldn't give them fodder to taunt me with, so I kept silent.

Carefully, I climbed to my feet and turned my back on them to get my backpack off the top of the toilet. Seeing it gave me pause. It had been a khaki green, before, but now dark purple blotches covered it, most of the contents of a bottle of grape juice. Pulling the straps around my shoulders, I turned around. The girls weren't there. I heard the bathroom door bang shut, cutting off the sounds of their glee, leaving me alone in the bathroom, drenched.

I approached the sink and stared at myself in the scratched, stained mirror that was bolted above it. I had inherited a thin lipped, wide, expressive mouth from my mother, but my large eyes and my gawky figure made me look a lot more like my dad. My dark hair was soaked enough that it clung to my scalp, neck and shoulders. I was wearing a brown hooded sweatshirt over a green t-shirt, but colored blotches of purple, red and orange streaked both. My glasses were beaded with the multicolored droplets of juice and soda. A drip ran down my nose and fell from the tip to land in the sink.

Using a paper towel from the dispenser, I wiped my glasses off and put them on again. The residual streaks made it just as hard to see, if not worse than it had been.

{\em Deep breaths, Taylor}, I told myself.

I pulled the glasses off to clean them again with a wet towel, and found the streaks were still there.

An inarticulate scream of fury and frustration escaped my lips, and I kicked the plastic bucket that sat just beneath the sink, sending it and the toilet brush inside flying into the wall. When that wasn't enough, I pulled off my backpack and used a two-handed grip to hurl it. I wasn't using my locker anymore: certain individuals had vandalized or broken into it on four different occasions. My bag was heavy, loaded down with everything I'd anticipated needing for the day's classes. It crunched audibly on impact with the wall.

\quotation{What the fuck!?} I screamed to nobody in particular, my voice echoing in the bathroom. There were tears in the corners of my eyes.

\quotation{The hell am I supposed to do!?} I wanted to hit something, break something. To retaliate against the unfairness of the world. I almost struck the mirror, but I held back. It was such a small thing that it felt like it would make me feel {\em more} insignificant instead of venting my frustration.

I'd been enduring this from the very first day of high school, a year and a half ago. The bathroom had been the closest thing I could find to refuge. It had been lonely and undignified, but it had been a place I could retreat to, a place where I was off their radar. Now I didn't even have that.

I didn't even know what I was supposed to do for my afternoon classes. Our midterm project for art was due, and I couldn't go to class like this. Sophia would be there, and I could just imagine her smug smile of satisfaction as I showed up looking like I'd botched an attempt to tie-dye everything I owned.

Besides, I'd just thrown my bag against the wall and I doubted my project was still in one piece.

The buzzing at the edge of my consciousness was getting worse. My hands shook as I bent over and gripped the edge of the sink, let out a long, slow breath, and let my defenses drop. For three months, I'd held back. Right now? I didn't care anymore.

I shut my eyes and felt the buzzing crystallize into concrete information. As numerous as stars in the night sky, tiny knots of intricate data filled the area around me. I could focus on each one in turn, pick out details. The clusters of data had been reflexively drifting towards me since I was first splashed in the face. They responded to my subconscious thoughts and emotions, as much of a reflection of my frustration, my anger, my {\em hatred} for those three girls as my pounding heart and trembling hands were. I could make them stop or direct them to move almost without thinking about it, the same way I could raise an arm or twitch a finger.

I opened my eyes. I could feel adrenaline thrumming through my body, blood coursing in my veins. I shivered in response to the chilled soft drinks and juices the trio had poured over me, with anticipation and with just a little fear. On every surface of the bathroom were bugs; Flies, ants, spiders, centipedes, millipedes, earwigs, beetles, wasps and bees. With every passing second, more streamed in through the open window and the various openings in the bathroom, moving with surprising speed. Some crawled in through a gap where the sink drain entered the wall while others emerged from the triangular hole in the ceiling where a section of foam tile had broken off, or from the opened window with peeling paint and cigarette butts squished out in the recesses. They gathered around me and spread out over every available surface; primitive bundles of signals and responses, waiting for further instruction.

My practice sessions, conducted away from prying eyes, told me I could direct a single insect to move an antennae, or command the gathered horde to move in formation. With one thought, I could single out a particular group, maturity or species from this jumble and direct them as I wished. An army of soldiers under my complete control.

It would be so easy, {\em so easy} to just go Carrie on the school. To give the trio their just desserts and make them regret what they had put me through: the vicious e-mails, the trash they'd upended over my desk, the flute --my mother's flute-- they'd stolen from my locker. It wasn't just them either. Other girls and a small handful of boys had joined in, \quote{accidentally} skipping over me when passing out assignment handouts, adding their own voices to the taunts and the flood of nasty emails, to get the favor and attention of three of the prettier and more popular girls in our grade.

I was all too aware that I'd get caught and arrested if I attacked my fellow students. There were three teams of superheroes and any number of solo heroes in the city. I didn't really care. The thought of my father seeing the aftermath on the news, his disappointment in me, his shame? That was more daunting, but it still didn't outweigh the anger and frustration.

Except I was better than that.

With a sigh, I sent an instruction to the gathered swarm. Disperse. The word wasn't as important as the idea behind it. They began to exit the room, disappearing into the cracks in the tile and through the open window. I walked over to the door and stood with my back to it so nobody could stumble onto the scene before the bugs were all gone.

However much I wanted to, I couldn't really follow through. Even as I trembled with humiliation, I managed to convince myself to pick up my backpack and head down the hall. I made my way out of the school, ignoring the stares and giggles from everyone I walked past, and caught the first bus that headed in the general direction of home. The chill of early spring compounded the discomfort of my soaked hair and clothes, making me shiver.

I was going to be a superhero. That was the goal I used to calm myself down at moments like these. It was what I used to make myself get out of bed on a school day. It was a crazy dream that made things tolerable. It was something to look forward to, something to work towards. It made it possible to keep from dwelling on the fact that Emma Barnes, leader of the trio, had once been my best friend.


\chapter{Gestation 1.2}

My thoughts were on Emma on the bus ride home. For an outside observer, I think it's easy to trivialize the importance of a \quote{best friend}, but when you're a kid, there's nobody more important. Emma had been my \quote{BFF} from grade one all the way through middle school. It hadn't been enough for us to spend our time together at school, so we had alternated staying at each others houses every weekend. I remember my mother saying that we were so close we were practically sisters.

A friendship that deep is intimate. Not in the rude way, but just in terms of a no-holds-barred sharing of every vulnerability and weakness.

So when I got back from nature camp just a week before our first year at high school started, to find that she wasn't talking to me? That she was calling Sophia her best friend? Discovering that she was now using every one of those secrets and vulnerabilities I had shared with her to wound me in the most vicious ways she could think of? It was crushing. There's just no better way to say it.

Unwilling to dwell on it any longer, I turned my attention to my backpack, setting it on the seat beside me and sorting through the contents. Grape juice had stained it, and I had a suspicion I would have to get a new one. I had bought it just four months ago, after my old one had been taken from my locker, and it had been just twelve bucks, so it wasn't a huge issue. The fact that my notebooks, textbooks and the two novels I'd shoved into my bag were wet with grape juice was more troubling. I suspected that whichever girl had been holding the grape juice had aimed for the open top of my bag as she poured it. I noted the destruction of my art project -- the box I'd put it in was collapsed on the one side. That bit was my fault.

My heart sank as I found the notebook with the white and black speckled hardcover. The corner of the paper was soaked through with as much as a quarter of each page stained purple. The ink had diluted and the pages were already turning wavy.

That notebook was -- had been -- my notes and journal for my hero career. The testing and training I'd done with my powers, pages of crossed out name ideas, even the measurements I was using for my costume in progress. After Emma, Madison and Sophia had stolen my last backpack and stuffed it in a wastebasket, I had realized how big a danger it was to have everything written down like that. I had copied everything over into a new notebook in a simple cipher and wrote it bottom to top. Now that notebook was spoiled, and I was looking at having to copy some two hundred pages of detailed writing into a new notebook if I wanted to preserve the information. If I could even remember what was on all of the ruined pages.

The bus stopped a block away from my house, and I got off, trying to ignore the stares. Even with the gawking, the knowledge that my notebook was ruined and my general nervousness about missing afternoon classes without permission, I felt better as I got closer to home. It felt worlds better to know I could drop my guard, stop watching my back and that I could take a break from wondering when the next incident would happen. I let myself into the house and headed straight for the shower, not even removing my backpack or taking off my shoes until I was in the bathroom.

I stood under the stream with my clothes on the floor of the tub, hoping the water would help get the worst of the juice out. I pondered. I don't know who said it, but at one point I had come across this notion about taking a negative and turning it into a positive. I tried to take the day's events and turn them around in my head, to see if I couldn't find a more positive twist on it.

Okay, so the first thing that came to mind was \quotation{Yet another reason to kill the trio.} It wasn't a serious thought -- I was angry, but it wasn't like I was going to actually kill them. Somehow, I suspected that I'd hurt myself before I hurt them. I was humiliated, frustrated, pissed, and I always had a weapon available -- my power. It was like having a loaded gun in your hand at all times. Except my power wasn't that great, so maybe it was more like having a taser. It was hard not to think about using it when things got really bad. Still, I didn't think I had that killer instinct in me.

No, I told myself, forcing myself back to the subject of positive thinking. Were there any upsides? Art project wrecked, clothes probably unrecoverable, needing a new backpack\ldots{} notebook. Somehow my mind fixated on that last part.

I cranked the shower to off, then toweled dry, thinking. I wrapped the towel around me, and rather than head to my room to get dressed, I put my wet clothes into a laundry hamper, grabbed my backpack and headed downstairs, through the kitchen and into the basement.

My house is old, and the basement was never renovated. The walls and floor are concrete and the ceiling was exposed boards and electrical cords. The furnace used to be coal fueled, and there was still an old coal chute, two feet by two feet, where the coal trucks used to come by to unload the winter's supply of coal for heating the house. The chute was boarded up, but around the time I was copying my original \quote{superpower notebook} over in code, I had decided to play it safe in all respects and start getting creative with my privacy. It was then that I'd started using it.

I removed one screw and removed the square wooden panel with the peeling white paint that covered the low end of the coal chute. I retrieved a gym bag from inside and put the panel back in place without screwing it back in.

I emptied the contents of the gym bag on the disused workbench that the house's previous owner had left in our basement, then opened the windows that were at the same level as the driveway and front garden. I closed my eyes and spent a minute exercising my power. I wasn't just grabbing every creepy crawly in a two block radius, though. I was being selective, and I was gathering quite a few.

It would take time for all of them to arrive. Bugs could move faster than you thought when they moved with purpose in a straight line, but even so, two blocks was a lot of ground for something so small to cover. I busied myself with opening the bag and sorting out the contents. My costume.

The first of the spiders started coming in through the open windows and congregating on the workbench. My power didn't give me a knowledge of the official names of the bugs I was working with, but anyone could recognize the spiders that were crawling into the room. These were black widows. One of the more dangerous spiders you could find in the States. Their bite could be lethal, though it usually wasn't, and they tended to bite with little provocation. Even under my complete control, they spooked me. At my request, the dozens upon dozens of spiders got into place on the workbench and began drawing out lines of webbing, laying the lines across one another, and weaving them into one work.

Three months ago, after I'd recovered from the manifestation of my powers, I had started to prepare for the goal I had set for myself. It had involved an exercise routine, training my power, research, and preparing my costume. Costumes were harder than one might think. While members of official teams surely had sources for that stuff, the rest of us were left to either buy costumes, put them together piecemeal with stuff bought from stores or make them from scratch. Each option had its problems. If you bought a costume online, you ran the risk of being traced, which could blow your secret identity before you'd even put a costume on. You could put a costume together with stuff bought from stores, but very few people could do that and look good. The final option, putting a costume together yourself, was just a hell of a lot of work and you could run into the issues of the prior two options -- being traced or winding up with a lame costume -- depending on where you got your materials and how you went about it.

In the second week after I'd figured out my powers, when I still wasn't entirely sure what was going on, I had come across a segment on the discovery channel about a suit that was made to withstand attacks by bears. That segment talked about how the suit was made of synthetic spider silk, which had inspired this particular project. Why go synthetic when you can produce with the real thing?

Okay, so it had been harder than that. Not just any spider worked, and the black widow spiders themselves were hard to find. They weren't typically found in the northeastern states, where it was generally colder, but I was fortunate that that key element that made Brockton Bay a tourist destination and a hotspot for capes also made it a place where black widow spiders could live, if not thrive. Namely, it was warm. Thanks to the surrounding geography and the ocean bordering us on the east, Brockton Bay had some of the mildest winters you could find in the Northeastern States, and some of the most comfortably warm summers. Both the black widows and the people running around in skintight costumes were thankful for that.

With my power, I had ensured the spiders could multiply. I'd kept them in safe locations and fattened them on prey I directed straight to them. I had flipped that mental switch that told them to breed and lay eggs as if it was summer, fed more prey to the hundreds of young that had resulted and had earned countless costume spinners for my trouble. The biggest issue had been that black widows are territorial, so I'd had to spread them out to ensure they didn't kill each other when I wasn't around to control them. Once a week or so, on my morning runs, I rotated the locations of the local spiders so I had a fresh supply all filled with proteins for the production of the essential materials. This ensured that the spiders were always ready for working on the costume in the afternoon, after school.

Yeah, I needed a life.

But I had a badass costume.

It wasn't a great looking costume, just yet. The fabric was a dirty yellow-gray. The armored sections had been made out of finely arranged and layered shells and exoskeletons I'd cannibalized from the local insect population and then reinforced with dragline silk. In the end, the armored parts had wound up dark mottled brown-gray. I was okay with that. When the entire thing was done, I planned to dye the fabric and paint the armor.

The reason I was so pleased with my costume was the fact that it was flexible, durable, and incredibly lightweight, considering the amount of armor I had put on it. At one point I had screwed up the dimensions of one of the legs, and when I tried to cut it off to start fresh, I had found I couldn't cut it with an x-acto knife. I had needed to use wire cutters, and even that had been a chore. As far as I figured, it was everything a superhero wanted for a costume.

I wasn't exactly willing to test it out, but I harbored hopes that it was bulletproof. Or at least, that the armored sections over my vital areas were.

The plan was to finish my costume over the course of the month, then as the school year ended and the summer began, I would take the leap into the world of superheroics.

But the plan had changed. I took off my towel and hung it from the corner of the bench, then began pulling on my costume to test the fit for the hundredth time. The spiders obediently moved out of my way as I did so.

When I had been standing in the shower, trying to find the good aspects in the day's troubles, my thoughts had turned to my notebook. I had realized I was procrastinating. I was constantly planning, preparing, considering all of the possibilities. There would always be more preparations, more stuff to study or test. The destruction of my notebook had been the burning of a bridge. I couldn't go back and copy it into a fresh book or start a new one without delaying my game plan for at least a week. I had to move forward.

It was time to do it. I flexed my hand inside the glove. I'd go out next week -- no. No more delays. This weekend, I would be ready.


\chapter{Gestation 1.3}

My training schedule consisted of running every morning and every other afternoon. In the process, I had picked up a pretty good knowledge of the east side of the city. Growing up in Brockton Bay, my parents had told me stuff like \quotation{stick to the Boardwalk}. Even on my runs, I had scrupulously stayed on the Boardwalk and avoided the bad part of town. Now it was Sunday night and I was in costume and breaking the rules.

I had dyed and painted the costume on Friday, bought temporary costume pieces (belt, the straps for the mask and the lenses) on Saturday and finished the most necessary details over the course of my Sunday afternoon before heading out for the evening. The costume wasn't complete yet, lacking the full extent of the armor paneling I had planned out, but the armor covered the most essential areas -- my face, chest, spine, stomach and major joints. The mask design featured dull yellow lenses, the only color on the black and gray costume, as well as sections of armor designed to imitate a bug's mandibles while simultaneously protecting my jaw. The mask left my hair free, which did leave the back of my head more vulnerable, but that was just one of the sacrifices I'd had to make to go out in an unfinished costume.

It was just after midnight, and I was crossing the line between one of the nicest parts of town and the part of town where the crack whores and gangsters lived. The distance between the two was thinner than one might think.

The Boardwalk was where the tourists came. Running north-to-south along the beach, there were shops that sold dresses for over a thousand dollars, cafes with ludicrously expensive coffees and stretches of wooden walkways and beaches where tourists could get a great view of the ocean. From pretty much any point on the Docks, you could see one of Brockton Bay's landmarks, the Protectorate Headquarters. Besides being a marvel of architectural design with its arches and towers, the PHQ was a floating base of operations that a squadron of local superheroes called home, outfitted with a forcefield bubble and a missile defense system. There had never been occasion for either to be used, but I had to admit, it made you feel safer.

If you headed west from the Boardwalk, away from the water, you found yourself in the area the locals just called the \quote{Docks}. When the import/export business in Brockton Bay had dried up, there had been a whole lot of people who were suddenly out of work. The richest and most resourceful people in town had managed to make more money, turning the city's resources towards tech and banking, but all of the people who had been employed on the ships and in the warehouses had few options left to them. They faced leaving Brockton Bay, sticking around while scraping up what little work they could or turning to more illicit activity.

This all contributed to the boom in the local supervillain population. The potential for big money coupled with the number of eager-to-please mooks and henchmen made it the city to be for the villains in the late 90s. It took a few years for the hero presence to establish and organize themselves, but they did, and there was something of an equilibrium now. As far as cape population went, Brockton Bay wasn't in the top 5 cities in the U.S., but it was probably in the top ten.

Just moving from one block to the next, you could see the change in the area. As I made my way into the Docks, I could see the quality of my surroundings decline steeply. There were enough warehouses and apartments in the area for even the most destitute to find shelter, so the only people on the streets were unconscious drunks, whores and gang members. I steered clear of any and all people I saw and ventured further into the area.

As I walked, I was using my powers to draw a swarm together, but kept them out of the way, moving just over the nearby rooftops and through the interior of buildings. Anyone paying attention to the local cockroach population might think something was up, but there weren't many lights on. I doubted most of the buildings here even had power.

The lack of lights in the area was what made me stop and draw myself against the side of a building when I saw a spot of orange in the dark street ahead. The orange was the flame of a lighter, and I was able to make out several faces around it. They were Asian, some wearing hoodies, others wearing headbands or long sleeved shirts, but all wore the same colors. Red and green.

I knew who these guys were. They were members from the local gang that left the tags \quote{Azn Bad Boys}, ABB for short, all over the East end of the city. More than a few went to my school. As far as the criminal element in Brockton Bay went, they weren't small potatoes. While the typical gang members were just Koreans, Japanese, Vietnamese and Chinese forcibly recruited from Brockton Bay's high schools and lower class neighborhoods, the gang was led by a couple of people with powers. Gangs didn't tend to be that racially inclusive as far as who joined, so it said something that their leader had the ability to draw in members from so many different nationalities and keep them in line.

The street was unlit, so my ability to see was dependent on the moon and the few indoor lights that were still on and shining out onto the sidewalks. I started actively looking for their boss. There were more gang members coming out of a two-story building, and they were gathering in the street. They didn't have the atmosphere of people who were just hanging out, either. They were expressionless or scowling, and they weren't talking.

I spotted their boss as the gang pulled away from the door of the building to give him room. I only knew about this guy from what I had heard on the news and read online, but I recognized him immediately. He was a big guy, but not so big that he would send people running when he walked down the street, like some people with powers were. He was a little over six feet, though, which put him head and shoulders above most of the gang members. He had an ornate metal mask over his face, and wasn't wearing a shirt, despite the chill. Sprawling tattoos covered his body from the neck down, all depicting dragons from Eastern mythology.

He went by \quote{Lung}, had successfully gone toe to toe with whole teams of heroes and had managed to keep himself out of jail, as evidenced by his presence here. As for his powers, I only knew what I could scrounge up online, and there were no guarantees there. I mean, for all I knew, he could have misled people about what his powers did, he could have a power he was keeping up his sleeve for an emergency, or he could even have a very subtle power that people couldn't see at work.

The information online and in the papers had told me this: Lung could gradually transform. Maybe it was based on adrenaline, his emotional state, or something, but whatever it was, it made his powers more potent the longer he was in a fight. He healed at a superhuman rate, got stronger, got tougher, got bigger, and he grew armor plating complete with blades at each fingertip. Rumor had it that he even grew wings if he fought long enough. If that wasn't enough, he was a pyrokinetic, which meant he could create flame out of thin air, shape it, intensify it, and so on. That power apparently got stronger as he transformed, too. As far as I knew, there wasn't an upper limit to how strong he could get. He only started returning to normal when there was nobody left to fight.

Lung wasn't the only one with powers in the ABB. He had a flunky, a scary sociopath called Oni Lee, who could teleport or create doubles of himself -- I wasn't a hundred percent sure on the details -- but Oni Lee had a distinctive look, and I didn't see him in the crowd. If there was anyone else with powers that I needed to watch out for, I hadn't seen or heard anything about them in my research.

Lung began talking in a deep, commanding voice. I couldn't make out the words, but it sounded like he was giving instructions. As I watched, one of the gang members drew a butterfly knife from his pocket, and another of them put his hand on his waistband. Between the gloom and the fact that I was standing half a block away, I couldn't see well, but a dark shape stood out against his green t-shirt. Chances were it was a gun handle. My pulse sped up a bit as I saw the gun, which was silly. Lung was more dangerous than fifty people with guns.

I decided to move away from where I was and find a better vantage point to monitor their conversation, which seemed like a good compromise between my curiosity and my self preservation. I slowly backed away from where I was, glancing over my shoulder to make sure nobody was watching, and then circled around the rear of the building I was lurking beside.

My investigation paid off. Halfway down the alley, I saw a fire escape that was leading up the back of the building that Lung and his gang were standing in front of. The feet of my costume had soft soles, so I was nearly silent as I ascended.

The roof was covered in gravel and cigarette butts, which made me think I wouldn't be nearly so quiet walking over it. Instead, I walked on the raised outside lip of the roof. As I neared the part of the roof directly above Lung and his gang of \quote{Azn Bad Boys}, I crouched and crawled forward on my stomach. It was dark enough that I doubted they would see me if I jumped up and down and waved my arms, but there was no reason to be stupid.

Being at the top of a two story building when they were on the ground floor made it hard to hear them. Lung had a strong accent, as well, which meant I had to wait until he had spoken a few sentences before I could figure out what he was saying. It helped that his mooks were utterly, respectfully silent as he spoke.

Lung was snarling, \quotation{\ldots{}the children, just shoot. Doesn't matter your aim, just shoot. You see one lying on the ground? Shoot the little bitch twice more to be sure. We give them no chances to be clever or lucky, understand?}

There was a murmur of assent.

Someone else lit up a cigarette, and then leaned over to light a cigarette for the guy next to him. In those moments that his hand wasn't cupped around the flame, I could see the gathered faces of just a dozen or so of the gangsters gathered around Lung. In hands, waistbands and holsters, I could see the dark metal of guns reflecting the orange flame. If I had to hazard a guess, all of them had weapons.

They were going to kill {\em kids}?


\chapter{Gestation 1.4}

I felt a chill. A part of me really wished that I had thought to get my hands on a disposable cell phone. I didn't have a utility belt, but the spade shaped section of armor that hung over my spine hid a set of EpiPens, a pen and notepad, a tube of pepper spray meant to hang off a key chain and a zippered pouch of chalk dust. I could have fit a cell phone back there. With a cell phone, I could have alerted the real heroes about the fact that Lung was planning to take a score of his flunkies to go and shoot kids.

At least, that's what I had heard. I was in a state of disbelief, turning the words around in my head to think of a different context that would make sense of it. It wasn't so much the fact that {\em he} would do something like that. I just had a hard time wrapping my head around the idea that anyone would.

Lung answered a question for one of his gang members, lapsing briefly into another language. He grabbed one of his minion's arms and twisted it to an angle where he could get a look at the guy's watch, so I guessed it had something to do with their timing or when they were leaving. The gang member who'd had his arm twisted winced as Lung let it go, but didn't complain.

What was I supposed to do? I doubted I could find any place in the Docks that would be willing to let me inside to use their phone. If I headed to the Boardwalk, I wasn't sure I would find any places that were still open, and I didn't have change for a payphone. That was another oversight I would have to correct for the next time I went out. Cell phone, spare change.

A car pulled up, and another three guys dressed in gang colors got out and and joined the crowd. Shortly after, the group -- twenty or twenty five in total -- started walking north, passing below me as they walked down the street.

I was out of time to consider my options. As much as I didn't want to face it, there was really only one option that I could have no regrets about. I shut my eyes and focused on every bug on the neighborhood, including the sizable swarm I had gathered on the way into the Docks. I took control of each of them.

{\em Attack.}

It was dark enough that I could only tell where the swarm was with my power. That meant I couldn't even tune out the swarm if I wanted to have any idea about what was going on. My brain was filled with horrendous amounts of information, as I sensed each bite, each sting. As the thousands of insects and arachnids swarmed over and around the group, I could almost see the outlines of each person, just by sensing the shapes of the surfaces the bugs were crawling on, or the areas the vermin wasn't occupying. I focused on keeping the more venomous types at bay for the time being -- I didn't need any allergic thugs going into anaphylactic shock from a bee sting or getting serious complications from the bite of a brown recluse spider.

I sensed the fire through the swarm before I realized what I was looking at with my eyes. My power told me of the bugs' recognition of the heat, but I didn't even have time to devote conscious thought to block out the instincts the fire set in motion before the damage was done. The primitive thought processes of my bugs were reduced to confused impulses to alternately flee and to pursue the heat and the light they so often used for navigation. Many bugs died or were crippled by the heat. From my vantage point, I could see Lung lashing out with streams of fire from his hands, directing them at the sky.

I suppressed a laugh, feeling heady with adrenaline. Was that all he could do? I directed the swarm to gather, so those who weren't already biting and stinging were in the midst of the gang. If he wanted to turn his flames on the swarm, he would have to set his own people on fire.

The heated air and the smells gave me enough information, by way of my insects, to tell where Lung was in the crowd. I took a deep breath, and then sent in the reserves. I took a share of the venomous types I'd held at bay and directed them to Lung. A handful of bees, wasps, a number of the more poisonous spiders, like black widows and brown recluses, and dozens of fire ants.

He healed fast when his power was working. Everything I'd read online said that people with healing abilities would shrug off the effects of poisons or drugs, so I knew I'd have to pump him full of enough venom to overwhelm that aspect of his power. Besides, he was a big guy. I judged he could take it.

From the information that I could glean from my bugs, Lung already had maybe a quarter of his body covered in armor. Triangular sections of metallic plating were piercing through his skin, where they would continue to grow and overlap until he was nigh impenetrable. If they weren't already, his fingertips and toes would become like blades or metal claws.

I felt a sadistic glee as I organized the attack on Lung. I directed the flying insects to attack his face. With distaste, I focused the crawling ants and spiders on\ldots{} other vulnerable areas. I did my best to ignore the feedback that I got from that particular attack, as I most definitely did not want the same kind of topographical map that the swarm had provided just a minute ago. Lung was bad news, and I needed him out of action as soon as possible. That meant delivering the hurt.

Rationale aside, I did feel a stab of guilt about taking pleasure in someone else's pain. I quieted that moment's remorse by reminding myself that Lung had spread tragedy, addiction and death to innumerable families. He had been planning to kill kids.

Lung exploded. No metaphor there. He detonated in a blast of rolling fire that set his clothes, several pieces of litter and one of his gang members alight. Almost every bug in his immediate vicinity died or was crippled by the wave of extreme heat. From my vantage point on the roof, I watched as he turned himself into a human bomb a second time. The second explosion turned his clothes to rags and sent his people fleeing for cover. He stepped out of the smoke with his hands burning like torches, the silvery scales that covered nearly a third of his body reflecting the flame.

Damn, damn, damn. He was fireproof? Or skilled enough at using fire to superheat the air around him without burning himself? The meager scraps of clothing that covered him were burning away, and fire licked and danced around his hands without him seeming to care.

He roared. It wasn't the monstrous roar one might expect, but a very human sound of rage and frustration. As human as it sounded, though, it was loud. All the way down the street neighborhood, lights and flashlights flickered on in response to the explosions and the roar. I even saw a few faces peering through windows to see the action. Idiots. If Lung's next attack shattered any glass, they could get hurt.

From where I was crouched on the side of the roof, I directed some of the more harmless insects to attack Lung. He lashed out with fire the moment they started crawling on him, which I had more or less expected. He was managing to kill the majority of the bugs with each burst of flame, and knowing what I did about his powers, I knew his flames would only get bigger, hotter and more dangerous.

In a typical fight, you figure someone would get weaker as the fight dragged on. They would take their lumps, get tired, exhaust their bag of tricks. With Lung, it was the opposite. I found myself regretting that I had used only a relatively small number of the more venomous bugs, because it was becoming clear that what I'd used wasn't having much effect. He had no idea where I was, so I figured I still had the upper hand, but my options and the number of bugs in my swarm were running out. Despite my earlier glee, I wasn't sure I could win this anymore.

I hissed through my teeth, all too aware that time was running out. Before long, Lung would set fire to the city block, become immune to bites and stings in general, or destroy my entire swarm. I had to get creative. I had to get meaner.

I focused my attention on a lone wasp, and piloted it around Lung's back, up behind his head and then had it circle around to his face and straight at his eyeball. The wasp touched his eyelash, and he blinked before it could hit the target. As a consequence, the stinger only sank into his eyelid, prompting yet another explosion of fire and a scream of rage.

Again. I thought. A honeybee this time. I wasn't sure if he eventually got armor plated eyelids, but maybe I could use the stings to make his eyes swell shut? He wouldn't be able to fight if he couldn't see.

The bee struck home this time, sinking his stinger into the ball of Lung's eye. It surprised me in that it didn't stick or kill the bee, so I had the bee sting again, and this time the barbs let it stick in the skin at the corner of his eye, at the side of his nose. The bee died that time, leaving some tiny organs and a venom sac hanging from the stinger.

I expected him to explode again. He didn't. Instead, he set himself on fire, head to toe. I waited a moment, poised to attack with the next wasp to attack the moment he dropped his guard, but as the seconds passed, I realized he wasn't planning on extinguishing himself. My heart sank.

Surely he was burning up all of the oxygen in his vicinity. Didn't he need to breathe? What the hell was the fuel source for his fire?

Standing in the street, he turned around, searching for me, with the flames that licked and rolled over his body casting light where there had been only gloom. Abruptly, he hunched over. I wondered if -- I hoped -- the various toxins and venoms in his system had done the trick. Then his back separated into two. A meaty looking gap appeared along his spine, followed by an eruption of long metallic scales all down the gap. After bristling for a few moments, the scales lay flat like dominoes falling. He stood and stretched, and I could swear he was a foot taller, now with an armor plated spine.

Still on fire, head to toe.

If the \quote{constantly on fire} thing had tipped the balance of the fight to futile, watching Lung grow and look stronger than ever had pushed me to the point of being spooked. I started thinking about an exit strategy. Rationally, I figured, Lung's men were scattered to the four winds and they were probably in pretty rough shape. Whatever Lung had been planning for tonight, chances were he wasn't going to be able to carry out whatever plans he'd had after this debacle. I had more or less accomplished what I needed to, and I figured I could run and find a way to contact the PHQ just in case.

That was the rational perspective. Justifications aside, I just wanted to leave, right then. If things dragged on and I stayed put, there was a very real chance that Lung would give evidence to the rumor that he could grow wings, at which point I would be spotted for sure. I wouldn't be able to beat Lung at this point, anyway, which left only a graceless retreat as the remaining option.

Lung had his back turned to me, so I lifted myself up, slowly. Crouching, I backed up to retreat to the fire escape, watching Lung carefully as I set foot on the gravel of the roof.

As if a gunshot had gone off, Lung whirled around to stare at me. One of his eyes was just a glowing line behind his mask, but the other was like an orb of molten metal.

A victorious roar filled the air, less human than the outcry he had made earlier, and I felt a kind of resignation. Enhanced hearing. The package of powers the bastard got from his transformation included superhuman hearing.

\chapter{Gestation 1.5}

You don't properly appreciate what superhuman strength means until you see someone leap from the sidewalk to the second floor of a building on the far side of the street. He didn't make it all the way to the roof, but he came to a point maybe three quarters of the way up. I wasn't sure just how Lung kept from falling, but I could only guess that he just buried his fingertips into the building's exterior.

I heard scraping and crunching as he ascended, and looked to my only escape route. I didn't harbor any delusions as far as my ability to get down the fire escape before Lung came over the top of the roof and deduced where I'd run off to. Worse, at that point he could probably just beat me to the street level by jumping off the roof, or even just shoot fire at me through the gaps in the metal while I was halfway down. The irony of the fire escape being anything but didn't escape me.

I wished I could fly. My school offered the choice between Chemistry, Biology and Physics, with Basic Science for the underachievers. I hadn't picked Physics, but I was still pretty sure that no matter how many I could gather together, jumping off the roof with a swarm of flying insects gripping me would be just as ineffective as the 9 year old superhero wannabes you heard about in the news, jumping off ledges with umbrellas and bedsheets.

For the time being, I was stuck where I was.

Reaching inside the convex armor that covered my spine, I ran my fingers over the things I had buckled in there. The EpiPens were meant to treat anaphylactic shock from allergic reactions to bee stings and the like, and likely wouldn't do a thing to Lung, even if I could get close enough and find a point to inject. Worst case scenario, the injections would supercharge his power by prompting a surge of whatever hormones or endorphins fueled his power. Not useful, dangerous at best. I had a pouch of chalk dust that was meant for climbers and gymnasts, I had seen it in the sports store when I was buying the lenses for my mask. I had gloves and didn't think I needed the dryness and extra traction, but I had gotten the idea that it could be useful to throw at an invisible enemy, and bought it on a whim. In retrospect, it had been kind of a dumb purchase, since my power let me find foes like that with my bugs. As a tool against Lung\ldots{} I wasn't sure if it would explode like regular dust could when exposed to flame, but fire didn't hurt him anyways. Scratch that option.

I tugged the little canister of pepper spray free from my armor. It was a black tube, three inches long, not much thicker around than a pen, with a trigger and a safety switch. It had been a gift from my dad, after I had started to go on my morning jogs for training. He had warned me to vary my route, and had given me the pepper spray for protection, along with a chain to clip it to my belt loop so it couldn't be taken and used against me by an attacker. In costume, I had opted not to keep the chain for the sake of moving quietly. Using my thumb, I flicked the safety off and positioned the tube so I was ready to fire. I crouched to make myself a smaller target, and waited for him to show himself.

Lung's hands, still on fire, were the first thing to show up, gripping the edge of the roof hard enough to bend the material that covered the roof's raised lip. His hands were quickly followed by his head and torso as he hauled himself up. He looked like he was made of overlapping knives or spades, smouldering yellow-orange with the low temperature flame. There was no skin to be seen, and he was easily seven or eight feet tall, judging by the length of his arms and torso. His shoulders alone were three feet across at the very least. Even the one eye that he had open looked metallic, a glowing, almond shaped pool of liquid-hot metal.

I aimed for the open eye, but the spray fired off at a sharp angle, just glancing off his shoulder. Where the spray grazed him, it ignited into a short lived fireball.

I swore under my breath and fumbled with the device. While he brought his leg over the edge, I adjusted my angle and shot again. This time -- with a small tweak of my aim mid-shot -- I hit him in the face. The ignited spray rolled off of him, but the contents still did the trick. He screamed, letting go of the roof with one hand, clutching the side of his face where his good eye was.

It had been vain to hope that he would slip and fall. I just counted myself lucky that however metallic his face looked, there were still parts of it vulnerable to the spray.

Lung hauled himself over the edge of the roof. I had him hurting\ldots{} I just couldn't do anything about it. My bugs were officially useless, there was nothing left in my utility sheath, and I would hurt myself more than I hurt Lung if I attacked him. Making a mental note to pick myself up a concealable knife or baton if I managed to live through this, I bolted for the fire escape.

\quotation{Muh\ldots{} Motherfucker!} Lung screamed. With my back turned, there was no way to see it, but the roof was briefly illuminated before the wave of flame hit me from behind. Knocked off balance, I skidded on the gravel and hit the raised lip of the roof, just by the fire escape. Frantically, I patted myself down. My costume wasn't on fire, but my hair -- I hurriedly ran my hands over it to make sure it hadn't been ignited.

Small mercies, I thought, that there was no tar used on the roof. I could just imagine the flames igniting the rooftop, and just how little I'd be able to do if it happened.

Lung stood, slowly, still covering part of his face with his hand. He walked with a slight limp as he approached me. Blindly, he lashed out with a broad wave of flame that rolled over half the roof. I covered my head with my hands and brought my knees to my chest as the hot air and flame rushed over me. My costume seemed to take the brunt of it, but it was still hot enough I had to bite my lip to stop from making a sound.

Lung stopped advancing, slowly turning his head from one side to another.

\quotation{Cock. Sucker,} he growled in his heavily accented voice, his cussing interrupted by his panting for breath, \quotation{Move. Give me something to aim for.}

I held my breath and stayed as still as possible. What could I do? I still had the pepper spray in my hand, but even if I got him again, I was running the risk that he would lash out and bake me alive before I could move. If I moved first, he would hear me and I would get knocked around by another blast of flame, probably before I could get to my feet.

Lung moved his hand from his face. He blinked a few times, then looked around, then blinked a few more times. It was a matter of seconds before he could see well enough to make me out from the shadows. Wasn't pepper spray supposed to put someone down for thirty minutes? How was this monster not an A-Lister?

He suddenly moved, flames wreathing his hands, and I screwed my eyes shut.

When I heard the crackling whoosh of the flame and wasn't burned alive, I opened my eyes again. Lung was firing streams of flame, aiming for the edge of the roof of the adjacent building, a three story apartment. I looked to see what he was aiming at, but couldn't make anything out in the gloom or in the brief second of light Lung's flames afforded.

With no warning, a massive shape landed atop Lung with an impact I could swear people heard at the other end of the street. The size of a van, the \quote{massive object} was animal rather than vehicle, resembling a cross between a lizard and a tiger, with tangles of muscle and bone where it ought to have skin, scales or fur. Lung was now on his knees, holding one of the beast's sizable claws away from his face with his own clawed hand.

Lung used his free hand to strike the creature across the snout. Even though he was smaller than the beast, the impact made it rear back. It took a few short steps back in reaction, and then rhino-charged him off the edge of the roof. They hit the street with an audible crash.

I stood, aware I was shaking like a leaf. I was so unsteady on my feet, from the mixed relief and fear, that I almost fell over again as two more impacts shook the roof. Two more creatures, similar to the first in texture, but slightly different in size and shape, had arrived on the rooftop. These two each had a pair of riders. I watched as the people slid off the backs of the animals. There were two girls, a guy, and a fourth I identified as male only because of the height. The tall one approached me, while the others hurried to the edge of the roof to watch Lung and the creature duke it out.

\quotation{You really saved us a lot of trouble,} he told me. His voice was deep, masculine, but muffled by the helmet he wore. He was dressed entirely in black, a costume I realized was basically motorcycle leathers and a motorcycle helmet. The only thing that made me think it was a costume was the visor of his helmet. The full-face visor was sculpted to look like a stylized skull, and was as black as the rest of his costume, with only the faint highlights of reflected light on the surface to give a sense of what it was. It was one of those costumes that people put together out of what they can scrounge up, and it wasn't half bad if you didn't look too close. He reached out a hand towards me, and I leaned away, wary.

I didn't know what to say, so I stuck to my policy of not saying anything that could get me into a worse situation.

Withdrawing his hand, the man in black jerked his thumb over one shoulder, \quotation{When we got word Lung was aiming to come after us tonight, we were pretty freaked. We were arguing strategy for the better part of the day. We eventually decided, fuck it, we'd meet him halfway. Wing it. Not my usual way of doing things, but yeah.}

Behind him, one of the girls whistled sharply and pointed down at the street. The two monsters the group had been riding on bounded across the roof and leaped down to the street to join the fight.

The guy in black kept talking, \quotation{Wouldn't you know, his flunky Lee is there with a half dozen guys, but Lung and the rest of his gang are nowhere to be found,} he laughed, a surprisingly normal sound for someone wearing a mask with a skull on it.

\quotation{Lee's no slouch in a fight, but there's a reason he's not leader of the ABB. He got spooked without his boss there and ran. I guess you're responsible for that?} Skull-mask waited for a response from me. When I didn't offer one, he ventured towards the edge of the roof and looked down, then spoke without turning to look at me, \quotation{Lung is getting creamed. The fuck you do to him?}

\quotation{Pepper spray, wasp and bee stings, fire ants and spider bites,} the second of the girls said, answering the question for me. She was dressed in a skintight outfit that combined black with a pale shade of blue or purple -- I couldn't tell in the dark -- and her dark blond hair was long and windblown. The girl grinned as she added, \quotation{He's not holding up too well. Gonna feel a helluvalot worse tomorrow.}

The man in black suddenly turned to look at me, \quotation{Introductions. That's Tattletale. I'm Grue. The girl with the dogs-} he pointed to the other girl, the one who had whistled and directed the monsters. She wasn't in costume unless I counted a plaid skirt, army boots, a torn-up sleeveless T-shirt and a hard plastic, dollar-store rottweiler mask as a costume. \quotation{-We call her Bitch, her preference, but in the interests of being P.G., the good guys and media decided to call her Hellhound instead. Last and certainly least, we have Regent.}

I finally caught up with what he was saying. Those monsters were {\em dogs}?

\quotation{Fuck you, Grue,} Regent retorted, with a chuckle and a tone of voice that made it clear he wasn't really that offended. He was wearing a white mask, not quite as decorative or made up as the ones I associated with the carnivals in Venice, but similar. He'd placed a silver coronet around his short black curls, and wore a ruffled white shirt with skintight leggings tucked into knee-high boots. The outfit was very renaissance faire. He had a build that made me think more of a dancer than a bodybuilder.

Introductions done, Grue looked at me for several long moments. After a few seconds, he asked me, \quotation{Hey, you okay? You hurt?}

\quotation{The reason she's not introducing herself isn't because she's hurt,} Tattletale told him, as she continued to lean over the edge of the roof and watch whatever was going on at the street level, \quotation{It's because she's shy.}

Tattletale turned around and it looked like she was going to say something else, but she stopped, turning her head. The smile she'd been wearing faded, \quotation{Heads up. We've gotta scram.}

Bitch nodded in response and whistled, one short whistle followed by two long ones. After a brief pause, the building was suddenly rattled by impacts. In just moments, the three creatures of hers leaped from the alleys to either side of the building and onto the roof.

Grue turned towards me. I was still standing on the opposite end of the roof, by the fire escape. \quotation{Hey, want a ride?}

I looked at the creatures -- dogs? They were bloodied, snarling creatures out of a nightmare. I shook my head. He shrugged.

\quotation{Hey,} Tattletale said to me, seating herself just behind Bitch, \quotation{What's your name?}

I stared at her. My voice caught in my throat before I was able to get the words out, \quotation{I don't\ldots{} I haven't picked one yet.}

\quotation{Well, Bug, a cape is gonna show up in less than a minute. You did us a solid by dealing with Lung, so take my advice. Someone from the Protectorate shows up, finds two bad guys duking it out, they're not going to let one walk away. You should get out of here,} She said. She flashed me a smile. She had one of those vulpine grins that turned up at the corners. Behind her simple black domino style mask, her eyes were glittering with mischief. If she had red hair, she would have made me think of a fox. She kind of did, anyways.

With that, they leaped over my head, one of the three beasts hitting or stepping on the fire escape on the way down, eliciting a screech of metal on metal.

When I realized what had just happened, I could have cried. It was easy enough to pin down Regent, Tattletale and Bitch as teenagers. It wasn't much of an intuitive leap to guess that Grue had been one too. The \quote{children} Lung had mentioned, the ones I had gone to so much effort to save tonight, were bad guys. Not only that, but they had mistaken me for one, too.


\chapter{Gestation 1.6}

I heard the cape arrive on his souped up motorcycle. I didn't want to be seen fleeing the scene of a fight, and risk being labeled one of the bad guys by {\em yet another} person, but I wasn't about to get closer to the street either, in case Lung was feeling better. Since there was nowhere to go, I just stayed put. Just resting felt good.

If you'd asked me just a few hours ago about how I thought I would feel meeting a big name superhero, I would have used words like excited and giddy. The reality was that I was almost too exhausted to care.

It looked as though he flew up onto the roof, but the six-foot long weapon the man held kind of jerked as he landed. I was pretty sure I saw the tines of a grappling hook retreating back into the end of the weapon. So this was what Armsmaster looked like in person, I thought.

The largest superhero organization in the world was the Protectorate, spanning Canada and the States, with ongoing talks about including Mexico in the deal. It was a government sponsored league of superheroes with a base in each \quote{cape city}. That is, they had a team set up in each city with a sizable population of heroes and villains. Brockton Bay's team was officially \quote{The Protectorate East-North-East}, and were headquartered in the floating, forcefield-shrouded island that you could see from the Boardwalk. This guy, Armsmaster, was the guy in charge of the local team. When the core group of the top Protectorate members from around Canada and the States assembled in that classic \quote{v} formation for the photo shoots, Armsmaster was one of the guys in the wings. This was a guy who had his own action figures. Poseable Armsmaster with interchangeable Halberd parts.

He really did look like a superhero, not like some guy in a costume. It was an important distinction. He wore body armor, dark blue with silver highlights, had a sharply angled v-shaped visor covering his eyes and nose. With only the lower half of his face exposed, I could see a beard trimmed to trace the edges of his jaw. If I had to judge, with only the lower half of his face to go by, I'd guess he was in his late twenties or early thirties.

His trademark and weapon was his Halberd, which was basically a spear with an axe head on the end, souped up with gadgets and the kind of technology you generally only saw in science fiction. He was the kind of guy who appeared on magazine covers and did interviews on TV, so you could find almost anything about Armsmaster through various media, short of his secret identity. I knew his weapon could cut through steel as though it was butter, that it had plasma injectors for stuff that the blade alone couldn't cut and that he could fire off directed electromagnetic pulses to shut down forcefields and mechanical devices.

\quotation{You gonna fight me?} He called out.

\quotation{I'm a good guy,} I said.

Stepping closer to me, he tilted his head, \quotation{You don't look like one.}

That stung, especially coming from him. It was like Michael Jordan saying you sucked at basketball. \quotation{That's\ldots{} not intentional,} I responded, not a little defensively, \quotation{I was more than halfway done putting the costume together when I realized it was already looking more edgy than I'd intended, and I couldn't do anything about it by then.}

There was a long pause. Nervously, I turned my eyes from that opaque visor. I glanced at his chest emblem, a silhouette of his visor in blue against a silver background, and was struck with the ridiculous thought that I had once owned a pair of underpants with his emblem on the front.

\quotation{You're telling the truth,} he said. It was a definitive statement, which startled me. I wanted to ask how he knew, but I wasn't about to do or say anything that might change his mind.

He approached closer, looking me over as I sat there with my arms around my knees, he asked, \quotation{You need a hospital?}

\quotation{No,} I said. \quotation{Don't think so. I'm as surprised as you are.}

\quotation{You're a new face,} he said.

\quotation{I haven't even come up with a name yet. You know how hard it is to come up with a bug-themed name that doesn't make me sound like a supervillain or a complete dork?}

He chuckled, and it sounded warm, very normal, \quotation{I wouldn't know. I got into the game early enough that I didn't have to worry about missing out on all of the good names.}

There was a pause in the conversation. I suddenly felt awkward. I don't know why, but I admitted to him, \quotation{I almost died.}

\quotation{That's why we have the Ward program,} he said. There was no judgement in his tone, no pressure. Just a statement.

I nodded, more to give a response than out of any agreement with the answer. The Wards were the under-eighteen subdivision of the Protectorate, and Brockton Bay did have its own team of Wards, with the same naming convention as the Protectorate; The Wards East-North-East. I had considered applying to join, but the notion of escaping the stresses of high school by flinging myself into a mess of teenage drama, adult oversight and schedules seemed self-defeating.

\quotation{You get Lung?} I asked, to change the subject from the Wards. I was pretty sure that he was obligated to try and induct new heroes into either the Protectorate or the Wards, depending on their age, to promote the whole agenda of organized heroes who are accountable for their actions, and I really didn't want him to get on my case about joining.

\quotation{Lung was unconscious, beaten and battered when I arrived. I pumped him full of tranquilizers to be safe and temporarily restrained him under a steel cage I welded to the sidewalk. I'll pick him up on my way back.}

\quotation{Good,} I said, \quotation{With him in jail, I'll feel like I accomplished something today. Only reason I started the fight was because I overheard him telling his men to shoot some kids. Only realized later that he was talking about some other villains.}

Armsmaster turned to look at me. So I told him, walking him through the fight in general, the arrival of the teenage bad guys, and their general descriptions. Before I finished, he was pacing back and forth on the roof.

\quotation{These guys. They knew I was coming?}

I nodded, once. As much respect as I had for Armsmaster, I wasn't in much of a mood to repeat myself.

\quotation{That explains a lot,} he said, staring off into the distance. After a few moments, he went on to explain, \quotation{They're slippery. On those few occasions we do manage to get in a toe to toe fight with them, they either win, or they get away more or less unscathed, or both. We know so little about them. Grue and Hellhound were working on their own before they joined the group, so there's some information there, but the other two? They're nonentities. If the girl Tattletale has some way of detecting or tracking us, it would go a long way towards explaining why they're doing as well as they are.}

It kind of surprised me to hear one of the top level heroes admitting to being anything less than perfectly on top of things.

\quotation{It's funny,} I said, after a few moment's thought, \quotation{They didn't seem that hardcore. Grue said they were kind of panicking when they heard Lung was coming after them, and they were casually joking around while the fight was going on. Grue was making fun of Regent.}

\quotation{They said all this in front of you?} he asked.

I shrugged, \quotation{I think they thought I was helping them out. The way Tattletale talked, I think she thought I was a bad guy too or something.} With a touch of bitterness, I said, \quotation{Dunno, I guess it was the costume that led them to that assumption.}

\quotation{Could you have taken them in a fight?} Armsmaster asked me.

I started to shrug, and winced a little. I was feeling a little sore in the shoulder, where I'd tumbled on the roof after being blasted by Lung's flames. I said, \quotation{Like you said, we don't know a lot about them, but I think that girl with the dogs-}

\quotation{Hellhound,} Armsmaster said.

\quotation{I think she could have kicked my ass on her own, so no. I probably couldn't have fought them.}

\quotation{Then count it as a good thing that they got the wrong impression,} Armsmaster said.

\quotation{I'll try to look at it that way,} I said, struck by how he easily he was able to employ the whole \quote{take a negative and turn it into a positive} mindset I'd been trying to maintain. I envied that.

\quotation{That a girl,} he said, \quotation{And while we're looking forward, we need to decide where we go from here.}

My heart sank. I knew he was going to bring up the Wards again.

\quotation{Who gets the credit for Lung?}

Caught off guard, I looked up at him. I started to speak, but he held up his hand.

\quotation{Hear me out. What you've done tonight is spectacular. You played a part in getting a major villain into custody. You just need to consider the consequences.}

\quotation{Consequences,} I muttered, even as the word {\em spectacular} rang in my ears.

\quotation{Lung has an extensive gang throughout Brockton Bay and neighboring cities. More than that, he has two superpowered flunkies. Oni Lee and Bakuda.}

I shook my head, \quotation{I know about Oni Lee, and Grue mentioned fighting him. I've never heard of Bakuda.}

Armsmaster nodded, \quotation{Not surprising. She's new. What we know about her is limited. She made her first appearance and demonstration of her powers by way of a drawn out terrorism campaign against Cornell University. Lung apparently recruited her and brought her to Brockton Bay after her plans were foiled by the New York Protectorate. This is\ldots{} something of a concern.}

\quotation{What are her powers?}

\quotation{Are you aware of the Tinker classification?}

I started to shrug, but remembered my sore shoulder and nodded instead. It was probably more polite, too. I said, \quotation{Covers anyone with powers that give them an advanced grasp of science. Lets them make technology years ahead of its time. Ray guns, ice blasters, mechanized suits of armor, advanced computers.}

\quotation{Close enough,} Armsmaster said. It struck me he would be a Tinker, if his Halberd and armor were any indication. That, or he got his stuff from someone else. He elaborated, \quotation{Well, most Tinkers have a specialty or a special trick. Something they're particualrly good at or something that they can do, which other Tinkers can't. Bakuda's specialty is bombs.}

I stared at him. A woman with a power that let her make bombs that were technologically decades ahead of their time. No wonder he saw it as a concern.

\quotation{Now I want you to consider the danger involved in taking the credit for Lung's capture. Without a doubt, Oni Lee and Bakuda will be looking to accomplish two goals. Freeing their boss and getting vengeance on the one responsible. I suspect you're now aware\ldots{} these are scary people. Scarier in some ways than their boss.}

\quotation{You're saying I shouldn't take the credit,} I said.

\quotation{I'm saying you have two options. Option one is to join the Wards, where you'll have support and protection in the event of an altercation. Option two is to keep your head down. Don't take the credit. Fly under the radar.}

I wasn't prepared to make a decision like that. Usually, I went to sleep at eleven or so, waking up at six thirty to get ready for my morning run. At my best guess, it was somewhere between one and two in the morning. I was emotionally exhausted from the highs and lows of the evening, and I could barely wrap my head around the complications and headaches that would come from joining the Wards, let alone having two insanely dangerous sociopaths coming after me.

On top of that, I wasn't so ignorant as to miss Armsmaster's motives. If I opted to not take the credit for Lung's capture, Armsmaster would, I was sure. I didn't want to get on the bad side of a major player.

\quotation{Please keep my involvement in Lung's capture secret,} I told him, painfully disappointed to have to say it, even as I knew it made the most sense.

He smiled, which I hadn't expected. He had a nice smile. It made me think that he could win the hearts of a lot of women, whatever the top two-thirds of his face looked like. \quotation{I think you'll look back and see this was a smart decision,} Armsmaster said, turning to walk to the other end of the roof, \quotation{Call me at the PHQ if you're ever in a pinch.} He stepped off the edge of the roof and dropped out of sight.

Call me if you're ever in a pinch. He'd been saying, without openly admitting, that he owed me one. He would take the lion's share of the credit for Lung's capture, but he owed me one.

Before I was all the way down the fire escape, I heard the thrum of his motorcycle, presumably carrying Lung towards a life of confinement. I could hope.

It would take me a half hour to get home. On the way, I would stop and pull on the sweatshirt and jeans I had hidden. I knew my dad went to sleep even earlier than I did, and he slept like a log, so I had nothing to worry about as far as wrapping up the night.

It could have gone worse. Strange as it sounds, those words were a security blanket I wrapped around myself to keep myself from dwelling on the fact that tomorrow was a school day.


\chapter{Interlude 1}

“We don't know how long he had been there. Suspended in the air above the Atlantic Ocean. On May twentieth, 1982, an ocean liner was crossing from Plymouth to Boston when a passenger spotted him. He was naked, his arms to his sides, his long hair blowing in the wind as he stood in the sky, nearly a hundred feet above the gently cresting waves. His skin and hair can only be described as a burnished gold. With neither body hair nor clothes to cover him, it is said, he seemed almost artificial.

“After a discussion including passenger and crew, the liner detoured to get closer. It was a sunny day, and passengers crowded to the railings to get a better look. As if sharing their curiosity, the figure drew closer as well. His expression was unchanging, but witnesses at the scene reported that he appeared deeply sad.

“\quote{I thought he was going to crack his facade and cry any moment}, said Grace Lands, \quote{But when I reached out and touched his fingertips, I was the one who burst into tears.}

“\quote{That boat trip was a final journey for me. I had cancer, and I wasn't brave enough to face it. Can't believe I'm admitting this in front of a camera, but I was going back to Boston, where I was born, to end things myself. After I met him, I changed my mind. Didn't matter anyways. I went to a doctor, and he said there was no sign I ever had the disease.}

“\quote{My brother, Andrew Hawke, was the last passenger to make any sort of contact with him, I remember. He climbed up onto the railing, and, almost falling off, he clasped the hand of the golden man. The rest of us had to grab onto him to keep him from falling. Whatever happened left him with a quiet awe. When the man with the golden skin flew away, my brother stayed silent. The rest of the way to Boston, my brother didn't say a word. When we docked, and the spell finally broke, my brother babbled his excitement to reporters like a child.}

“The golden man would reappear several more times in the coming months and years. At some point, he donned clothing. At first, a sheet worn over one shoulder and pinned at either side of the waist, then more conventional clothes. In 1999, he donned the white bodysuit he still wears today. For more than a decade, we have wondered, where did our golden man get these things? Who was he in contact with?

“Periodically at first, then with an increasing frequency, the golden man started to intervene in times of crisis. For events as small as a car accident, as great as natural disasters, he has arrived and used his abilities to save us. A flash of light to freeze water reinforcing a levee stressed by a hurricane. A terrorist act averted. A serial murderer caught. A volcano quelled. Miracles, it was said.

“His pace increased, perhaps because he was still learning what he could do, perhaps because he was getting a greater sense of where he was needed. By the middle of the 1990s, he was traveling from crisis to crisis, flying faster than the speed of sound. In fifteen years, he has not rested.

“He has been known to speak just once in thirty years. After extinguishing widespread fire in Alexandrovsk, he paused to survey the scene and be sure no blazes remained. A reporter spoke to him, and asked, \quote{Kto vy?} -- what are you?

“Shocking the world, caught on camera in a scene replayed innumerable times, he answered in a voice that sounded as though it might never have uttered a sound before. Barely audible, he told her, \quote{Scion}.

“It became the name we used for him. Ironic, because we took a word that meant descendant, and used it to name the first of many superpowered individuals -- parahumans -- to appear across Earth.

“Just five years after Scion's first appearance, the superheroes emerged from the cover of rumor and secrecy to show themselves to the public. Though the villains followed soon after, it was the heroes who shattered any illusions of the parahumans being divine figures. In 1989, attempting to quell a riot over a basketball game in Michigan, the superhero known to the public as Vikare stepped in, only to be clubbed over the head. He died not long after of a brain embolism. Later, he would be revealed to be Andrew Hawke.

\quotation{The golden age of the parahumans was thus short lived. They were not the deific figures they had appeared to be. Parahumans were, after all, people with powers, and people are flawed at their core. Government agencies took a firmer hand, and state-}

The television flicked off, and the screen went black, cutting the documentary off mid sentence. Danny Hebert sighed and sat down on the bed, only to stand just a moment later and resume pacing.

It was three fifteen in the morning, and his daughter Taylor was not in her bedroom.

Danny ran his hands through his hair, which was thinned enough at the top to be closer to baldness than not. He liked to be the first to arrive at work, watching everyone arrive, having them know he was there for them. So he usually went to bed early; he'd turn in at ten in the evening, give or take depending on what was on TV. Only tonight, a little past midnight, he'd been disturbed from restless sleep when he had felt rather than heard the shutting of the back door of the house, just below his bedroom. He had checked on his daughter, and he'd found her room empty.

So he had waited for his daughter to return for three hours.

Countless times, he had glanced out the window, hoping to see Taylor coming in.

For the twentieth time, he felt the urge to ask his wife for help, for advice, for support. But her side of the bed was empty and it had been for some time. Daily, it seemed, he was struck by the urge to call her cell phone. He knew it was stupid -- she wouldn't pick up -- and if he dwelt on that for too long, he became angry at her, which just made him feel worse.

He wondered, even as he knew the answer, why he hadn't gotten Taylor a cell phone. Danny didn't know what his daughter was doing, what would drive her to go out at night. She wasn't the type. He could tell himself that most fathers felt that way about their daughters, but at the same time, he knew. Taylor wasn't social. She didn't go to parties, she wouldn't drink, she wasn't even that interested in champagne when they celebrated the New Year together.

Two ominous possibilities kept nagging at him, both too believable. The first was that Taylor had gone out for fresh air, or even for a run. She wasn't happy, especially at school, he knew, and exercise was her way of working through it. He could see her doing it on a Sunday night, with a fresh week at school looming. He liked that her running made her feel better about herself, that she seemed to be doing it in a reasonable, healthy way. He just hated that she had to do it here, in this neighborhood. Because here, a skinny girl in her mid-teens was an easy target for attack. A mugging or worse -- he couldn't even articulate the worst of the possibilities in his own thoughts without feeling physically sick. If she had gone out at eleven in the evening for a run and hadn't come back by three in the morning, then it meant something had happened.

He glanced out the window again, at that corner of the house where the pool of illumination beneath the streetlight would let him see her approaching. Nothing.

The second possibility wasn't much better. He knew Taylor was being bullied. Danny had found that out in January, when his little girl had been pulled out of school and taken to the hospital. Not the emergency room, but the psychiatric ward. She wouldn't say by whom, but under the influence of the drugs they had given her to calm down, she had admitted she was being victimized by bullies, using the plural to give him a clue that it was a they and not a he or a she. She hadn't mentioned it -- the incident or the bullying -- since. If he pushed, she only tensed up and grew more withdrawn. He had resigned himself to letting her reveal the details in her own time, but months had passed without any hints or clues being offered.

There was precious little Danny could do on the subject, either. He had threatened to sue the school after his daughter had been taken to the hospital, and the school board had responded by settling, paying her hospital bills and promising they would look out for her to prevent such events from occurring in the future. It was a feeble promise made by a chronically overworked staff and it didn't do a thing to ease his worries. His efforts to have her change schools had been stubbornly countered with rules and regulations about the maximum travel times a student was allowed to have between home and a given school. The only other school within a reasonable distance of Taylor's place of residence was Arcadia High, and it was already desperately overcrowded with more than two hundred students on a list requesting admittance.

With all that in mind, when his daughter disappeared until the middle of the night, he couldn't shake the idea that the bullies might have lured her out with blackmail, threats or empty promises. He only knew about the one incident, the one that had landed her in the hospital, but it had been grotesque. It had been implied, but never elaborated on, that more had been going on. He could imagine these boys or girls that were tormenting his daughter, egging one another on as they came up with more creative ways to humiliate or harm her. Taylor hadn't said as much aloud, but whatever had been going on had been mean, persistent and threatening enough that Emma, Taylor's closest friend for years, had stopped spending time with her. It galled him.

Impotent. Danny was helpless where it counted. There was no action he could take -- his one call to the police at two in the morning had only earned him a tired explanation that the police couldn't act or look for her without something more to go on. If his daughter was still gone after twelve hours, he'd been told, he should call them again. All he could do was wait and pray with his heart in his throat that the phone wouldn't ring, a police officer or nurse on the other end ready to tell him what had happened to his daughter.

The slightest of vibrations in the house marked the escape of the warm air in the house to the cold outdoors, and there was a muffled whoosh as the kitchen door shut again. Danny Hebert felt a thrill of relief coupled with abject fear. If he went downstairs to find his daughter, would he find her hurting or hurt? Or would his presence make things worse, her own father seeing her at her most vulnerable after humiliation at the hands of bullies? She had told him, in every way except articulating it aloud, that she didn't want that. She had pleaded with him, with body language and averted eye contact, unfinished sentences and things left unsaid, not to ask, not to push, not to see, when it came to the bullying. He couldn't say why, exactly. Home was an escape from that, he'd suspected, and if he recognized the bullying, made it a reality here, maybe she wouldn't have that relief from it. Perhaps it was shame, that his daughter didn't want him to see her like that, didn't want to be that weak in front of him. He really hoped that wasn't the case.

So he ran his fingers through his hair once more and sat down on the corner of the bed, elbows on his knees, hands on his head, and stared at his closed bedroom door. His ears were peeled for the slightest clue. The house was old, and it hadn't been a high quality building when it had been new, so the walls were thin and the structure prone to making noise at every opportunity. There was the faintest sound of a door closing downstairs. The bathroom? It wouldn't be the basement door, with no reason for her to go down there, and he couldn't imagine it was a closet, because after two or three minutes, the same door opened and closed again.

After something banged on the kitchen counter, there was little but the occasional groan of floorboards. Five or ten minutes after she had come in, there was the rhythmic creak of the stairs as she ascended. Danny thought about clearing his throat to let her know he was awake and available should she knock on his door, but decided against it. He was being cowardly, he thought, as if his clearing of his throat would give reality to his fears.

Her door shut carefully, almost inaudibly, with the slightest tap of door on doorframe. Danny stood, abruptly, opening his door, ready to cross the hall and knock on her door. To verify that his daughter was okay.

He was stopped by the smell of jam and toast. She had made a late night snack. It filled him with relief. He couldn't imagine his daughter, after being mugged, tormented or humiliated, coming home to have toast with jam as a snack. Taylor was okay, or at least, okay enough to be left alone.

He let out a shuddering sigh of relief and retreated to his room to sit on the bed.

Relief became anger. He was angry at Taylor, for making him worry, and then not even going out of her way to let him know she was okay. He felt a smouldering resentment towards the city, for having neighborhoods and people he couldn't trust his daughter to. He hated the bullies that preyed on his daughter. Underlying it all was frustration with himself. Danny Hebert was the one person he could control in all of this, and Danny Hebert had failed to do anything that mattered. He hadn't gotten answers, hadn't stopped the bullies, hadn't protected his daughter. Worst of all was the idea that this might have happened before, with him simply sleeping through it rather than laying awake.

He stopped himself from walking into his daughter's room, from shouting at her and demanding answers, even if it was what he wanted, more than anything. Where had she been, what had she been doing? Was she hurt? Who were these people that were tormenting her? He knew that by confronting her and getting angry at her, he would do more harm than good, would threaten to sever any bond of trust they had forged between them.

Danny's father had been a powerful, heavyset man, and Danny hadn't gotten any of those genes. Danny had been a nerd when the term was still young in popular culture, stick thin, awkward, short sighted, glasses, bad fashion sense. What he had inherited was his father's famous temper. It was quick to rise and startling in its intensity. Unlike his father, Danny had only ever hit someone in anger twice, both times when he was much younger. That said, just like his father, he could and would go off on tirades that would leave people shaking. Danny had long viewed the moment he'd started to see himself as a man, an adult, to be the point in time where he had sworn to himself that he wouldn't ever lose his temper with his family. He wouldn't pass that on to his child the way his father had to him.

He had never broken that oath with Taylor, and knowing that was what kept him contained in his room, pacing back and forth, red in the face and wanting to punch something. While he'd never gotten angry at her, never screamed at her, he knew Taylor had seen him angry. Once, he had been at work, talking to a mayor's aide. The man had told Danny that the revival projects for the Docks were being cancelled and that, contrary to promises, there were to be layoffs rather than new jobs for the already beleaguered Dockworkers. Taylor had been spending the morning in his office on the promise that they would go out for the afternoon, and had been in a position to see him fly off the handle in the worst way with the man. Four years ago, he had lost his temper with Annette for the first time, breaking his oath to himself. That had been the last time he had seen her. Taylor hadn't been there to see him shouting at her mother, but he was fairly certain she'd heard some of it. It shamed him.

The third and last time that he had lost his temper where Taylor had been in a position to know had been when she had been hospitalized following the incident in January. He'd screamed at the school's principal, who had deserved it, and at Taylor's then-Biology teacher, who probably hadn't. It had been bad enough that a nurse had threatened to call for a police officer, and Danny, barely mollified, had stomped from the hallway to the hospital room to find his daughter more or less conscious and wide eyed in reaction. Danny harbored a deep fear that the reason Taylor hadn't offered any details on the bullying was out of fear he would, in blind rage, do something about it. It made him feel sick, the notion that he might have contributed something to his daughter's self imposed isolation in how she was dealing with her problems.

It took Danny a long time to calm down, helped by telling himself over and over that Taylor was okay, that she was home, that she was safe. It was something of a blessing that, as the anger faded, he felt drained. He climbed into the left side of the bed, leaving the right side empty out of a habit he'd yet to break, and pulled the covers up around himself.

He would talk to Taylor in the morning. Get an answer of some sort.

He dreamed of the ocean.
