% Arc 2: Insinuation

\part{Insinuation}

\chapter{Insinuation 2.1}

I woke to the muffled sound of the radio in the bathroom. Reaching over to my alarm clock, I turned it around. 6:28. Which made today a weekday like any other. My alarm was set for six thirty, but I almost never needed it, because my dad was always in the shower at the same time. Routines defined us.

As a wave of fatigue swept over me, I wondered if I might be sick. It took me a few moments of staring up at the ceiling to remember the events of last night. Small wonder I was tired. I had gotten home, snuck inside and gone to bed at close to three thirty, just three hours ago. With all that had happened, I hadn't slept those full three hours, either.

I forced myself out of bed. As a slave to my routine, it would be wrong to do otherwise. I made myself change into sweats and walk down to the kitchen sink to wash my face, fighting to keep awake. I was sitting at the kitchen table, pulling on my sneakers, when my dad came downstairs in his bathrobe.

My dad is not what you'd call an attractive man. Beanpole thin, weak chin, thinning dark hair that was on the cusp of baldness, big eyes and glasses that magnified those eyes further. As he entered the kitchen, he looked surprised to see me there. That's just the way my dad always looked: constantly bewildered. That, and a little defeated.

\quotation{Good morning, kiddo,} he said, entering the kitchen and leaning down to kiss the crown of my head.

\quotation{Hey, dad.}

He was already stepping towards the fridge as I replied. He looked over his shoulder, \quotation{A little glum?}

\quotation{Hunh?}

\quotation{You sound down,} he said.

I shook my head, \quotation{Tired. I didn't sleep well.}

There was the slap of bacon hitting the frying pan. It was sizzling by the time he spoke, \quotation{You know, you could go back to bed, sleep in for another hour or so. You don't have to go on your run.}

I smiled. It was equal parts annoying and sweet, that my dad hated me running. He worried about my safety, and couldn't turn down a chance to drop hints that I should stop, or be safer, or join a gym. I wasn't sure if he'd worry more or less if I told him about my powers.

\quotation{You know I do, dad. If I don't go today, it'll be that much harder to make myself get up and do it tomorrow.}

\quotation{You've got the, uh\ldots{}}

\quotation{I've got the tube of pepper spray in my pocket,} I said. He bobbed his head in acknowledgement. It was only moments later that I realized I didn't have it. The pepper spray was with my costume, in the coal chute in the basement. I felt a pang of guilt at realizing I'd lied to my dad.

\quotation{O.J.?} he asked.

\quotation{I'll get it,} I said, heading to the fridge for the orange juice. While I was at the fridge, I also grabbed some applesauce. As I returned to the table, my dad slapped some french toast on the frying pan to join the bacon. The room filled with the aroma of the cooking food. I helped myself to the applesauce.

\quotation{You know Gerry?} my dad asked.

I shrugged.

\quotation{You met him once or twice when you've visited me at work. Big guy, burly, Black Irish?}

Shrugging again, I took a bite of french toast. My dad was part of the Dockworkers Association, as the Union spokesperson and head of hiring. With the state of the Docks being what they were, that meant my dad was pretty much in charge of telling everyone that there were no jobs to be had, day after day.

\quotation{Rumor's going around he found work. Guess with who.}

\quotation{Dunno,} I said, around a mouthful of food.

\quotation{He's going to be one of Über and Leet's henchmen.}

I raised my eyebrows. Über and Leet were local villains with a video game theme. They were pretty much as incompetent as villains could be while staying out of jail. They barely even rated as B-list.

\quotation{They going to make him wear a uniform? Bright primary colors, Tron style?}

My dad chuckled, \quotation{Probably.}

\quotation{We're supposed to talk about how the powers thing has influenced our lives in class today. Maybe I'll mention that.}

We ate in silence for a minute or two.

\quotation{I heard you come in late last night,} he said.

I just gave him a small nod and took another bite of french toast, even as my heart rate tripled and my mind searched for excuses.

\quotation{Like I said,} I finally opened my mouth, looking down at my plate, \quotation{I just couldn't sleep. I couldn't get my thoughts to settle down. I got out of bed and tried pacing, but it didn't help, so I stepped outside and walked around the neighborhood.} I wasn't totally lying. I'd had nights like that. Last night just hadn't been one of them, and I had gone walking around the neighborhood, even if it was in a different way than I'd implied.

\quotation{Christ, Taylor,} my father answered, \quotation{This isn't the kind of area where you can walk around in the middle of the night.}

\quotation{I had the pepper spray,} I protested, lamely. That wasn't a lie, at least.

\quotation{What if you get caught off guard? What if the guy has a knife, or a gun?} my father asked.

Or pyrokinesis and the ability to grow armor plating and claws? I felt a little knot of ugliness in the pit of my stomach at my father's concern for me. It was all the more intense because it was so justified. I had almost died last night.

\quotation{What's going on, that has you so anxious you can't sleep?} he questioned me.

\quotation{School,} I said, swallowing around a lump in my throat, \quotation{Friends, the lack thereof.}

\quotation{It's not better?} he asked, carefully stepping around the elephant in the room, the bullies.

If it was, I wouldn't be having problems, would I? I just gave him a one shoulder shrug and forced myself to take another bite of french toast. My shoulder twinged a little as it made the bruises from last night felt. As much as I didn't feel like eating, I knew my stomach would be growling at me before lunch if I didn't. That was even without accounting for the energy I burned running, let alone the escapades of last night.

When my dad realized I didn't have an answer for him, he resumed eating. He only had one bite before he put his fork down again with a clink on the plate.

\quotation{No more going out in the middle of the night,} he said, \quotation{Or I'm putting a bell on the doors.}

He would, too. I just nodded and promised myself I would be more careful. When I had come in, I had been so tired and sore that I hadn't given any thought to the click of the door, the rattle of the lock or the creaks of floorboards that were older than me.

\quotation{Okay,} I said, adding, \quotation{I'm sorry.} Even with that, I felt a twinge of guilt. My apology was sincere in feeling, but I was making it with the knowledge that I would probably do the same thing again. It felt wrong.

He gave me a smile that seemed almost like an unspoken \quote{I'm sorry too}.

I finished off my plate and stood up to put it in the sink and run water over it.

\quotation{Going on your run?}

\quotation{Yeah,} I said, put my dishes in the beaten up old dishwasher and bent down to give my dad a hug on my way to the door.

\quotation{Taylor, have you been smoking?}

I shook my head.

\quotation{Your hair is, uh, burnt. At the ends, there.}

I thought back to the previous night. Getting hit in the back by one of Lung's blasts of flame.

Shrugging, I suggested, \quotation{Stove, maybe?}

\quotation{Be safe,} my dad said, emphasizing each word. I took that as my cue to go, heading out the side door and breaking into an all out run the moment I was past the chain link gate at the side of the house.


\chapter{Insinuation 2.2}

The run had helped to wake me up, as did the hot shower and a cup of the coffee my dad had left in the pot. Even so, the fatigue didn't help the feeling of disorientation over just how normal the day seemed as I made my way to school. Just a matter of hours ago, I had been in a life and death fight, I had even met Armsmaster. Now it was a day like any other.

I felt a bit nervous as I got to homeroom. Having basically skipped two classes the previous Friday, failing to turn in a major assignment, I figured that Mrs. Knott probably knew already. I didn't feel relieved when Mrs. Knott glanced up at me and gave a tight smile before turning her attention back to her computer. That just meant the humiliation would be redoubled if and when class was interrupted by someone coming down from the office. A part of me just wanted to miss this class too, just to avoid the potential humiliation and avoid drawing attention.

All in all, I felt anxious as I made my way to my computer, which kind of sucked because Computer class was one of the few parts of the school day I didn't usually dread. For one thing, it was the one class in which I was doing well. More to the point, neither Madison, Sophia nor Emma were in this class, though some of their friends were. Those girls didn't usually feel the need to harass me without the trio around, and I was further removed from them because I was in the advanced stream of the class. A good three quarters of the people in the room were computer illiterate, being from families that didn't have the money for computers or families that didn't have much interest in the things, so they practiced typing without looking at the keyboard and had lessons in using search engines. By contrast, I was in the group that was learning some basic programming and spreadsheets. It didn't do a lot for my already geeky reputation, but I could deal.

Mrs. Knott was an alright teacher, if not the most hands on; she was usually content to give us advanced students an in-class assignment and then focus on the more rambunctious majority for the rest of the class. This suited me just fine -- I usually wrapped up the assignment in a half hour, leaving me an hour to use as I saw fit. I had been recalling and going over the events of the previous night during my morning run, and the first thing that I did when the ancient desktop finished its agonizing load process was to start digging for information.

The go-to place for news and discussion on capes was Parahumans Online. The front page had constant updates on recent, international news featuring capes. From there, I could go to the wiki, where there was information on individual capes, groups and events, or to the message boards, which broke down into nearly a hundred sub-boards, for specific cities and capes. I opened the wiki in one tab, then found and opened the message board for Brockton Bay in another.

I had the sense that either Tattletale or Grue were the leader of the group I had run into. Turning my attention to Tattletale, I searched the wiki. The result I got was disappointingly short, starting with a header reading \quotation{This article is a stub. Be a hero and help us expand it.} There was a one sentence blurb on how she was a alleged villain active in Brockton Bay, with a single blurry picture. The only new information for me was that her costume was lavender. A search of the message boards turned up absolutely nothing. There wasn't even a hint as to what her power was.

I looked up Grue. There was actually information about him, but nothing detailed or definitive. The wiki stated he had been active for nearly three years, dealing in petty crimes such as robbing small stores and doing some work as an enforcer for those who wanted a little superpowered muscle along for a job. Recently, he had turned to higher scale crime, including corporate theft and robbing a casino, together with his new team. His power was listed as darkness generation in the sidebar under his picture. The picture seemed crisp enough, but the focus of it, Grue, was just a blurry black silhouette in the center.

I searched for Bitch, next. No results. I did another search for her more official title, Hellhound, and got a wealth of information. Rachel Lindt had never made any real attempt to hide her identity. She had apparently been homeless through most of her criminal career, just living on the streets and moving on whenever police or a cape came after her. The sightings and encounters with the homeless girl ended around a year ago -- I figured that was when she joined forces with Grue, Tattletale and Regent. The picture in the sidebar was taken from surveillance camera footage -- an unmasked, dark haired girl who I wouldn't have called pretty. She had a squarish, blunt-featured face with thick eyebrows. She was riding atop one of her monstrous \quote{dogs} like a jockey rides a horse, down the middle lane of a street.

According to the wiki entry, her powers manifested when she was fourteen, followed almost immediately by her demolishing the foster home she had been living in, injuring her foster mother and two other foster children in the process. This was followed by a two year series of skirmishes and retreats across Maine as various heroes and teams tried to apprehend her, and she either defeated them or successfully evaded capture. She had no powers that would have made her any stronger or faster than the average Jane, but she was apparently able to turn ordinary dogs into the creatures I had seen on the rooftop. Monsters the size of a car, all muscle, bone, fang and claw. A red box near the bottom of the page read, \quotation{Rachel Lindt has a public identity, but is known to be particularly hostile, antisocial and violent. If recognized, do not approach or provoke. Leave the area and notify authorities as to her last known location.} At the very bottom of the page was a list of links that were related to her: two fansites and a news article relating to her early activities. A search of the message boards turned up too many results, leaving me unable to sift through the crap, the arguments, the speculation and the villain worship to find any genuine morsels of information. If nothing else, she was notorious. I sighed and moved on, making a mental note to do more investigation when I had the time.

The last member of the group was Regent. Given what Armsmaster had said about the guy being low profile, I didn't expect to find much. I was surprised to find less than that. Nothing. My search on the wiki turned up only a default response, \quotation{There are no results matching this query. 32 unique IP addresses have searched the Parahumans.net Wiki for \quote{Regent} in 2011. Would you like to create the page?} The message board didn't turn up anything else. I even did a search for alternate spellings of his name, such as Regence and Recant, in case I had heard it wrong. Nothing turned up.

If my mood had been on the sour side as I got to homeroom, the dead ends only made it worse. I turned my attention to the in-class assignment, making a working calculator in Visual Basic, but it was too trivial to distract me. The work from Thursday and Friday had already given us the tools to do the job, so it was really just busywork. I didn't mind learning stuff, but work for the sake of doing work was annoying. I did the bare minimum, checked it for any bugs, moved the file to the \quote{completed work} folder and returned to surfing the web. All in all, the work barely took fifteen minutes.

I looked up Lung on the wiki, which I had done often enough before, as part of my research and preparation for being a superhero. I'd wanted to be sure I knew who prominent local villains were and what they could do. The search for \quote{Lung} redirected to a catch-all page on his gang, the ABB, with quite a bit of detailed information. The information on Lung's powers was pretty in line with my own experience, though there was no mention of the super-hearing or him being fireproof. I debated adding it, but decided against it. There were security concerns with my submission being tracked back to Winslow High, and then to me. I figured it would probably be deleted as unsupported speculation, anyways.

The section beneath the description of Lung and his powers covered his subordinates. He was estimated to have forty or fifty thugs working for him across Brockton Bay, largely drawn from the ranks of Asian youth. It was pretty unconventional for a gang to include members of the variety of nationalities that the ABB did, but Lung had made it a mission to conquer and absorb every gang with Asian members and many without. Once he had the manpower he needed, the non-Asian gangs were cannibalized for assets, their members discarded. Even though there were no more major gangs in the east end of town to absorb, he was still recruiting zealously. His method, now, was to go after anyone older than twelve and younger than sixty. It didn't matter if you were a gang member or not. If you were Asian and you lived in Brockton Bay, Lung and his people expected you to either join or to pay tribute one way or another. There had been local news reports on it, newspaper articles, and I could remember seeing signs in the guidance counselor's office detailing where people who were targeted in this way could go for help.

Lung's lieutenants were listed as Oni Lee and Bakuda. I already had some general knowledge about Oni Lee, but I was intrigued to see there were recent updates to his wiki entry. There were specific details on his powers: He could teleport, but when he did so, he didn't disappear. As he teleported, his original self, for lack of a better term, would stay where it was and remain active for five to ten seconds before disintegrating into a cloud of carbon ash. Essentially, he could create another version of himself anywhere nearby, while the old version could stick around long enough to distract or attack you. If that wasn't scary enough, there was an report of him holding a grenade in his hand as he repeatedly duplicated himself, with his short lived duplicates acting as suicide bombers. Topping it all off, Oni Lee's wiki page had a similar red warning box to the one that Bitch/Hellhound had on hers, minus the bit about his public identity. From what they knew about him, authorities had seen fit to note him a sociopath. The warning covered the same essential elements: exceedingly violent, dangerous to approach, should not be provoked, and so on. I glanced at his picture. His costume consisted of a black bodysuit with a black bandoleer and belt for his knives, guns and grenades. The only color on him was an ornate Japanese-style demon mask, crimson with two green stripes down either side. Except for the mask, his costume gave off the distinct impression of a ninja, which just added weight to the notion that this was a guy who could and would slide a knife between your ribs.

Bakuda was a new entry, added to the ABB wiki page just ten days ago. The picture only showed her from the shoulders up, a girl with straight black hair, large opaque goggles over her eyes and a metal mask with a gas mask styled filter covering the lower half of her face. A braided cord of black, yellow and green wires looped over one of her shoulders. I couldn't pinpoint her ethnicity with the mask and goggles, and her age wasn't any easier to figure out.

The wiki had a lot of the same details Armsmaster had mentioned to me. Bakuda had essentially held a university ransom and she did it with her superhuman ability to design and fabricate high tech bombs. There was a link to a video titled \quote{Bomb Threat @ Cornell}, but I didn't think it wise to play it in school, especially without headphones. I made a mental note to check it out when I got home.

The next thing that caught my eye was the section heading titled \quote{Defeats and Captures}. I scrolled down to read it. According to the wiki, Lung had apparently suffered a number of minor defeats at the hands of various teams, ranging from the Guild to the local teams of New Wave, the Wards and the Protectorate, but consistently managed to evade capture until last night. A blurb read, \quote{ Armsmaster successfully ambushed and defeated the leader of the ABB, who was weakened from a recent encounter with a rival gang. Lung was taken to the PHQ for holding until the villain's trial by teleconference. Given Lung's extensive and well documented criminal history, it is expected he will face imprisonment in the Birdcage should he be found guilty at trial.}

I took a deep breath and let it out slowly. I wasn't sure what to think. By all rights, I should have been angry that Armsmaster took the credit for the fight that could have cost me my life. Instead, I felt a building excitement. I felt like shaking the shoulder of the guy sitting next to me and point to the screen, saying, \quotation{Me, I made that possible! Me!}

With a renewed enthusiasm, I switched tabs to the message board and began looking to see what people were saying about it. A post by a fan or minion of Lung threatened violence against Armsmaster. There was a request by someone asking for more information on the fight. I was given pause by one post that asked whether Bakuda could or would use a large scale bomb and the threat of potentially thousands or hundreds of thousands dead, to ransom Lung back.

I tried to put that out of my mind. If it happened, it would be the responsibility of heroes better and more experienced than I.

It struck me that there was one person I hadn't looked for. Myself. I opened up the advanced search page for the Parahumans.net message board and did a search for multiple terms. I included insect, spider, swarm, bug, plague, and a mess of other terms that had struck me when I had been trying to brainstorm a good hero name. I narrowed the timeframe of posts to search for posts made within the past 12 hours and hit Search.

My efforts turned up two posts. One referred to a villain called Pestilence, active in the UK. Apparently Pestilence was one of the people who could use \quote{magic}. That is, he was if you believed magic was real, and not just some convoluted or deluded interpretation of a given set of powers.

The second post was in the \quote{Connections} section of the message board, where rescued damsels left their contact information for their dashing heroes, where conventions and fan gatherings were organized and where people posted job offers for capes and the cape-obsessed. Most were cryptic or vague, referring to stuff only the people in question would know.

The message was titled, simply, \quotation{Bug}

I clicked it and waited impatiently for the outdated system and overloaded school modem to load up the page. What I got was brief.

{\it Subject: Bug

Owe you one. Would like to repay the favor. Meet?

Send a message,

Tt.}

The post was followed by two pages of people commenting. Three people suggested it was something important, while a half dozen more people decried them as tinfoil hats, Parahumans.net's term for conspiracy theorists.

It was meaningful, though. I couldn't interpret it any other way; Tattletale had found a way to get in contact with me.


\chapter{Insinuation 2.3}

I didn't have any time to contemplate the message I'd received from Tattletale. The bell rang and I had to hurry to properly log off and shut down before heading to my next class. As I gathered my stuff, I realized I had been so caught up in researching on the villains I'd met last night and in Tattletale's message that I had forgotten to worry about getting into trouble for skipping class. I felt a kind of resignation as I realized I would have to face the music later in the day, anyways.

Madison was already in her seat as I got to the classroom. She had a pair of girls crouching by either side of her desk, and all three of them broke into giggles as they saw me. Bitches.

My seat of choice was the far right, front row, closest to the door. Lunch hour and immediately after school was when the trio tended to give me the hardest time, so I tried to sit as close as possible to the door, for a quick escape. I spotted a puddle of orange juice on the seat, with the empty plastic bottle lying just underneath the chair. Madison was going for a two for one. It was both a \quote{prank} and a reminder of how they had doused me with juice and soft drinks last Friday. Irritated, I carefully avoided looking at Madison and took an empty seat a few rows back.

Mr Gladly entered the room, he was short and young enough you could almost mistake him for another high school student. It took a few minutes for him to start the class, and he immediately ordered us to break into groups of four to share our homework with one another and to prepare to share it with the rest of the class. The group that had the most to contribute would win the prize he had mentioned on Friday, treats from the vending machine.

It was stuff like this that made Mr. Gladly my least favorite teacher. I got the impression he'd be surprised to hear he was {\em anyone's} least favorite teacher, but that was just one more point against him in my book. I don't think he comprehended why people might not like him, or how miserable group work was when you didn't identify with any of the groups or cliques in the school. He just figured people liked doing group work because it let them talk and hang out with their friends in class.

While the class got sorted, I figured I'd avoid standing around like a loser with no group to join and get something else out of the way. I approached the desk at the front of the room.

\quotation{Mr. Gladly?}

\quotation{Call me Mr. G. Mr. Gladly is my dad,} he informed me with a sort of mock sternness.

\quotation{Sorry, uh, Mr. G. I need a new textbook.}

He gave me a curious look, \quotation{What happened to your old one?}

Soaked with grape juice by a trio of harpies. \quotation{I lost it,} I lied.

\quotation{Replacement textbooks are thirty five dollars. I don't expect it now, but\ldots{}}

\quotation{I'll have it for you by the end of the week,} I finished for him.

He handed me a textbook, and I looked over the room before joining the only group with room for more: Sparky and Greg. We had been in a group several times before, as the leftovers when all the friends and cliques had banded together.

Sparky had apparently picked up his nickname when a third grade teacher used it in an ironic sense, and it had stuck, to the point where I doubted anyone but his own mother even knew his real name. He was a drummer, long haired, and was so out of touch with reality that you could stop talking in the middle of a sentence and he wouldn't notice. He just went through life in a daze, presumably until he could do his thing, which was his band.

Greg was just the opposite. He was smarter than average, but he had a way of saying every thought that came into his head -- his train of thought didn't have any brakes. Or tracks. It would have been easier to be in a group with just Sparky and essentially do the work by myself than it would be to work with Greg.

I got my share of the homework out of my new backpack. Mr. Gladly had asked us to come up with a list of ways that capes had influenced society. In between the various steps of my getting ready for my first night out in costume, I had taken the time to fix up my art project and had come up with a fairly comprehensive list for Mr. Gladly's homework. I had even used newspaper and magazine clippings to support my points. I felt pretty good about it.

\quotation{I didn't get much done,} Greg said, \quotation{I got distracted by this new game I got and it is really really good, it's called Space Opera, have you played it?}

A full minute later he was still on the same topic, even though I wasn't playing any attention to him or giving him any feedback on what he was saying, \quotation{\ldots{}you have to understand it's a genre, and it's one I've really been getting into it lately, since I started watching this anime called -- Oh, hey, Julia!} Greg broke off from his monologue to wave with enough energy and excitement that I felt a little embarrassed to just be sitting next to him. I turned in my seat to see one of Madison's friends coming in, late.

\quotation{Can I be in Madison's group?} Julia asked Mr. Gladly.

\quotation{That wouldn't be fair. Greg's group only has three people. Help them,} Mr. Gladly said.

Julia walked over to where we were sitting and made a face. Just loud enough for us to hear, she muttered a disgusted, \quotation{Ew.} I felt much the same about her joining us.

It was downhill from there. Madison's group moved so the four of them were sitting right next to our group, which let Julia talk with them while still sitting with us. The presence of all the popular and attractive girls in the class just got Greg more wound up, and he began trying to insert himself into their conversation, only to get shut down or ignored. It was embarrassing to watch.

\quotation{Greg,} I said, trying to distract him from the other group, \quotation{Here's what I did over the weekend. What do you think?}

I handed him the work I had done. To his credit, he gave it a serious read.

\quotation{This is really good, Taylor,} He said, when he was done.

\quotation{Let me see,} Julia said. Before I could stop him, Greg dutifully handed my work over to her. I watched her glance over it, then toss it onto Madison's table. There were a few giggles.

\quotation{Give that back,} I said.

\quotation{Give what back?} Julia said.

\quotation{Madison,} I said, ignoring Julia, \quotation{Give it back.}

Madison, cute and petite and crush of choice for half the guys in our grade, turned and managed a combined look and tone of such condescension that a grown man would have flinched, \quotation{Nobody is talking to you, Taylor.}

That was that. Short of running to the teacher and complaining, I wasn't going to get my work back, and anyone who considered that an option has clearly never been in high school. Greg looked between me and the girls with a kind of panic before settling into a funk, Sparky had his head down on his desk, either asleep or close to it, and I was left fuming. I made an attempt at trying to to salvage things, but getting Greg to focus was impossible, as he constantly tried to apologize and made lame attempts to convince the other group to give my work back. Our time ran out, and Mr. Gladly picked out people from each group to stand up and go over what they had come up with.

I sighed as Mr. Gladly picked Greg to do our group's presentation, and was forced to watch Greg botch it badly enough that Mr. Gladly asked him to sit down before he was finished. Greg was one of those kids I always figured made teachers groan inwardly when they raised their hands in class. The sort of kid that took twice as long to answer as anyone else, and was often only half-right or so off-tangent that it derailed the discussion. I couldn't imagine what had possessed Mr. Gladly to pick Greg to do our group's presentation.

What made things worse was that I then got to watch Madison rattle off {\em my} very impressive sounding list of ways capes had changed the world. She cribbed almost all of my stuff; fashion, economics, Tinkers and the tech boom, the fact that movies, television and magazines had been tweaked to accommodate cape celebrities, and so on. Still, she got it wrong when explaining how law enforcement had changed. My point had been that with qualified capes easing the workload and taking over for most high profile crises, law enforcement of all stripes were more free to train and expand their skill sets, making for smarter, more versatile cops. Madison just made it sound like they got a lot of vacation days.

Mr. Gladly named another group as the winners, by virtue of the sheer number of things they had come up with, though he made a point of saying the quality of Madison's work was nearly good enough to count. From there, he moved on to his lecture.

I was steamed and I could hardly focus on the lecture, as my power crackled and tugged at my attention from the periphery of my consciousness, making me acutely aware of every bug within a tenth of a mile. I could tune it out, but the extra concentration that took, coupled with the anger I felt towards Madison and Mr. Gladly, was distracting enough that I couldn't focus on the lecture. I took a cue from Sparky and put my head down on the desk. Being as exhausted from the previous night's activity as I was, it was all I could do to keep from dozing off. Still, spending the class half asleep made it go by faster. I was startled when the bell rang.

As everyone gathered their things and began to file out, Mr. Gladly approached me and quietly said, \quotation{I'd like you to stick around for a few minutes, please.}

I just nodded and put my books away, then waited for the teacher to finish negotiating where to meet the prize winners from the class contest so he could pay for their prizes.

When it was just me and Mr. Gladly in the classroom, he cleared his throat and then told me, \quotation{I'm not stupid, you know.}

\quotation{Okay,} I replied, not sure how to respond.

\quotation{I have something of an idea of what goes on in my classroom. I don't know exactly who, but I know some people are giving you a pretty hard time.}

\quotation{Sure,} I said.

\quotation{I saw the mess left on your usual seat today. I remember a few weeks back when glue was smeared on your desk and chair. There was also the incident that happened at the start of the year. All of your teachers had a meeting about that.}

I couldn't meet his gaze as he brought that last event up. I looked at my feet.

\quotation{And I'm guessing there's more that I don't know about?}

\quotation{Yeah,} I said, still looking down. It was hard to explain how I felt about this conversation. I was gratified, I think, that someone had brought it up, but annoyed that that someone was Mr. Gladly. I felt kind of embarrassed too, like I had walked into a door and someone was trying too hard to make sure I was okay.

\quotation{I asked you after the glue incident. I'm asking you again. Would you be willing to go to the office with me, to talk with the principal and vice principal?}

After a few moments of consideration, I looked up and asked him, \quotation{What would happen?}

\quotation{We'd have a discussion about what's been going on. You would name the person or people you believe responsible, and each of them would be called in to talk to the principal, in turn.}

\quotation{And they'd get expelled?} I asked, though I already knew the answer.

Mr. Gladly shook his head, \quotation{If there was enough proof, they would be suspended for several days, unless they've done something very serious. Further offenses could lead to longer suspensions or expulsion.}

I gave a rueful chuckle, feeling the frustration welling up, “Great. So they {\em might} miss a few days of school, and only if I can prove they were behind it all\ldots{} and whether they get suspended or not, they feel a hundred percent justified in whatever else they do to the rat for revenge.”

\quotation{If you want things to get better, Taylor, you have to start somewhere.}

\quotation{That isn't a starting point. It's shooting myself in the foot,} I said, pulling my bag over my shoulder. When he didn't immediately respond, I left the classroom.

Emma, Madison, Sophia and a half dozen other girls were standing in the hall, waiting for me.


\chapter{Insinuation 2.4}

\quotation{Nobody likes her. Nobody wants her here,} Julia said.

\quotation{Such a loser. She didn't even turn in the major project for art, last Friday,} Sophia responded.

\quotation{If she's not going to try, then why is she even coming to school?}

Despite the way the conversation sounded, they were talking to me. They were just pretending to talk to one another. It was both calculating in how they were managing plausible deniability while at the same time they were acting totally juvenile by pretending I wasn't there. A blend of immaturity mixed with craftiness in a way only high schoolers could manage. I would have laughed at the ridiculousness of it, if it hadn't been at my expense.

The moment I had left the classroom, Emma, Madison and Sophia had crowded me into a corner, with another six girls backing them up. I was unable to squeeze past them without getting pushed or elbowed back, so I couldn't do much more than lean against the window, listening while eight of the girls were rattling off an endless series of taunts and jibes. Before one girl was even finished, another started up. All the while, Emma stayed back and stayed quiet, the slightest of smiles on her face. I couldn't meet the eyes of any of the other girls without them barking a fresh torrent of insults directly to my face, so I just glared at Emma.

\quotation{Ugliest girl in our grade.}

They were barely thinking about what they were saying and a lot of the insults were wildly off the mark or contradictory. One would say I was a slut, for example, then another might say a guy would puke before he touched me. The point wasn't being witty, being smart or being on target. It was more about delivering the feeling behind the words over and over, hammering it in. If I'd had just a moment to butt in, maybe I could have come up with retorts. If I could just kill their momentum, they probably wouldn't get back into the easy rhythm again. That said, I couldn't find the words, and there weren't any openings in the conversation where I wouldn't just be talked over.

While this particular tactic was new to me, I'd been putting up with stuff like this for a year and a half, now. At a certain point, I'd come to the conclusion that it was easier to sit back and take it, when it came to most things. They {\em wanted} me to fight back, because everything was stacked in their favor. If I stood up for myself and they still \quote{won}, then it only served to feed their egos. If I came out ahead in some way, then they got more persistent and mean for the next time. So for much the same reason I hadn't fought Madison for the homework she had taken from me, I just leaned against the wall next to the window and waited for them to get bored with their game or get hungry enough to leave and go have their lunches.

\quotation{What does she use to wash her face? A Brillo pad?}

\quotation{She should! She'd look better!}

\quotation{Never talks to anybody. Maybe she knows she sounds like a retard and keeps her mouth shut.}

\quotation{No, she's not that smart.}

No more than three feet behind Emma, I could see Mr. Gladly leaving his classroom. The tirade didn't stop as I watched him tuck a stack of folders under one arm, find his keys and lock the door.

\quotation{If I were her, I'd kill myself,} one of the girls announced.

Mr. Gladly turned to look me in the eyes.

\quotation{So glad we don't have gym with her. Can you imagine seeing her in the locker room? Gag me with a spoon.}

I don't know what expression I had on my face, but I know I didn't look happy. No less than five minutes ago, Mr. Gladly had been trying to convince me to go with him to the office and tell the principal about the bullying. I watched him as he gave me a sad look, shifted the file folders to his free hand and then walked away.

I was stunned. I just couldn't wrap my head around how he could just ignore this. When he had been trying to help me, had he just been covering his own ass, doing what was required of him in the face of a situation he couldn't ignore? Had he just given up on me? After trying to help, in his own completely ineffective way, after I turned his offer for help down twice, he just decided I just wasn't worth the effort?

\quotation{You should have seen her group fail in class just now. It was painful to watch.}

I clenched my fist, then forced myself to relax it. If we were all guys, this scenario would be totally different. I was in the best shape of my life. I could have swung a few punches from the very start, caused a bloody nose or two, maybe. I know I would have lost the fight in the end, getting shoved to the ground by force of numbers and kicked while I was down, but things would have ended there, instead of dragging on like they were here. I'd hurt physically for days afterwards, but I'd at least have had the satisfaction of knowing some of the others were hurting too, and I wouldn't have to sit through this barrage of insults. If there was enough damage done, the school would have to take notice, and they wouldn't be able to ignore the circumstances of a one-against-nine fight. Violence gets attention.

But things didn't work that way here. Girls played dirty. If I decked Emma, she would run to the office with some fabricated story, her friends backing up her version of events. For most, ratting to the faculty was social suicide, but Emma was more or less top dog. If she went to the principal, people would only take things more seriously. By the time I got back to school, they would have spread the story through the grapevine in a way that made me look like a total psycho. Things would get worse. Emma would be seen as the victim and girls who had previously ignored the bullying would join in on Emma's behalf.

\quotation{And she smells,} one girl said, lamely.

\quotation{Like expired grape and orange juice,} Madison cut in with a little laugh. Again, bringing up the juice? I suspected that one had been her idea.

It seemed like they were running out of steam. I figured it was just a minute or two before they got bored and walked away.

It seemed Emma got the same impression, because she stepped forward. The group parted to give her room.

\quotation{What's the matter, Taylor?} Emma said, \quotation{You look upset.}

Her words didn't seem to fit the situation. I had maintained my composure for however long they had been at it. What I'd been feeling was more a mixture of frustration and boredom than anything else. I opened my mouth to say something. A graceless \quotation{Fuck you} would have sufficed.

\quotation{So upset you're going to cry yourself to sleep for a straight week?} she asked.

My words died in my throat as I processed her words.

Almost a year before we had started high school, I had been at her house, the both of us eating breakfast and playing music way too loud. Emma's older sister had come downstairs with the phone. We'd turned down the music, and my dad had been on the other end, waiting to tell me in a broken voice that my mom had died in a car accident.

Emma's sister had given me a ride to my place, and I bawled the entire way there. I remember Emma crying too, out of sympathy, maybe. It could have been the fact that she thought my mom was the coolest adult in the world. Or perhaps it was because we really were best friends and she had no idea how to help me.

I didn't want to think about the month that had followed, but fragments came to mind without my asking. I could remember overhearing my dad berating my mother's body, because she'd been texting while driving, and she was the only one to blame. At one point, I barely ate for five straight days, because my dad was such a wreck that I wasn't on his radar. I'd eventually turned to Emma for help, asking to eat at her place for a few days. I think Emma's mom figured things out, and gave my dad a talking to, because he started pulling things together. We'd established our routine, so we wouldn't fall apart as a family again.

It was a month after my mom had died that Emma and I had found ourselves sitting on the bridge of a kid's play structure in the park, our rear ends cold from the damp wood, sipping coffee we'd bought from the Donut Hole. We didn't have anything to do, so we had just been walking around and talking about whatever. Our wandering had taken us to the playground, and we were resting our heels.

\quotation{You know, I admire you,} she had said, abruptly.

\quotation{Why?} I had responded, completely mystified about the fact that someone gorgeous and amazing and popular like her could find something to admire in me.

\quotation{You're so resilient. After your mom died, you were totally in pieces, but you're so together after a month. I couldn't do that.}

I could remember my admission, \quotation{I'm not resilient. I can hold it together during the day, but I've cried myself to sleep for a straight week.}

That had been enough to open the floodgates, right there. She gave me her shoulder to cry on, and our coffee was cold before I was done.

Now, as I gaped at Emma, wordless, her smile widened. She remembered what I had said, then. She knew the memories it would evoke. At some point, that recollection had crossed her mind, and she had decided to weaponize it. She'd been waiting to drop it on me.

Fuck me, it worked. I felt the trail of a tear on my cheek. My power roared at the edges of my consciousness, buzzing, pressuring me. I suppressed it.

\quotation{She is! She's crying!} Madison laughed.

Angry at myself, I rubbed my hand over my cheek to brush the tear away. More were already welling up, ready to take its place.

\quotation{It's like you have a superpower, Emma!} one of the girls tittered.

I had taken off my backpack so I could lean against the wall. I reached to pick it up, but before I could, a foot hooked through the strap and dragged it away from me. I looked up and saw the owner of the foot -- dark skinned, willowy Sophia -- smirking at me.

\quotation{Oh em gee! What's she doing?} one of the girls said.

Sophia was leaning against the wall, one foot casually resting on top of my backpack. I didn't think it was worth fighting her over, if it gave her an opportunity to continue her game of keep-away. I left the bag where it was and shoved my way through the gathered girls, bumping an onlooker with my shoulder hard enough to make him stumble. I ran into the stairwell and out the doors on the ground floor.

I fled. I didn't check, but chances were they were watching from the window at the end of the hallway. It didn't really matter. The fact that I had just promised to pay thirty five bucks of my own money for a World Issues textbook to replace the one that had been soaked with grape juice wasn't my top concern. Even if it was pretty much all the money I had left after buying the pieces for my costume. My art midterm was in my bag as well, newly repaired. I knew I wouldn't get any of it back in one piece, if at all.

No, my primary concern was getting out of there. I wasn't going to break the promise I had made to myself. No using powers on them. That was the line I wasn't crossing. Even if I did something utterly innocuous, like give them all lice, I didn't trust myself to stop there. I didn't trust myself to keep from offering blatant hints that I had powers or spoiling my secret identity just to see the looks on their faces when they realized the girl they had been tormenting was a bona-fide superhero. It was something I couldn't help but daydream about, but I knew the long term ramifications would spoil that.

Perhaps most important, I rationalized, was keeping the two worlds separate. What use was escapism, if the world I was escaping to was muddled with the people and things I was trying to avoid?

Before the thought of going back to school had even crossed my mind, I found myself wondering what I was going to do to fill my afternoon.


\chapter{Insinuation 2.5}

If you looked at Brockton Bay as a patchwork of stellar and squalor, upper class and lower class with no middle ground, then downtown was one of the nice areas. The streets and sidewalks were wide, and that meant that even with skyscrapers in every other lot, there was a great deal of blue overhead.

Following my retreat from school grounds, I hadn't been sure what to do. My dad worked an unreliable schedule, so I couldn't spend the rest of the afternoon at home unless I wanted to risk having to explain what I was doing home on a school day. I didn't want to hang around the general area of my school, so that had left me the options of the half-hour walk to downtown or a trip to the Boardwalk. Between my morning runs and the previous night's escapades, I had seen enough of the Boardwalk, so I'd decided to head downtown.

I didn't want to dwell on the subject of school or Emma, so I turned my focus to the recent message from Tattletale. She wanted to meet, presumably to repay the favor she felt she owed me. I considered the possibility that it was a trap, but I couldn't imagine any angle where it would be. She just didn't have any reason to go after me. The worst case scenario was that it wasn't Tattletale, but that wasn't the impression I'd had. What she said in the message seemed to flow with what I had seen of her last night. I would be careful, nonetheless.

It was perplexing. These guys were, in large part, virtual unknowns. From what I knew of Grue and Hellhound, they were both marginally successful B-list villains who had been barely scraping by. Now both were on a team that was pulling high profile heists and confounding even the likes of Armsmaster. The two of them seemed totally different in methodology and style, and if I was remembering right, both Grue and Hellhound had lived in different cities prior to teaming up and setting roots in Brockton Bay. That raised the question: who or what had drawn these four very different individuals together?

It was possible that Tattletale or Regent were the uniting factors, but I couldn't really imagine it, having seen what I did of their group dynamic. Grue had poked fun at Regent rather than treat him like a leader, and while I couldn't put my finger on it, the more I imagined Tattletale uniting that group of unconnected people with powers, the harder I found it to picture. In fact, when I thought about it, hadn't Grue said they had fought for a considerable amount of time over how to deal with Lung? It didn't really sound like they had any leadership worth speaking about.

It wasn't hard to sympathize with Armsmaster. The whole scenario there was just bizarre, and it was made worse by the fact that there were practically no details as far as Tattletale or Regent went. Information, it seemed, was a major factor when dealing with capes.

The streets were busy with people on their lunch break. Businessmen and businesswomen were heading to restaurants and fast food places. My stomach growled as I passed a line of people waiting their turn at a street vendor. I checked my pockets and winced at the realization that I didn't have enough for even a hot dog. My lunch had been in my backpack.

I stopped myself before I could finish that train of thought and put myself into a worse mood by dwelling on what had happened at school. Still, as I thought back to the circle of villains and Tattletale's message, the amusing thought crossed my mind that I could ask them to repay the favor by buying me lunch. It wasn't a serious thought, but the ridiculousness of the mental image -- me eating a burger with a group of supervillains -- put a dumb smirk on my face. I was pretty sure I looked like a moron to anyone on the street who happened to glance at me.

As I thought on it, though, the notion that I might actually consider taking Tattletale up on her offer of a meeting nagged at me. The more I thought on it, the scarier the idea got, and the more it seemed to make sense.

What if I did take them up on the offer? I could meet them, talk with them, see what they had to offer, and all the while, fish for information. If I got anything worth sharing, I could turn around and give it to Armsmaster so he could use it against them. Just going by what Armsmaster had said about these guys and the scarcity of information on them, it would be a pretty major coup for the good guys.

Okay, so they would likely see my ploy as a monumental betrayal if and when I pulled it off. I would be making enemies. That said, I suspected that when it came out that I was a hero and not a villain, they would count it as such regardless. Didn't it make sense to leverage as much information as I could from them before they caught on, as far as their misconception went?

I turned around and headed in the direction of the public library. It was only a few blocks away.

The library was busy, which made sense, given the number of offices and businesses around, the number of people wanting some quiet during their lunch hour, and people doing research or casual browsing they couldn't do at their workplaces. I would have included Brockton Bay's biggest and fanciest high school, the nearby Arcadia High, in that generalization, but I doubted many students were spending their lunch breaks at the library.

The Central Library looked almost more like a museum or art gallery than anything else, with tall ceilings, pillars and massive pieces of artwork hung to frame the hallways between the major sections of the building. I headed up to the second floor, where there were about twenty computers and a line of people waiting their turn to use them. I anticipated a fifteen or twenty minute wait, but as the clock approached one o'clock, people headed back to work and the line rapidly thinned out. A free computer came up within a few minutes of my joining the line. I let the person behind me go on ahead, waiting a bit longer so I could get a station with a little more privacy.

By the time I sat down, I had a pretty good idea of what I wanted to write. I found the message with the search function and clicked on the username \quote{Tt}. A drop down menu appeared, and I chose \quote{send private message}. It gave me the option of making an account, signing in with an already existing account, or sending the message as an anonymous guest. I chose the last option, then typed:

{\bf Subject:} Re:Bug

Bug here. Would like to meet, but want proof you are Tt. I'll reciprocate if needed.

I didn't send it right away, taking a moment to consider. Getting decent proof would prevent any potential problems like the message turning out to be a trap laid by, say, Bakuda. Leaving the burden of proof on Tattletale and leaving it up to her to decide if she wanted verification I was indeed \quote{Bug} meant I didn't have to worry about coming up with exactly how one might prove their identity. I reread it twice over, then sent the message.

The reply came only two or three minutes later. It was fast enough that I couldn't imagine Tattletale taking the time to check and double check every aspect of her message the way I had mine. Was that recklessness on her part, or just the benefit of experience?

I closed the tabs I had opened in the meantime and checked to see what she had written. It was a private message, from her to me, and it set my fight or flight instincts in high gear:

{\bf Subject:} re:Bug

Proof? Last night you didn't say anything until I asked your name. Big guy had a mess of nasty bites and you pepper sprayed him and I told my pal G that when he asked. Good enough?

G R and me will meet you at the same spot we crossed paths last night, k? Don't have to get gussied up if you catch my drift. Rest of us will be in casual wear.

If we meet at 3 will that give you enough time to get there from library with everything you need? let me know

Ta ta

My heart pounded. She knew where I was, and she was letting me know. Why? More to the point, how? Had I unwittingly entered an online exchange with a savvy hacker? I knew my way around computers, my mom had made sure I had one since before I could read and write, but I would be lying if I said I could tell if I was being hacked or do anything about it.

I would have interpreted the casual mention of my location as a veiled threat if it didn't run contrary to everything else in her messages to me. Besides -- Tattletale was talking about meeting me in casual clothes. I took that to mean they wouldn't be in costume. I couldn't understand why, but at the same time, it was hard to imagine her threatening me with one breath just a sentence after she'd offered to meet me in a way that made her totally vulnerable.

Tattletale had unwittingly raised the stakes for my scheme. My primary goal was to gather information on them, and here I was getting a chance to see them with their masks off. It was too good to be true, which made me wonder what kind of safeguards they had in place to protect themselves.

I just had no idea what I would be getting into.

The screensaver came up while I stared at the monitor with thoughts racing through my head. The words \quote{BROCKTON BAY CENTRAL LIBRARY} scrolled across the screen in varying colors.

If I went, best case scenario, I could get enough information to turn them in. I'd get mucho cred from the good guys and respect from an international celebrity. If I'd judged Armsmaster right, I'd get even more brownie points if I gave him the info and let him -- or helped him -- make the bust. On the flip side of the coin, the worst case scenario was that it was a trap, or they'd figure out what I was doing. It would mean a fight, maybe a beating. There was an outside possibility I could get killed, but somehow that didn't concern me as much as it maybe should have. Part of the reason for my lack of concern, I think, was that the possibility existed any time I went out in costume. That, and from my interactions with them last night, I didn't get a \quote{killer} vibe from them.

On the topic of the status quo\ldots{} if I didn't go, what would happen? This particular window of opportunity would likely pass, as far as being able to get the dirt on Tattletale and her gang. That was okay, as I thought on it. It was a high risk, high reward venture anyways. Taking that path would mean turning down the meet, then killing time for the rest of the afternoon, trying to avoid dwelling on the fact that I had missed two straight afternoons of classes and might, maybe, miss more. It was depressing to think about.

\quotation{Excuse me?}

Startled, I looked up. A middle aged woman in a red jacket stood just behind me. As I met her eyes she asked, \quotation{Are you done?} She gestured at the computer, where the screensaver was still scrolling.

Heady with the relief that she hadn't been, irrationally enough, Tattletale, I smiled and told her, \quotation{Give me thirty seconds.}

{\it Subject: Re:Bug

See you at three.}


\chapter{Insinuation 2.6}

I showed up in costume. I didn't care if they thought it was rude or paranoid, I would rather be capable of surviving having a knife pulled on me than play nice.

I had caught a bus from the library to my house and put my costume on under my clothes. Most of the armor panels of my costume were separate pieces, held in place by straps that ran into slits in the fabric of the costume. Not all of them were, though. I'd made some of the armor part of the bodysuit, I'd made narrow, rigid sections of armor running along the center of my chest, back, shins, wrists, hips and the tops of my shoulders. so that when I strapped the larger pieces on, grooves on the underside of the armor would fit over them and help keep them from flopping around. I checked myself in the mirror before I left, and didn't think anyone would notice unless I held a strange posture and they were paying a great deal of attention to what I was wearing. I wore loose fitting clothes over the costume, -- one of my larger pairs of jeans and a sweatshirt, and even with that, I felt painfully conspicuous

I changed much the way I had the previous night, finding an empty alley, quickly pulling on my mask, pulling off my outer clothes, and stuffing the clothes into one of my dad's old backpacks. I'd hidden the backpack before I went patrolling last night, but today, I opted to take it with me. I headed out the other end of the alley.

When I was a short distance away from the site of last night's brawl, I sent a dozen flies out to scout. I focused on what they were sensing.

Bugs, it probably goes without saying, sense things in a very different way than we do. More than that, they sense and process things at a very different speed. The end result was that the signals my power were able to translate and send to me in a way my brain could understand were muted. Visual information came through as ink blot patches of monochrome light and dark, alternating between fuzzy and overly sharp. Sound was almost painful to focus on, breaking down to bass vibrations that made my vision distort and high pitch noises that weren't unlike nails on a chalkboard. Multiply that by a hundred, a thousand, ten thousand, and it was overwhelming. When my power was new to me, I hadn't been able to hold back. The sensory overload had never actually hurt me, even at its worst, but it had made me flat out miserable. These days, I had that part of my power turned off a good ninety nine percent of the time.

My preferred method of sensing things through my bugs was touch. It wasn't that their sense of touch translated much better than the hearing or sight part of things, but had more to do with the fact that I could tell where they were in relation to me. I was acutely aware when they were very still, if they were moving, or if something else was moving them. That was one thing that translated well.

So as I sent the bugs out to scout, the twelve sets of compound eyes first identified the trio as blurry silhouettes atop a larger, more defined shadow, backlit by a flare of white that had to be the sun. I directed the flies closer, towards the \quote{heads} of the figures, and they touched down on skin. None of the three were wearing masks, which I deemed reason to believe Tattletale had been telling the truth. They weren't in costume. There was no guarantee that the three were really Tattletale, Grue and Regent, but I felt confident enough to head around to the fire escape and climb up to the roof.

It was them, no doubt. I recognized them even without their costumes. Two guys and a girl. The girl had dirty blonde hair tied back into a loose braid, a smattering of freckles over the bridge of her nose and the same vulpine grin I recognized from the night prior. She wore a black long sleeved t-shirt with a grafitti-style design on it and a knee length denim skirt. I was surprised by the bottle-glass green of her eyes.

The smaller and younger of the two guys -- about my age -- was undoubtedly Regent. I recognized the mop of black curls. He was a good looking guy, but not in a way that would make me say he was handsome. He was pretty, with a triangular face, light blue eyes and full lips pulled into a bit of a scowl. I pegged him as having French or Italian heritage. I could see where he would have girls all over him, but I couldn't say I was interested, myself. The pretty boys -- Leonardo Decaprio, Marcus Firth, Justin Beiber, Johnny Depp -- had never done it for me. He was wearing a white jacket with a hood, jeans and sneakers, and was perched on the raised lip at the edge of the roof, a bottle of cola in hand.

Grue was startling in appearance, by contrast. Taller than me by at least a foot, Grue had dark chocolate skin, shoulder length cornrows and that masculine lantern jaw you typically associated with guy superheroes. He wore jeans, boots and a plain green t-shirt, which struck me as a bit cold for the spring. I did note that he had considerable muscle definition in his arms. This was a guy who worked out.

\quotation{And she arrives,} Tattletale crowed, \quotation{Pay up.}

Regent's scowl deepened for a second, and he fished in his pocket for a wad of bills, which he forked over to Tattletale.

\quotation{You bet on whether I would show up?} I ventured.

\quotation{We bet on whether you would come in costume,} Tattletale told me. Then, more to Regent than to me, she said, \quotation{and I won.}

\quotation{Again,} Regent muttered.

\quotation{It's your own fault for taking the bet in the first place,} Grue said, \quotation{Even if it wasn't Tattle, it was a sucker bet. Showing up in costume makes too much sense. It's what I would do.} He had a nice voice. It was an adult voice, even if his appearance gave me the sense of a guy in his late teens.

He extended his hand to me, \quotation{Hey, I'm Brian.}

I shook his hand, he wasn't shy about shaking my hand firmly. I said, \quotation{You can call me Bug, I guess. At least, until I come up with something better, or until I decide this isn't an elaborate trick.}

He shrugged, \quotation{Cool.} There wasn't the slightest trace of offense at my suspicion. I almost felt bad.

\quotation{Lisa,} Tattletale introduced herself. She didn't offer me her hand to shake, but I think it would have felt out of place if she had. It wasn't that she seemed unfriendly, but she didn't have the same aura of geniality about her that Grue did.

\quotation{I'm Alec,} Regent informed me, with a quiet voice, then he added, \quotation{And Bitch is Rachel.}

\quotation{Rachel is sitting this one out,} Grue said, \quotation{She didn't agree with the aim of our meeting, here.}

\quotation{Which raises the question,} I cut in, \quotation{What is the aim of this meeting? I'm a little weirded out with you guys revealing your secret identities like this, or at least, pretending to.}

\quotation{Sorry,} Grue\ldots{} Brian apologized, \quotation{That was my idea. I thought we would make a token show of trust.}

Behind the yellow tinted lenses of my mask, my eyes narrowed, flicking from Lisa to Alec to Brian. I couldn't draw any conclusions from their expressions.

\quotation{Why, exactly, do you need my trust?} I asked.

Brian opened his mouth, then closed it. He looked to Lisa, who bent down and picked up a plastic lunchbox. She held it out to me.

\quotation{I said we owed you. All yours, no strings attached.}

Without taking the box, I tilted my head to get a better look at the front, \quotation{Alexandria. She was my favorite member of the Protectorate when I was a kid. Is the lunchbox collectable?}

\quotation{Open it,} Lisa prompted me, with a roll of her eyes.

I took it. From the weight and the motion of the contents inside it, I immediately had a pretty good idea of what it was. I undid the clasps and opened the box.

\quotation{Money,} I breathed, caught off guard by suddenly having so much in my hands. Eight stacks of bills, tied with paper bands. Each of the paper bands had a number written on it in permanent marker. Two fifty each\ldots{}

Lisa answered before I had the number totaled up in my head, \quotation{Two grand.}

I closed the box and did the clasps. With no idea what to say, I stayed silent.

\quotation{You have two choices,} Lisa explained, \quotation{You can take that as a gift. A thank you for, intentionally or not, saving our ass from Lung last night. And maybe a bit of incentive to count us among your friends when you're out in costume and doing dastardly deeds.}

Her grin widened, as if she'd said something she found amusing. Maybe it was the irony of a villain talking about \quote{dastardly deeds}, or how corny the phrase was. She elaborated, \quotation{Between territory disputes, differences in ideology, general power struggles and egos, there's a rare few people in the local villain community who won't attack us on sight.}

\quotation{And the second option?} I asked.

\quotation{You can take this as your first installment in the monthly allowance you're entitled to as a member of the Undersiders,} Brian spoke up, \quotation{As one of us.}

I shifted my gaze between the three of them, looking for the joke. Lisa still had a bit of a smile, but I was getting the impression that was her default expression. Alec looked a little bored, if anything. Brian looked dead serious. Damn.

\quotation{Two thousand a month,} I said.

\quotation{No,} Brian cut in, \quotation{That's just what the boss pays us, to stick together and to stay active. We make, uh, considerably more than that.}

Lisa smirked, and Alec chuckled as he swished the contents of his coke bottle. I made mental note at the mention of this \quote{boss}.

Not wanting to get sidetracked, I quickly thought through the earlier part of our conversation in the context of the job offer.

I asked, \quotation{So Bitch didn't come because she was against the, er, recruitment?}

\quotation{Yeah,} Alec said, \quotation{We voted on it, and she said no.}

\quotation{On the plus side, the rest of us voted yes,} Brian hurried to add, giving Alec a dirty look, \quotation{She'll come around. She always votes against adding new members to the group, because she doesn't want to divide the money five ways.}

\quotation{So you've done this recruiting thing before,} I concluded.

\quotation{Uh, yeah,} Brian looked a touch embarrassed, he rubbed the back of his neck, \quotation{It didn't go well. We tried with Spitfire, and she got scared off before we even got to the job offer. Our fault, for bringing Rachel along that time.}

\quotation{And then she got recruited by someone else,} Alec added.

\quotation{Yeah,} Brian shrugged, \quotation{She got snagged by Faultline before we got a second chance. We've made an offer to Circus, too, and she told us in no uncertain terms that she worked alone.}

\quotation{Taught me a few new curse words while she did it, too,} Alec said.

\quotation{She was pretty vocal about how she flies solo,} Brian admitted.

\quotation{So you're going the extra mile, with no costumes as a show of trust and a cash bonus up front, to get me to join,} I said, as the pieces came together.

\quotation{That's the gist of it,} Brian agreed, \quotation{Long and short of it is, especially with Lung taken out of action and the ABB diminished by his being gone, there's bound to be some pushing and shoving over territory and status among the various gangs and teams. Us, Faultline's Crew, the remaining ABB, Empire Eighty-Eight, the solo villains, and any out of town teams or gangs that figure that they can worm in and grab a piece of the Bay. If it comes down to it, we want firepower. We haven't screwed up a job yet, but the way us three figure it, it's only a matter of time before we end up stuck in a fight we can't win, with Bitch as the only one of us who can really dish out the hurt.}

\quotation{I just don't get why you want me,} I said, \quotation{I control bugs. That's not going to stop Alexandria, Glory Girl or Aegis.}

\quotation{You fucked up Lung,} Lisa shrugged as she spoke, \quotation{Good enough for me.}

\quotation{Um, not really,} I replied, \quotation{In case you missed it, you're the ones who stopped him from executing me last night. That just goes to prove the point I was making.}

\quotation{Honey,} Lisa said, \quotation{Entire teams of capes have gone up against Lung and got their asses handed to them. That you managed as well as you did is fantastic. The fact that the asshole is lying in a hospital bed because of you is the icing on the cake.}

My reply stopped before it even left my mouth. I only managed a dumb, \quotation{Hunh?}

\quotation{Yeah,} Lisa raised an eyebrow, \quotation{You do know which bugs you had biting him, right? Black Widow, Brown Recluse, Browntail Moth, Mildei, Fire Ants-}

\quotation{Yeah,} I cut her off, \quotation{I don't know the official names, but I know exactly what bit him, what stung him and what the venoms do.}

\quotation{So why are you surprised? A couple of those bugs would be fucking dangerous if they bit just once, but you had them bite several times. Bad enough, but when Lung came into custody they had him checked over by the docs, and the idiot doctor in charge said something like, \quote{Oh, well, these do look like bug bites and stings, but the really venomous ones don't bite multiple times. Let's arrange to check on him in a few hours}.}

I could tell where the story was going. I put my hands over my mouth, whispering, \quotation{Oh my god.}

Tattletale grinned, \quotation{I can't believe you didn't know.}

\quotation{But he regenerates!} I protested, dropping my hands, \quotation{Toxins aren't supposed to be even one percent as effective against people who heal like he does.}

\quotation{They're effective enough, I guess, or his healing stopped working somewhere along the line} Lisa told me, “By the time they got to him, the big guy was just beginning to suffer from large scale tissue necrosis. His heart even stopped a few times. You {\em do} remember where you had the bugs bite him?”

I closed my eyes. I could see my reputation going down the tubes. One of the spiders I had been using was the brown recluse. Arguably the most dangerous spider in the United States, more than even the black widow. A single bite from a brown recluse could make a good chunk of the flesh around the bite blacken and rot away. I'd had my bugs biting Lung in the more sensitive parts of his anatomy.

\quotation{Let's just say that even with the ability to heal several times faster than your average person, Lung is going to be sitting down to use the toilet.}

\quotation{Okay, that's enough,} Brian stopped Lisa before she could go on, \quotation{Lung is going to recover, right?}

With the look Brian was giving Lisa, I thought she might lie, regardless of the truth. She shrugged and told me, \quotation{He's already recuperating. Slowly, but he's on the mend, and he should be in good working order in six months to a year.}

\quotation{You'd better hope he doesn't escape,} Alec said, his voice still quiet but bemused, \quotation{Because if someone made my man bits fall off, I'd be out for blood.}

Brian pinched the bridge of his nose, \quotation{Thank you for that, Alec. Way you two are going, our potential recruit is going to run off to have a panic attack before the idea of becoming an Undersider even crosses her mind.}

\quotation{How do you know this?} I asked, within a heartbeat of the thought crossing my mind. When Brian turned my way with an expression like he thought he had said something to offend me, I clarified, \quotation{Tattletale, or Lisa, or whatever I'm supposed to call you. How do you know this stuff about Lung\ldots{} or about the fact that I was at the Library, or that the cape was on his way, last night?}

\quotation{Library?} Brian interjected, giving Lisa another dark look.

Lisa ignored Brian's question and winked at me, \quotation{Girl's gotta have her secrets.}

\quotation{Lisa's half the reason we haven't failed a job yet,} Alec said.

\quotation{And our boss is a large part of the rest,} Lisa finished for him.

\quotation{So you say,} Brian grumbled, \quotation{But let's not go there.}

Lisa smiled at me, \quotation{If you want the full scoop, I'm afraid the details on what we do only come with team membership. What I can tell you is that we're a good group. Our track record is top notch, and we're in it for fun and profit. No grand agenda. No real responsibility.}

I pursed my lips, behind my mask. While I had picked up some info, I felt like I had a lot more questions. Who was this boss they mentioned? Was he or she setting up other teams of highly successful villains, in Brockton Bay or elsewhere? What made these guys as effective as they were, and was it something I could steal or copy for myself?

It wasn't like I was signing the deal in blood or anything. I stood to gain so much.

\quotation{Alright then, count me in,} I told them.


\chapter{Insinuation 2.7}

As I agreed to join the Undersiders, there was some whooping and cheering. I felt a touch guilty, for acting under false pretenses. I also felt pleased with myself, in an irrational way.

\quotation{Where do we go from here?} Lisa asked Brian.

\quotation{Not sure,} Brian said, \quotation{It's not like we've done this before. I suppose we should let Rachel know, but she said she might work today.}

\quotation{If the new girl is okay with it, let's stop by our place,} Lisa suggested, \quotation{See if Rache is there, celebrate the new recruit and get her filled in.}

\quotation{Sure,} I said.

\quotation{It's just a few blocks away,} Brian said, \quotation{But we would stand out if you came with in costume.}

I stared at him for a moment, not wanting to comprehend his statement. If I took too long to respond, I realized, I would ruin this plan before it went anywhere. Whatever the case, I could have kicked myself. Of course this was the natural progression of events. Joining their team would mean I would be expected to share my identity, since they already had. Until I did, they wouldn't be able to trust me with their secrets.

I could have blamed the lapse in judgement and foresight on my lack of sleep or the distraction of the events earlier in the day, but that didn't change matters. I had maneuvered myself into a corner.

\quotation{Alright,} I said, sounding calmer than I felt. I hoped. \quotation{This costume is kinda uncomfortable under clothes. Can I get some privacy?}

\quotation{You want an alley, or\ldots{}} Lisa asked, trailing off.

\quotation{I'll change here, just take a minute,} I said, impulsively, as I glanced around. The buildings on the street were mostly one and two stories tall, with the only buildings taller than the one we were on being the one half a block away, and the one right next to us. There weren't any windows on the building next to us with a great angle for seeing me change, and I doubted anyone on the distant building could see me as more than a figure two inches tall. If someone could see me change out of costume and make out enough details to identify me, I'd be surprised.

As the three of them headed to the fire escape, I pulled out the clothes I'd stuffed into the backpack. Armor panels aside, my costume was essentially one piece, with the exceptions being the belt and the mask. I kept the mask on as I undid the belt and peeled off the main costume. I wasn't indecent -- I was wearing a black tank top and black biking shorts underneath, in part for extra warmth. Silk wasn't the best insulator on its own. I stepped into my jeans and pulled on the sweatshirt, then rubbed my arms and shoulders to brush off the mild chill. I put my costume and the plastic lunchbox in my backpack.

I felt a stab of regret at not having chosen better clothes to wear than a loose fitting sweatshirt and jeans that were too big for me. That regret quickly turned to a pang of anxiety. What would they think when they saw the real me? Brian and Alec were good looking guys, in very different ways. Lisa was, on the sliding scale between plain and pretty, more pretty than not. My own scale of attractiveness, by contrast, put me somewhere on a scale that ranged from \quote{nerd} to \quote{plain}. My opinion of where I fit on that scale changed depending on the mood I was in when I was looking in the mirror. They were cool, confident, assured people. I was\ldots{} me.

I stopped myself before I could get worked up. I wasn't regular old Taylor, here. In the here and now, I was the girl who had put Lung in the hospital, accidental as it was. I was the girl who was going undercover to try and get the details on a particularly persistent gang of supervillains. I was, until I came up with a better name to go by, Bug, the girl the Undersiders wanted on their team.

If I said I made my way down the fire escape filled to the brim with confidence, I'd be lying. That said, I had managed to hype myself up enough to get myself down the ladder, mask still on, costume in my bag. I stood before them, glanced around to make sure nobody else was around, and then pulled off my mask. I had a few terrifying heartbeats where I was half-blind, their facial features just smudges, before I put on the glasses I'd had in my bag.

\quotation{Hi,} I said, lamely, using my fingers to comb my hair back into order, \quotation{I guess it wouldn't work if you kept calling me Bug or new girl. I'm Taylor.}

Using my real name was a big gamble on my part. I was afraid it would be another thing I would be kicking myself for five minutes from now, much like the realization that I'd have to go uncostumed. I rationalized it by telling myself that I was already in this wholesale. Being truthful about that one thing might well save my hide if any of them decided to do some digging on me, or if I ran into someone I knew while in their company. I figured, hoped, that by the time this whole thing was over, I could maybe pull some strings with someone like Armsmaster and avoid having them leak my real name. Not impossible to imagine, given the level of security around some of the prisons they had for criminal parahumans. In any event, I would cross that bridge when I got to it.

Alec offered the slightest roll of his eyes as I introduced myself, while Brian just grinned. Lisa, though, put one of her arms around my shoulders and gave me a one-armed squeeze of a hug. She was a little older than I was, so she was just tall enough to be at the perfect height to do it. What caught me off guard was how nice the gesture felt. Like I had been needing a hug from someone who wasn't my dad for a long time.

We walked deeper into the Docks as a group. While I had lived on the periphery of the area my entire life, and while most people would say the neighborhood I lived in was part of the \quote{Docks}, I had never really been in the areas that gave this part of the city such a bad reputation. At least, I hadn't if I discounted last night, and it had been dark then.

It wasn't an area that had been kept up, and kind of gave off an impression of a ghost town, or what a city might look like if war or disaster forced people to abandon it for a few years. Grass and weeds grew between slats in the sidewalk, the road had potholes you could hide a cat in, and the buildings were all faded, consisting of peeling paint, cracked mortar and rusty metal. The desaturated colors of the buildings were contrasted by splashes of vividly colored graffiti. As we passed what had once been a main road for the trucks traveling between the warehouses and the docks, I saw a row of power lines without wires stretching between them. At one point weeds had crawled most of the way up the poles, only to wither and die at some point. Now each of the poles had a mess of dead brown plants hanging off of them.

There were people, too, though not too many were out and about. There were those you expected, like a homeless bag lady with a grocery cart and a shirtless old man with a beard nearly to his navel, collecting bottles and cans from a dumpster. There were others that surprised me. I saw a woman that looked surprisingly normal, in clothes that weren't shabby enough to draw attention, herding four near-identical infant children into a factory building with a faded sign. I wondered if they were living there or if the mom was working there and just couldn't do anything with her kids but bring them with her. We passed a twenty-something artist and his girlfriend, sitting on the sidewalk with paintings propped up around them. The girl waved at Lisa as we walked by, and Lisa waved back.

Our destination was a red brick factory with a massive sliding metal door locked shut by a coil of chain. Both the chain and door had rusted so much that I expected that neither offered any use. The size of the door and the broadness of the driveway made me think that large trucks or small boats would have been backed up through the entryway back in the factory's heyday. The building itself was large, stretching nearly half the block, two or three stories tall. The background of the sign at the top of the building had faded from red to a pale orange-pink, but I could make out the bold white letters that read \quote{Redmond Welding}.

Brian let us in through a small door on the side of the building, rather than the big rusted one. The interior was dark, lit only by rows of dusty windows near the ceiling. I could make out what had been massive machines and treadmills prior to being stripped to their bare bones. Sheets covered most of the empty and rusted husks. Seeing the cobwebs, I reached out with my power and felt bugs throughout. Nobody had been active in here for a long time.

\quotation{Come on,} Brian urged me. I looked back and saw that he was halfway up a spiral staircase in the corner. I headed up after him.

After seeing the desolation of the first floor, seeing the second floor was a shock. It was a loft, and the contrast was startling. The exterior walls were red brick, and there was no ceiling beyond a roof and a skeleton of metal girders overhead to support it. In terms of general area, the loft seemed to have three sections, though it was hard to define because it was such an open layout.

The staircase opened up into what I would have termed the living room, though the one room alone had nearly as much floor space as the ground floor of my house did. The space was divided by two couches, which were set at right angles from one another, both facing a coffee table and one of the largest television sets I had ever seen. Below the television set were a half dozen video game consoles, a DVD player and one or two machines I didn't recognize. I supposed they might have a TiVo, though I'd never seen one. Speakers larger than the TVs my dad and I had at home sat on either side of the whole setup. Behind the couches were tables, some open space with rugs and shelves set against the walls. The shelves were only half filled with books and magazines, while the rest of the shelf space was filled with odds and ends ranging from a discarded shoe to candles.

The second section was a collection of rooms. It was hard to label them as such, though, because they were more like cubicles, three against each wall with a hallway between them. They were a fair size, and there were six doors, but the walls of each room were only eight or so feet tall, not reaching all the way up to the roof. Three of the doors had artwork spray painted on them. The first door had a crown done in a dramatic graffiti style. The second door had the white silhouette of a man and a woman against a blue background, mimicking the \quote{mens} and \quote{womens} washroom signs that were so common. The third had a girl's face with puckered lips. I wondered what the story was, there.

\quotation{Nice art,} I said, pointing at the door with the crown on it, feeling kind of dumb for making it the first thing I'd said as I entered the room.

\quotation{Thanks,} Alec replied. I guess that meant it was his work.

I took another second to look around. The far end of the loft, the last of the three sections, had a large table and some cabinets. Though I couldn't take a better look without crossing the whole loft, I gathered that their kitchen was in the far end of the loft.

Throughout, there was mess. I felt almost rude for paying attention to it, but there were pizza boxes piled on one of the tables, two dirty plates on the coffee table in front of the couch, and some clothes draped over the back of one of the couches. I saw pop cans -- or maybe beer cans -- stacked in a pyramid on the table in the far room. It wasn't so messy that I thought it was offensive, though. It was mess that made a statement\ldots{} like, \quote{This is our space.} No adult supervision here.

\quotation{I'm jealous,} I admitted, meaning it.

\quotation{Dork,} Alec said, \quotation{What are you jealous for?}

\quotation{I meant it's cool,} I protested, a touch defensively.

Lisa spoke before Alec could reply, \quotation{I think what Alec means is that this is your place now too. This is the team's space, and you're a member of the team, now.}

\quotation{Oh,} I said, feeling dumb. Lisa and Alec headed to the living room, while Brian walked off to the far end of the loft. When Lisa gestured for me to follow her, I did. Alec lay down, taking up an entire couch, so I sat on the opposite end of the couch from Lisa.

\quotation{The rooms,} Lisa said, \quotation{Far side, in order of closest to farthest, are Alec, bathroom, mine.} That meant Alec's room was the one with the crown, and Lisa's door had the face with the puckered lips. She went on, \quotation{On the side closer to us, Rachel's room, Rachel's dogs' room, and the storage closet.}

Lisa paused, then glanced at Alec and asked, \quotation{You think she-}

\quotation{Duh,} Alec cut her off.

\quotation{What?} I asked, feeling lost.

\quotation{We'll clean out the storage closet,} Lisa decided, \quotation{So you have a room.}

I was taken aback. \quotation{You don't have to do that for me,} I told her, \quotation{I've got a place.}

Lisa made a face, almost pained. She asked me, \quotation{Can we just do it anyways, and not make a fuss? It'd be a lot better if you had your own space here.}

I must have looked confused, because Alec explained, \quotation{Brian has an apartment, and was pretty firm about not needing or wanting a room here\ldots{} but he and Lisa have been arguing regularly because of it. He has nowhere to sleep but the couch if he gets hurt and can't go to his place, and there's no place to put his stuff, so it gets left all over. Take the room. You'll be doing us a favor.}

\quotation{Okay,} I said. I added, \quotation{Thank you,} as much for the explanation as for the room itself.

\quotation{Last time he went up against Shadow Stalker, he came back here and bled all over a white couch,} Lisa groused, \quotation{nine hundred dollar couch and we had to replace it.}

\quotation{Fucking Shadow Stalker,} Alec commiserated.

Brian came back from the other end of the loft, raising his voice to be heard as he approached, \quotation{Rache's not here, and neither are her dogs. She must be walking them or working. Dammit. I get stressed when she's out.} He approached the couches and saw Alec sprawled on the one.

\quotation{Move your legs,} Brian told him.

\quotation{I'm tired. Sit on the other couch,} Alec mumbled, one arm over his face.

Brian glanced at Lisa and I, and Lisa scooted over to make room. Brian glared down at Alec and then sat between us girls. I shifted my weight and tucked one leg under me to give him room.

\quotation{So,} Brian explained, \quotation{Here's the deal. Two grand a month, just to be a member of the team. That means you help decide what jobs we do, you go on the jobs, you stay active, you're available if we need to call.}

\quotation{I don't have a phone,} I admitted.

\quotation{We'll get you one,} he said, like it wasn't even a concern. It probably wasn't. \quotation{We generally haul in anywhere from ten grand to thirty-five grand for a job. That gets divided four ways\ldots{} five ways now that you're on the team.}

I nodded, then exhaled slowly, \quotation{It's not small change.}

Brian nodded, a small smile playing on his lips, \quotation{Nope. Now, how on the ball are you, as far as knowing what we're up against?}

I blinked a few times, then hedged, \quotation{For other local capes? I've done research online, read the cape magazines religiously for a few years, more since getting my powers\ldots{} but I dunno. If the past twenty four hours have taught me anything, it's that there's a lot I don't know, and will only find out the hard way.}

Brian smiled. I mean, really smiled. It made me think of a boy rather than a nearly-grown man. He replied, \quotation{Most don't get that, you know? I'll try to share what I know, so you aren't caught off guard, but don't be afraid to ask if there's anything you're not sure about, alright?}

I nodded, and his smile widened. He said, through a good natured chuckle, \quotation{Can't tell you how much of a relief it is that you take this stuff seriously, since some people -} he stopped to lean over and kick the side of the couch Alec was lying on, \quotation{-need me to twist their arms to get them listening, and some people,} he jerked his thumb over his right shoulder, \quotation{think they know everything.}

\quotation{I do know everything,} Lisa said, \quotation{It's my power.}

\quotation{What?} I said, interrupting Brian. My heartbeat quickened, though I hadn't exactly been relaxed to begin with, \quotation{You're omniscient?}

Lisa laughed, \quotation{No, no. I do know things though. My power tells me stuff.}

Swallowing hard, hoping I wasn't drawing attention by doing so, I asked, \quotation{Like?} Like why I was joining their team?

Lisa sat forward and put her elbows on her knees, \quotation{Like how I knew you were at the library when I sent me the messages. If I felt like it, and if I had the know how, I'm sure I could have figured it out by breaking into the website database and digging through the logs to find the address you connected from, but my power just let me skip that step like that.} She snapped her fingers.

\quotation{And why exactly did you mention you knew where she was?} Brian queried, his voice a touch too calm.

\quotation{I wanted to see how she'd react. Messing with her a little,} Lisa grinned.

\quotation{God dammit-} Brian started, but Lisa waved him off.

\quotation{I'm filling the newbie in,} she waved him off, \quotation{Yell at me later.}

Not giving him a chance to reply, she turned to me and explained, \quotation{My power fills in the gaps in my knowledge. I generally need some info to start from, but I can use details my power feeds me to figure out more stuff, and it all sort of compounds itself, giving me a steady flow of info.}

I swallowed, \quotation{And you knew that a cape was on the way last night?}

\quotation{Yeah,} she said, \quotation{Call it a well educated guess.}

\quotation{And you knew the stuff about what happened in the PHQ the same way?}

Lisa's smile widened, \quotation{I'll admit I cheated there. Figuring out passwords is pretty easy with my power. I dig through the PHQ's digital paperwork and enjoy a little reality TV by way of their surveillance cameras when I'm bored. It's useful because I'm not only getting the dirt from what I see, hear and read, but my power fills in the details on stuff like changes in their routine and the team politics.}

I stared at her, a good part of me horrified that I'd gotten into an undercover situation opposite a girl with superpowered intuition.

Taking my silence for awe, she grinned her vulpine smile, “It's not {\em that} amazing. I'm really best with concrete stuff. Where things are, timing, encryption, yadda yadda. I can read something out of changes in body language or routine, but it's less reliable and kind of a headache. Enough information overload without, you know?”

I did know, her explanation echoed my own thoughts regarding my ability to see and hear things through my bugs. Still, her words didn't make me feel that much better.

\quotation{And,} Brian said, still glowering at Lisa, \quotation{Even if she knows a lot, that doesn't mean Lisa can't be a dumbass sometimes.}

Lisa punched him in the arm.

\quotation{So what are your powers then?} I asked Brian and Alec, hoping for a change in topic.

They didn't get a chance to tell me. I heard barking from downstairs. A matter of heartbeats later I was standing, three paces from the couch. Three snarling dogs had me backed against the wall, drool flying from their mouths as their teeth gnashed and snapped for my hands and face.


\chapter{Insinuation 2.8}

\quotation{Call off your dogs!} Brian shouted.

The largest of the dogs, an ugly Rottweiler or a mutt with strong Rottweiler blood, seized my wrist in its jaws. My knees almost buckled in response to the pain, which only worsened when it abruptly snapped its head to one side and wrenched my arm. I fell, and in a heartbeat, the other two dogs -- a German Shepherd and a hairless terrier with a missing ear and eye -- were on me.

The German Shepherd set to barking and snapping at my face, occasionally catching the hair that hung in front of my face to pull at it. The other started raking at me with its claws and nipping with its teeth, trying to find somewhere on my legs, body or backside that it could set its teeth into.

While those two were at it, the Rottweiler still had my wrist in its teeth, and it began pulling, as though it wanted to drag me somewhere. I grit my teeth at the pain and tried to think something I could do that would amount to more than curling up into a fetal position to protect my arms, legs and face.

\quotation{Call off the fucking dogs!} I heard Brian bellow, again.

A tooth or claw scratched my ear. I think that's what spooked me, because my composure broke and I cried out.

Just a second or two later, a much longer span of time than it sounds like when a pack of dogs is tearing at you, there was a whistle. Hearing the noise, the dogs abruptly backed off. The one eyed terrier offered one hostile bark followed by a long growl even as it walked away, as if it still had enough mean left in it that it had to let it out somehow.

Lisa and Alec helped me to my feet. I was shaking like a leaf. One of my hands gripped the forearm of my other arm, as much to stop the worst of the trembling as to cradle the injury. I had tears in the corners of my eyes and I was clenching my teeth so hard my jaw ached.

On the opposite side of the room, Brian was rubbing the back of one of his hands. The three dogs were sitting in a neat line ten feet away from a girl who was lying on the ground. The girl had blood running from both of her nostrils. I recognized her from the picture I had seen on her wiki page. Rachel Lindt. Hellhound. Bitch.

“I {\em fucking} hate it,” Brian growled at the girl, putting emphasis on the swear, \quotation{When you make me do that.}

Bitch propped herself up a bit, half against the wall opposite me, so she had a better view of the room. A better view of me. Seeing her in person just confirmed my impressions of her from her picture online. She wasn't attractive. An unkind person might call her butch, and I wasn't feeling particularly kindly towards her. Most of her features looked like they would have been better fit on a guy rather than a girl. She had a square face, thick eyebrows, and a nose that had been broken more than once -- maybe broken again just a moment ago, given the blood trickling from her nostrils. Even as far as her physical build went, she was solidly built without being fat. The trunk of her body alone was bigger around than mine was with my arms down at my sides, just by virtue of having a thicker, broader torso and having more meat on her bones. She was wearing boots, black jeans with tears all over them, and a green army jacket over a gray hooded sweatshirt. Her auburn hair was cut shortish.

I took a deep breath. Then, speaking slowly so I wouldn't stumble over my words or let a tremor into my voice, I asked \quotation{Why the fuck did you do that?}

She didn't reply. Instead, she licked her upper lip clean of blood and smiled. It was a mean, smug sneer of a smile. Even though she was the one lying on the ground with a bloody nose, she somehow had it in her head that she'd beat me. Or something.

\quotation{God fucking dammit!} Brian was shouting. He went on to say something else, but I didn't really hear it over the buzzing of my power in my ears. I realized I was clenching my fist, and habitually forced myself to relax it.

Then, like I had done so many times over the past few days and weeks, I searched for a reason to justify why I was backing down. It was almost reflexive. When the bullies got on my case, I always had to take a moment to collect myself and tell myself why I couldn't or shouldn't retaliate.

For a few moments, I felt adrift. Around the same time that I realized I couldn't find a reason to back off, I realized I had already wrenched free of Lisa and Alec's support and crossed half of the room at a run. I reached for my bugs and realized I'd been using my power without thinking about it. They were already gathering at the stairs and by the windows. All it took was a thought, and they started flowing into the room in greater numbers. Cockroaches, earwigs, spiders and flies. Not as many as I might have liked, I hadn't been using my power for long enough to gather those from further around the neighborhood, but it was enough to count.

Bitch saw me approaching and raised her fingers to her mouth, but I didn't give her a chance to signal her animals. I kicked for her face like I might kick a soccer ball, and she aborted the whistle to cover her head with her arms. My foot bounced off of one of her arms and her entire body recoiled as she flinched.

Because I hadn't slowed down before reaching her, I had to use my hands to stop myself from running into the wall. A line of red hot pain ran down my arm at the impact, starting at the point where the Rottweiler had bitten my wrist. Reminded of the dogs, I glanced to my right, and saw the largest of them standing, ready to come to his master's aid. I brought a large share of my bugs in between myself and the beasts. The last I saw of them before the swarm blocked most of my view, the dogs were rapidly backing away from the swarm, startled.

Finding myself standing over Bitch, braced against the wall, I pressed the attack. Her arms were covering her face and chest, but I saw her exposed ear as a target and brought my foot down on it. Her head bounced against the floor, and blood bloomed from the top of her ear. The sight of the blood almost stopped me, but I knew that backing down now would give her a chance to set them on me again with a whistle. My toe found her exposed stomach, and as she drew her knees upward to protect her belly, I aimed a sharp kick between her legs. I managed to get kicks to connect firmly with ribs three times before she brought an elbow down to protect it.

I didn't get a chance to do any more damage, because the dogs had gotten over their fear of the bugs and were closing in, circling around me and Bitch as the swarm extended. I abandoned my assault on Bitch to step away and face them. I knew I could set my bugs on them, but something told me the dogs weren't about to yelp and run away while their master was being hurt. I might have the swarm attack them, but if the pain of the bites and stings didn't stop them, they'd attack me and I'd be in the same situation I'd been in a minute ago. I doubted Bitch would call them off a second time.

A shadow fell over my vision, like a jet black curtain sweeping in front of me, blocking my view of half the room and the dogs. It dissolved into wisps of black smoke a second later, and I was startled to see Brian right in front of me, between me and the dogs.

\quotation{Enough,} he intoned. The little one-eared cyclops of a terrier snarled at him in response.

There was a sound I didn't recognize. It was only when Bitch tried again, more successfully, that I realized the first sound had been a weak attempt at a whistle. The dogs looked to their master and then retreated, still edging away from the swarm. I backed away a little as well, being careful to keep Brian between myself and the mongrels.

Bitch coughed, then raised her head to look me in the eye. She rubbed her ear with one hand, and her palm was red with blood as she pulled it away. As the German Shepherd approached her, she rested the same hand on its head. The other two dogs moved closer to her, as if they could protect her, but their attention was fixed entirely on me and Brian.

When a good few seconds had passed and Bitch had made no further overtures of aggression towards me, I sent an instruction to the swarm to make their exit. I could see Brian visibly relax as they faded into the cracks.

\quotation{No more fighting,} he said, his voice calmer, \quotation{I'm directing that at you, Rachel. You deserved whatever Taylor gave you.}

She glared at him, coughed once, and then glanced at the other two before turning her angry gaze to the floor.

\quotation{Taylor, come sit down. I promise we'll-}

\quotation{No,} I interrupted him, \quotation{Fuck this. Fuck you guys.}

\quotation{Taylor-}

“You said she wasn't cool with me joining. You {\em never} said she was pissed off enough to try and kill me.”

Bitch and Brian started speaking at the same time, but Brian stopped when she started coughing. As her coughing fit subsided, Bitch looked up at me and snarled, \quotation{If I ordered them to kill you, Brutus would have torn out your throat before you could scream. I gave them the hurt command.}

I laughed a little, just a little more high pitched than I would've liked, \quotation{That's great. She has her dogs trained to hurt people. Seriously? Fuck you guys. Count this as another failed recruitment.}

I headed for the stairs, but I didn't get two steps before that curtain of black appeared again, blocking my way. Brian's powers in the wiki had been listed as darkness generation. I knew where the stairs and the railing for the stairs was, so I put my hand in front of me to make sure I wasn't walking into an opaque forcefield, and on finding it to be more like smoke, I kept moving. As I entered it, the blackness slithered over my skin, oily with a weird consistency to it. Combined with an absolute lack of light that left me unable to tell whether my eyes were open or shut, it was ominous.

As my hands made contact with the railing, a pair of hands settled on my shoulders. I wheeled around and knocked them away, my voice raised as I half-shouted, \quotation{Back off!}

Except the words barely reached {\em me.} The sound echoed as if from a distant place, and had a hollowness to it that made me think of someone shouting from the bottom of a deep well. The darkness didn't just block off the light. It swallowed up noises as well. I'd let go of the railing when I turned to face the other person in the darkness, and I had a moment's panic when I realized I couldn't tell where the stairs were anymore. The texture of the darkness was inconsistent, making it hard to identify the full scope of my movements. I was reminded of those times I had been underwater and lost track of which direction the surface was. I could tell which way was up, sure, but that was about it.

Sensory deprivation. When those two words came to my mind, I felt myself relax some. Brian's power mucked with your senses\ldots{} Sight, hearing, touch. I wasn't limited to those three. Reached out with my power, I identified where all of the bugs in the loft and the factory below were. Using them to ground myself like a sailor might use the constellations, I figured out where the stairs should be and found the railing. The hands hadn't grabbed for me again, so I hurried down, down the stairs and out of the oppressive darkness.

I was only a few paces from the door when Brian called for me, \quotation{Taylor!}

When I turned to face him, I saw he was alone.

\quotation{You're going to use your power on me again?} I asked, wary, angry.

\quotation{No. Not in the open, not uncostumed, and not on you. It was stupid of me to do it in the first place. I wasn't thinking, I just wanted to stop you from bolting. I can barely tell it's there, so I forget how it can affect other people.}

I started to turn away, ready to walk, but Brian took a quick step in my direction, and I stopped.

Brian tried again, \quotation{Look, I'm sorry. About using my power on you, about Bitch.}

I cut him off before he could get any further, \quotation{You don't have to worry. I won't tell anyone what you guys showed me tonight, I won't be attacking you guys if I run into you in costume. I'm pissed, but I'm not that pissed.} I wasn't sure how much of that was a lie, but it seemed like the thing to say.

When he didn't say anything in response, I added, \quotation{You guys offered me a choice. I could take the money and go, or I could join. Let me change my mind. After what your teammate just did, you owe me that much.}

\quotation{If it were up to me, I'd kick Bitch out and keep you,} Brian spoke.

His words were like a bucket of water in my face, waking me up. I'd been pissed, furious, and why? Because I'd felt betrayed and disappointed. The irony of that, given my whole reason for being there in the first place, didn't escape me. I wouldn't have been as disappointed and betrayed as I was if I didn't enjoy their companionship on some level. Here Brian was, expressing similar sentiments from the other side of things.

I let out a long sigh. I guessed, \quotation{But you won't?}

\quotation{It's complicated. As much as I want you on the team, we count on the boss for our allowances, information, equipment and for fencing anything we steal. We count on her to deploy our heavy hitters. We'd lose all that if we kicked her out.}

\quotation{I became a-} I almost said superhero, \quotation{cape to get away from that shit, from assholes like Bitch.} There was also the fact that Tattletale spooked me, but I couldn't say that out loud.

“Come back inside, Taylor. Please. I personally guarantee I won't let her pull another stunt like that or {\em I'll} quit the team. You're hurt, you're bleeding, your clothes are ripped, and you left your bag with the money upstairs. I'm trained in first aid. At least let us patch you up, get you in some new clothes.”

I glanced down at my arm. I had my right hand clasped around my other wrist, and there was blood on the sleeve of my sweatshirt. And my costume was still upstairs? Fuck.

\quotation{Fine,} I sighed, “But just so you know, I'm only coming back because {\em she} doesn't want me to. I quit, she wins, and I'm not fucking having that.”

Brian smiled and opened the door for me, \quotation{I'll take what I can get.}


\chapter{Insinuation 2.9}

As Brian and I returned to the loft, I felt more than a little apprehensive. It wasn't just that I was going to be around Bitch again, but I was also having to face Lisa and Alec. After shouting and talking about quitting the team, I was turning around and going back. A part of me wanted to apologize, but a larger part of me felt I shouldn't. I had been justified in everything I had said and done, right? Maybe it was just because I wasn't used to violence or raising my voice.

As I'd feared, there was a bit of an awkward silence as we reached the top of the stairs. Bitch was sitting in a chair beside one of the tables, her dogs nowhere to be seen. As she saw me, she scowled, but didn't say anything. Alec grinned as I came back, but I couldn't decide if it was because he was glad or if it was at my expense. I didn't know him well enough to guess either way.

\quotation{Glad you came back,} Lisa told me, a bit of a smile on her face, \quotation{Alec, can you go get the first aid kit? It might be in the storage closet.}

While Alec did that, Brian sat me down on the arm of the couch and I pulled off my sweatshirt to get a better look at the damage. I pulled the bottom of my tank top up around my ribs to get a look at where one of the dogs had been gotten at my stomach and back. My clothes had taken most of the damage, and I'd only suffered three or four shallow-ish scrapes. There was bruising and some raw areas where I felt tender, but I figured I'd recover from that in a day or two. I had a cut on my ear, which would be harder to hide, but I was pretty sure I could keep the incident from my dad without him raising hell.

There was only one spot of real damage, a puncture where it looked like a fang had buried itself deep in the top of my forearm and then dragged an inch or so down towards my wrist as it made its exit. The area around it was already turning colors with bruising. I wasn't sure how deep the puncture was, but I was pretty sure it should have been hurting more than it did. The blood from the injury had trickled down my arm, and was still welling out.

\quotation{Christ,} I said, mostly to myself.

\quotation{That was awesome, you know,} Alec told me, as he returned with the first aid kit, \quotation{I didn't think you had it in you to kick someone's ass.} I glared at him, but he just sat on the back of the sofa, his legs kicking like an excited kid.

\quotation{I think we're going to clean that and stitch it. Tattle's power should give us a better sense of whether stitches are necessary,} Brian said, quietly.

\quotation{Alright,} I agreed.

I would hardly describe getting stitches as a bonding experience, but Bitch more or less stayed quiet throughout the process. We were both sat down and told to sit still while Brian both cleaned and sewed up the hole in my arm and the tear my kick had made in Bitch's ear. Brian insisted I take two Tylenol, though the pain was still limited to a mild ache in my arm. I grudgingly obliged. I'd never liked taking pills, and never felt they made a real difference.

\quotation{You have first aid training?} I inquired, to make conversation and break the tense silence.

Alec complained, \quotation{We all do, Brian made us all take a comprehensive class less than a week after we were gathered as a team. Such a pain in the ass, believe me. He'll make you do it too.}

\quotation{I already did,} I admitted, \quotation{One of the first things I did.} I jumped a little at a snarling from my left, but it was just Rachel cussing as Lisa taped a cotton pad to her ear.

Brian just looked at me and flashed that boyish smile again. I looked away, embarrassed that a guy like him would get pleased like that on my account. He got up to head to the bathroom, garbage from the bandages, sutures, cotton swabs and ointments in his hands.

With Brian gone and Lisa absorbed in trying to patch up Bitch's ear, I was left with Alec. To make conversation, I said, \quotation{Alec. You were going to tell me what you do. You go by Regent, right?}

\quotation{The name is a long story, but what I do is this.} He looked over his shoulder at Brian, who was returning from the washroom with a damp washcloth in hand. Brian, mid-stride, stumbled and fell onto the floor.

\quotation{Way to look good in front of the new girl, gimpy!} Alec mocked his teammate, laughing. Grateful for the break in the tension, I couldn't help but laugh too. While Alec continued laughing, Brian got to his feet and ran up to the smaller boy, at which point he got Alec in a headlock and began punching him in the shoulder repeatedly. This abuse only made Alec laugh harder in between his cries of pain.

Lisa turned to me, smiling at the prank and play fighting between the boys, \quotation{It's a bit complicated to explain, but basically, Alec can get into people's nervous systems. This lets him fire off impulses that set off reflexes or make body parts jerk into motion. It's not a dramatic power, but with timing, he can make someone fall over midstep, drop something, lose their sense of balance or pull the trigger on a gun.}

I nodded, absorbing the information. It sounded very underwhelming to me, but I was willing to admit I could be underestimating it.

\quotation{Well,} I said, after a long pause, \quotation{I think I pretty much get what everyone can do, then. Correct me if I'm wrong, but Bitch can turn those dogs into those freakish monsters I saw the other night?}

Sitting a few feet away, Bitch muttered, \quotation{They aren't freakish.}

Lisa answered my question, ignoring her. “{\em Rachel} can do it with any dog, actually,” she said, stressing the name, \quotation{And no codenames when we're not in costume, ‘kay? Get in the habit of using the right name at the right times, and it's that much harder to slip.}

It was hard to think of Rachel by her real name. Bitch seemed really fitting given the stunt she had pulled. I apologized to Lisa, \quotation{Sorry.}

Lisa gave a small nod in response, then told me, \quotation{She can use her power on any dog, but only Brutus, Judas and Angelica are trained well enough that they'll listen to her when they're pumped up.}

Ah, so that was it. \quotation{And Brian makes that oily darkness that screws up your hearing. The Parahumans wiki said it was darkness generation.}

Brian smiled, \quotation{I put that into the wiki myself. It's not wrong, but it does catch people off guard when they think they know what you can do, and there's something more to it.}

Lisa added, \quotation{It's not just hearing. It also cuts off radio signals and dampens the effects of radiation.}

\quotation{That's what her power tells her, anyways. I haven't had much chance to test that part of things. I get by as is,} Brian said. He turned his hand palm up and created a handful of the darkness. It was like smoke, but so absolutely black that there was no texture to it. It was like someone had taken a scalpel to reality and the blackness was what was there when everything else was gone. I couldn't even gauge the dimensions of it, unless I looked at it from a different perspective. Even then, with the way the darkness shifted and billowed like smoke, it was hard to judge the shape.

More of it just kept pouring from his hand, climbing upwards to cover the top of the room. As the light from the windows near the upper edges of the room and the florescent bars on the ceiling was cut off, the room got a great deal darker.

He closed his hand into a fist, and the darkness thinned out and disintegrated into strands and tatters, and the room brightened again. I looked at the light coming in from the windows and was surprised it wasn't later.

\quotation{What time is it?} I asked.

\quotation{Nineteen minutes before five,} Lisa said. She didn't look at a watch or a clock as she said it, which was unsettling. It was a reminder that her power was constantly available to her.

Brian asked me, “Do you have somewhere you need to be?

\quotation{Home, I guess,} I admitted, \quotation{My dad will wonder where I am.}

\quotation{Call him,} Lisa suggested, \quotation{Now that the introductions are over with, you can just hang out for a bit, if you want.}

\quotation{We could order pizza,} Alec suggested. Then when Lisa, Brian and Bitch all made faces, he added, \quotation{Or maybe everyone's sick of pizza and we could order something else.}

\quotation{Stick around?} Brian made it a question.

I glanced at Bitch. She was sitting on the table behind one of the couches and looking like a mess, with a bloody bandage over one ear, blood smeared below her nose and lip, and a bit of green around the gills that suggested she was feeling a little worse for wear. With her in that state, I didn't feel particularly threatened. Staying meant I could work to get things more copacetic and maybe dig for a bit more information. I'd also missed socializing with people -- even if it was under false pretenses with a group that included an apparent sociopath. It had been a sucky day. Just chilling out sounded good.

\quotation{Okay,} I decided, \quotation{Yeah, I think I'd like to.}

\quotation{Phone's in the kitchen if you want to call your dad,} Lisa said.

I looked over my shoulder as I headed across the loft. The others got settled on the couches, with Alec turning on the TV while Lisa and Brian took a second to clean up.

I found the phone and dialed my dad.

\quotation{Hey dad,} I said, when I heard the phone being picked up.

\quotation{Taylor. Are you alright?} He sounded worried. It {\em was} unusual, I supposed, my not being home when he got back from work.

\quotation{I'm fine, dad. Is it cool if I hang out with some people tonight?}

There was a pause.

\quotation{Taylor, if there's anyone that's making you make this call\ldots{} the bullies or someone else, tell me everything is fine. If you're not in trouble, tell me your mother's full name.}

I felt momentarily embarrassed. Was it so unusual for me to hang out with people? I knew my dad was just trying to keep me safe, but it was bordering on the ridiculous.

\quotation{Annette Rose Hebert,} I told him, \quotation{Really dad, it's cool.}

\quotation{You're really okay?}

My gaze roved over the kitchen, taking in the details, as I gave him my assurances.

\quotation{Better than ever. I kind of made some friends,} I said.

My eyes settled on their dining room table. There was a stack of money, wrapped with a paper band just as the money in the lunchbox had been. Beside the money, plain as day, was the dark gray metal of a handgun.

My attention caught by the gun, I only barely caught my dad's question. \quotation{What are they like?}

\quotation{They seem like good people,} I lied.


\chapter{Interlude 2}

There were very few things, in Victoria Dallon's estimation, that were cooler than flying. The invisible forcefield that extended a few millimeters over her skin and clothes just made it better. The field kept the worst of the chill from touching her, but still let her feel the wind on her skin and in her hair. Bugs didn't splat against her face like they did against car windshields, even when she was pushing eighty miles an hour.

Spotting her target, she whooped and plunged for the ground, gaining speed where anyone else would be slowing down. She hit the asphalt hard enough to crack it and send fragments of it into the air, touching ground with her knee and foot, one arm extended. She stayed in that kneeling position for just heartbeats, letting her platinum curls and the cape that was draped over one of her shoulders flutter in the wake of air that had followed her descent. She met the eyes of her quarry with a steely glare.

She'd practiced that landing for {\em weeks} to get it right.

The man was a twenty something Caucasian with a shaved head, a dress shirt with the sleeves rolled up, jeans and work boots. He took one look at her and bolted.

Victoria grinned as he disappeared down the far end of the alley. She rose from her kneeling position, dusted herself off and ran her fingers through her hair to tidy it. Then she raised herself a foot off the ground and flew after him at an easy forty five miles an hour.

It didn't take a minute to catch him, even with the head start she had given him. She flew just past him, grazing him. An instant later, she came to a dead stop, facing him. Again, the wind made for a dramatic flourish as it stirred her hair, her cape and the skirt of her costume.

\quotation{The woman you attacked was named Andrea Young,} she spoke.

The man looked over his shoulder, as if gauging his escape routes.

\quotation{Don't even think about it, fugly,} she told him, \quotation{You know I'd catch you, and trust me, I'm already pissed off enough without you wasting my time.}

\quotation{I didn't do anything,} the man snarled.

\quotation{Andrea Young!} Victoria raised her voice. As she shouted, she exercised her power. The man quailed as though she'd slapped him. \quotation{A black college student was beaten so badly she needed medical attention! Her teeth were knocked out! You're trying to tell me that you, a skinhead with swollen knuckles, someone who was in the crowd watching paramedics arrive with an expression bordering on glee, you didn't do anything!?}

\quotation{I didn't do nothing worth caring about,} he sneered. His bravado was tempered by a second look over his shoulder, as though he'd very much like to be elsewhere right that moment.

She flew forward, her fists catching him by the collar. For just a moment, she contemplated slamming him up against a wall. It would have been fitting and satisfying to shove him hard enough against the brick to crack it, then drop him into the dumpster that sat at the wall's base.

Instead, she pulled up a little, bringing the two of them to a stop. They were now just high enough above the ground that he'd feel uncomfortable with the height. The dumpster, mostly empty, was directly below him, but she doubted he was paying attention to anything but her.

\quotation{I think it's a safe bet to say you're a member of Empire Eighty-Eight,} she told him, meeting his eyes with a hard stare, “or at least, you've got some friends who are. So here's what's going to happen. You're going to either tell me everything the triple-E's have been up to, or I'm going to break your arms and legs and {\em then} you're going to tell me everything.”

As she spoke, she ratcheted up her power. She knew it was working when he started squirming just to avoid her gaze.

\quotation{Fuck you, you can't touch me. There's laws against that shit,} he blustered, staring fixedly over one shoulder.

She turned up her power another notch. Her body thrummed with current -- waves of energy that anyone in her presence would experience as an emotional charge of awe and admiration. For those with a reason to be afraid of her, it would be a feeling of raw intimidation instead.

\quotation{Last chance,} she warned him.

Unfortunately, fear affected everyone differently. For this particular asshole, it just made him dig in his heels and become obstinate. She could see it in his body language before he opened his mouth -- this was the sort of guy who reacted to anything that spooked or unsettled him with an almost mindless refusal to bend.

\quotation{Lick my hairy, sweaty balls,} he snarled, before punctuating it with a spat, \quotation{Cunt.}

She threw him. Since she could bench press a cement mixer, though it was hard to balance something so large and unwieldy, even a casual toss on her part could get some good distance. He flew a good twenty five or thirty yards down the back road before hitting the asphalt, and rolled for another ten.

He was utterly for still for long enough that Victoria had begun to worry that he'd somehow snapped his neck or broken his spine as he'd rolled. She was relieved when he groaned and began to pull himself to his feet.

\quotation{Ready to talk?} she asked him, her voice carrying down the alley. She didn't move forward from where she hovered in the air, but she did let herself drop closer to the ground.

Pressing one hand against his leg to support himself as he straightened up, he raised his other hand and flipped her the bird, then turned and began to limp down the alley.

What was this asshole thinking? That she would just let him go? That, what, she would just bend to his witless lack of self preservation? That she was helpless to do any real harm to him? To top it off, he was going to insult her and try to walk away?

\quotation{Screw you too,} she hissed through her teeth. Then she kicked the dumpster below her hard enough to send it flying down the little road. It rotated lazily through the air as it arced towards the retreating figure, the trajectory and rotation barely changing as it knocked him flat. It skidded to a halt three to five yards beyond him, the metal sides of the dumpster squealing and sparking as it scraped against the asphalt.

This time, he didn't get up.

\quotation{Fuck,} she swore, \quotation{Fuckity fuck fuck.} She flew to him and checked for a pulse. She sighed, and then headed to the nearest street. She found the street address, grabbed her cell from her belt and dialed.

\quotation{Hey sis? Yeah, I found him. That's, uh, sort of the problem. Yeah. Look, I'm sorr- ok, can we talk about this later? Yeah. I'm at Spayder and Rock, there's this little road that runs behind the buildings. Downtownish, yeah. Yeah? Thanks.}

Victoria returned to the unconscious skinhead, checked his pulse, and listened intently for changes in his breathing. It took a very long five minutes for her sister to arrive.

“{\em Again}, Victoria?” the voice disturbed her from her contemplations.

\quotation{Use my codename, please,} Victoria told the girl. Her sister was as different from her as night was from day. Where Victoria was beautiful, tall, gorgeous, blonde, Amy was mousy. Victoria's costume showed off her figure, with a white one-piece dress that came to mid-thigh (with shorts underneath) an over-the shoulder cape, high boots and a golden tiara with spikes radiating from it, vaguely reminiscent of the sun's rays or the statue of liberty. Amy's costume, by contrast, was only a shade away from being a burka. Amy wore a robe with a large hood and a scarf that covered the lower half of her face. The robe was alabaster white and had a medic's red cross on the chest and the back.

\quotation{Our identities are public,} Amy retorted, pushing the hood back and scarf down to reveal brown frizzy hair and a face with freckles spaced evenly across it.

\quotation{It's the principle of the thing,} Victoria replied.

“You want to talk about {\em principles,} Glory Girl?” Amy asked, in the most sarcastic tone she could manage, \quotation{This is the sixth -- sixth! -- time you've nearly killed someone. That I know about!}

\quotation{I'm strong enough to lift a SUV over my head,} Victoria muttered, \quotation{It's hard to hold back all the time.}

\quotation{I'm sure Carol would buy that line,} Amy said, making it clear in her tone she wasn't, \quotation{But I know you better than anyone. If you're having trouble holding back, the problem isn't here -} she poked Victoria in the bicep. \quotation{It's here-} she jabbed her sister in the forehead, hard. Victoria didn't even blink.

\quotation{Look, can you just fix him?} Victoria pleaded.

\quotation{I'm thinking I shouldn't,} Amy said, quietly.

\quotation{What?}

\quotation{There's consequences, Vicky. If I help you now, what's going to stop you from doing it again? I can call the paramedics. I know some good people from the hospital. They could probably fix him up alright.}

\quotation{Hey, hey, hey,} Victoria said, \quotation{That's not funny. He goes to the hospital, people ask questions.}

\quotation{Yeah, I'm well aware,} Amy said, her voice hushed.

\quotation{This isn't, like, me getting grounded. I'd get pulled into court on charges of aggravated assault and battery. That doesn't just fuck with me. It fucks with our family, all of New Wave. Everything we've struggled to build.}

Amy frowned and looked at the fallen man..

\quotation{I know you're not keen on the superhero thing, but you'd really go that far? You'd do that to us? To me?}

Amy pointed a finger at her sister, “That's not me. It's not my fault we're at this point. It's you. You're crossing the line, going too far. Which is {\em exactly} what people who criticize New Wave are scared of. We're not government sponsored. We're not protected or organized or regulated in the same way. Everyone knows who we are under our masks. That means we have to be accountable. The responsible thing for me to do, as a member of this team, is to let the paramedics take him, and let the law do as it sees fit.”

Victoria abruptly pulled Amy into a hug. Amy resisted for a moment, then let her arms go limp at her sides.

\quotation{This isn't just a team, Ames,} Victoria told her, “We're a family. We're {\em your} family.”

The man lying just a matter of feet away stirred, then groaned, long and loud.

\quotation{My adoptive family,} Amy mumbled into Victoria's shoulder, \quotation{And stop trying to use your frigging power to make me all squee over how amazing you are. Doesn't work. I've been exposed so long I'm immune.}

\quotation{It hurts,} the man moaned.

\quotation{I'm not using my power, dumbass,} Victoria told Amy, letting her go, \quotation{I'm hugging my sister. My awesome, caring and merciful sister.}

The man whined, louder, \quotation{I can't move. I feel cold.}

Amy frowned at Victoria, \quotation{I'll heal him. But this is the last time.}

Victoria beamed, \quotation{Thank you.}

Amy leaned over the man and touched her hand to his cheek, “Slingshot break to his ribs, fractured clavicle, broken mandible, broken scapula, fractured sternum, bruised lung, broken ulna, broken radius -“

\quotation{I get the point,} Victoria said.

\quotation{Do you?} Amy asked. Then she sighed, \quotation{I wasn't even halfway down the list. This is going to take a little while. Sit?}

Victoria crossed her legs and assumed a sitting position, floating a half foot above the ground. Amy just knelt where she was and rested her hand on the man's cheek. The tension went out of his body and he relaxed.

\quotation{How's the woman? Andrea?}

\quotation{Better than ever, physically,} Amy replied, \quotation{I grew her new teeth, fixed everything from the bruising to the scrapes, and even gave her a head to toe tune-up. Physically, she'll feel on top of the world, like she had been to a spa and had the best nutritionist, best fitness expert and the best doctor all looking after her for a straight month.}

\quotation{Good,} Victoria said.

\quotation{Mentally? Emotionally? It's up to her to deal with the aftermath of a beating. I can't affect the brain.}

\quotation{Well-} Victoria started to speak.

\quotation{Yeah, yeah. Not can't. Won't. It's complicated and I don't trust myself not to screw something up when I'm tampering with someone's head. That's it, that's all.}

Victoria started to say something, then shut her mouth. Even if they weren't related by blood, they {\em were} sisters. Only sisters could have these sorts of recurring arguments. They had gone through a dozen different variations on this argument before. As far as she was concerned, Amy was doing herself a disservice by not practicing using her powers on the brain. It was only a matter of time before her sister found herself in a situation where she needed to do some emergency brain surgery and found herself incapable. Amy, for her part, refused to even discuss it.

She didn't want to raise a sensitive issue when Amy was in the process of doing her a major favor. To change the subject, Victoria asked, \quotation{Is it cool if I question him?}

\quotation{Might as well,} Amy sighed.

Victoria tapped the man a few times on the forehead to get his attention. He could barely move his head, but his eyes lolled in her direction.

\quotation{Ready to answer my questions, or do me and my sister just walk away and leave you like this?}

\quotation{I\ldots{} sue you, he gasped out, then managed an added, “Whore.}

\quotation{Try it. I'd just love to see a skinhead with a few broken bones go up against a superheroine whose mom just happens to be one of the best lawyers in Brockton Bay. You know her, right?}

\quotation{Brandish,} he said.

\quotation{That's her name in costume. Normally she's Carol Dallon. She'd kick your ass in court, believe me,} Victoria said. {\em She} believed it. What the thug didn't understand was that even if he lost the case, the media circus that would be stirred up would do more damage than anything else. But she didn't need to inform him of that. She asked him, \quotation{So do I get my sister to leave you as you are, or are you willing to trade some information for relief from months of incredible pain and a lifetime of arthritis and stiffness in your bones?}

\quotation{And erectile dysfunction,} Amy said, just loud enough for the thug to hear her, \quotation{You fractured your ninth vertebra. That's going to affect all nerve function in extremities below your waist. If I leave you like you are, your toes will always feel a little numb, and you'll have a hell of a time getting it up, if you know what I mean.}

The skinhead's eyes widened a fraction, \quotation{You're fucking with me.}

\quotation{I have an honorary medical license,} Amy told him, her expression solemn, \quotation{I'm not allowed to fuck with you about stuff like that. Hippocratic oath.}

\quotation{Isn't that \quote{do no harm}?} the thug asked. Then he groaned, long, loud and with the slightest rattle in his breath, as she removed her hand from his body.

\quotation{That's just the first part of it, like how freedom of speech and the right to bear arms is just the first part of a very long constitution. It doesn't look like he's cooperating, Glory Girl. Should we go?}

\quotation{Fuck!} the man shouted, then winced, tenderly touching his side with one hand, \quotation{I'll tell you. Please, just\ldots{} do what you were doing. Touch me and make the pain go away, put me back together. Fix me?}

Amy touched him. He relaxed, and then he started talking.

\quotation{Empire Eighty-Eight is extending into the Docks on Kaiser's orders. Lung's in custody, and whatever happens, the ABB is weaker than it was. That means there's territory for grabs, and the Empire sure ain't making progress downtown.}

\quotation{Why not?} Victoria asked him.

\quotation{This guy, Coil. Don't know what his powers are, but he's got a private army. Ex-military, all of 'em. At least fifty, Kaiser said, and every one of 'em has top notch gear. Their armor's better than kevlar. You shoot 'em, they're back up in a few seconds. ‘Least when you shoot a pig, you can be pretty sure you broke a few ribs. But that's not the fucked up thing. These guys? They've got these lasers hooked up to the machine guns they carry around. If they don't think bullets are doing it, or if they're after people who are behind cover, they fire off these purple laser beams that can cut through steel. Tear through any cover you're standing behind and burn through you too.}

\quotation{Yeah. I know about him. His methods get expensive,} Victoria said, \quotation{Top of the line soldiers, top of the line gear.}

The thug nodded weakly, \quotation{But even with money to burn, he's fighting us over Downtown territories. Constant tug of war, neither of us making much headway. Been going on for months. So Kaiser thinks we should take the Docks now that the ABB are on the outs, gain some ground somewhere easier. Don't know any more than that, as far as his plans.}

\quotation{Who else is up to something? Faultline?}

\quotation{The bitch with the freaks in her crew? She's a mercenary, different goals. But maybe. If she wanted to branch out, now would be the time to do it. With her rep, she'd even do alright.}

\quotation{Then who? There's a power vacuum in the docks. Kaiser's declared he wants to seize it, but I'm willing to bet he's warned you about others making a play.}

The skinhead laughed, then winced, “Are you dense, girl? Everyone's going to make a play. It's not just the major gangs and teams that are looking for a slice of the pie, there. It's {\em everyone}. The Docks are ripe for the taking. The location's worth as much money as you'd get downtown. It's the go to place if you want to buy black market. Sex, drugs, violence. And the locals are already used to paying protection money. It's just a matter of changing who they pay to. The Docks are rich territory, and we're talking the potential for a full scale fucking war over it.”

He looked up at the blond superheroine and laughed. Her lips set into a firm line.

He continued, \quotation{You want to know my guess? Empire Eighty Eight is going to take the biggest slice of the Docks, because we're strong enough to. Coil's going to stick his thumb in just to spite us, ABB is going to hold on to some. But you're also going to have a bunch of the little guys trying to take something for themselves. Über and Leet, Circus, the Undersiders, Squealer, Trainwreck, Stain, others you've never heard of? They're going to stake out their ground, and one of two things is going to happen. Either there's war, in which case civilians get hurt and things get bad for you, or there's alliances between the various teams and solo villains and shit gets even worse for you.}

He broke into laughter yet again.

\quotation{Come on, Panacea,} Victoria said as she stood up, touched ground with her boots and brushed her skirt straight, \quotation{We've gotten enough.}

\quotation{You sure? I'm not done yet,} Amy told her.

\quotation{You fixed the bruises and scrapes, broken bones?} Everything that could get her in trouble, in other words.

“Yeah, but I didn't fix {\em everything},” Amy replied.

\quotation{Good enough,} Victoria decided.

\quotation{Hey!} the skinhead shouted, \quotation{The deal was you'd fix me if I talked! Did you fix my cock?} He tried to struggle to get to his feet, but his legs buckled under him, \quotation{Hey! I can't fuckin' walk! I'll fucking sue you!}

Victoria's expression changed in an instant, and her power flooded out, blindsiding the thug. For an instant, his eyes were like those of a panicked horse, all whites, rolling around, unfocused. She grabbed him by the shirt collar, lifted him up and growled into his ear, her voice just above a whisper, “Try it. My sister just healed you\ldots{} most of you, with a touch. Did you ever wonder what else she could do? Ever think, maybe, she could break you just as easily? Or change the color of your skin, you racist fuck? I'll tell you this, I'm not {\em half} as scary as my little sister is.”

She let him go. He collapsed in a heap on the ground.

As the two sisters walked away, Victoria pulled her cell phone out of a pouch on her belt with her free hand. Turning to Amy, she said, \quotation{Thank you.}

\quotation{Play safe, Victoria. I can't bring people back from the dead, and once you've gone that far\ldots{}}

\quotation{I'll be good. I'll be better,} Victoria promised as she dialed with one hand. She put the phone to her ear, \quotation{Hello? Emergency services? Requesting special line. New Wave, Glory Girl. Incapacitated criminal for you to pick up, no powers. No, no rush, I can hold.}

Looking over her shoulder, Victoria noted the thug, still floundering and half-crawling, \quotation{He's not going to get up?}

\quotation{He'll be numb from the waist down for another three hours. His left arm will be iffy for about that long, too, so he's not going to move unless he can drag himself somewhere with just one limb. He'll also have numb toes for a good month or so, too,} Amy smiled.

\quotation{You didn't actually\ldots{}}

\quotation{No. Nothing was broken, and I didn't screw up anything, beyond a temporary numbness. But he doesn't know that. Fear and doubt will complete the effect, and the suggestion becomes a self fulfilling prophecy.}

\quotation{Amy!} Victoria laughed, hugging her sister with one arm, \quotation{Weren't you just saying you weren't going to mess with people's heads?}



