\part[p1]{Agitation}

\chapter[c3.1]{Agitation 3.1}

Tuesday morning found me running again, first thing. I woke up at my regular time, apologized to my dad for not having breakfast with him, and headed out the door, hood of my sweatshirt up to hide the mess of my uncombed hair.

There was something appealing about being out and about before the city had woken up. I didn't usually get out quite this early, so it was a refreshing change. As I headed east at a brisk jog, there were no cars or people on the street. It was six thirty in the morning, and the sun had just finished rising, so the shadows were long. The air was cool enough for my breath to fog. It was like Brockton Bay was a ghost town, in a good way.

My training regimen had me running every morning, and alternating between more running and doing other exercises in the afternoons, depending on which day of the week it was. The primary goal was to build my stamina. In February, Sophia had goaded some boys into trying to catch me, I think the goal had been to duct tape me to a telephone pole. I had escaped, helped mostly by the fact that the boys hadn't really cared enough to run after me, but I found myself winded after having run just a block. It had been a wake-up call that came about just when I was starting to think about going out in costume. Not long after, I had started training. After a few starts and stops, I had settled into a routine.

I was more fit, now. While I could hardly say I was heavy, before, I'd had the unfortunate combination of a slight bulge for a belly, small breasts and broomstick-thin arms and legs. It had added up to me looking something like a frog forced to stand up on its hind legs. Three and a half months had burned away the body fat, leaving me very lean, and had given me the stamina to run at a steady jog without leaving me panting for breath.

I didn't aim to just jog, though. I steadily increased my pace with every block I ran as I headed towards the water. By the fifth block, I was running.

My general approach was not to get too worried about counting the miles or measuring the times. That just felt like it was distracting me from my own awareness of my body and its limits. If it felt too easy, I just pushed myself a step further than I had the previous day.

The route I took varied every day, at my father's insistence, but it usually took me to the same place. In Brockton Bay, going east took you to one of two places. You either ended up at the Docks, or you ended up at the Boardwalk. Because most areas of the Docks were not the sort of place that you just breezed through, given the vagrants, gang members and general crime, I stuck to main roads leading past the Docks and to the Boardwalk. It was usually close to seven by the time I got to the bridge that went over Lord Street. From there, it was a block to the Boardwalk.

I slowed down as the sidewalk ended and the wooden platform began. Though my legs were aching and I was out of breath, I forced myself to keep a low and steady pace rather than just stop.

Along the boardwalk, people were starting their day. Most places were still closed, with the top notch security systems, steel shutters and iron grates protecting all of the expensive stores, but there were cafes and restaurants opening up. Other stores had vans parked in front, and were busy loading in their shipments. There were only a few people out and about, which made it easy to find Brian.

Brian was leaning on the wooden railing, looking over the beach. Balanced on the railing next to him was a paper bag and a cardboard tray with a coffee in each of the four pockets. I stopped beside him, and he greeted me with a broad smile.

\quotation{Hey, you're right on time,} Brian said. He looked different than he had when I saw him on Monday. He was wearing a sweater under a felt jacket, his jeans didn't have any rips or tears in them, and his boots were shined. On Monday, he had given me the impression of a regular person who lived at the Docks. The fashionable, well fit clothes he wore today made him look like someone who belonged on the Boardwalk alongside the customers who shopped in stores where nothing cost less than a hundred dollars. The contrast and the ease with which he seemed to make the transition was startling. My estimation of Brian rose a notch.

\quotation{Hey,} I said, feeling just a touch embarrassed at having taken so long to respond, and feeling painfully under-dressed in his presence. I hadn't expected him to dress so well. I hoped my being out of breath was enough of an excuse for the delay in response. There was nothing I could do about feeling unfashionable.

He gestured towards the paper bag, \quotation{I got donuts and croissants from the cafe over there, and a coffee if you want it.}

\quotation{I want,} I said, then I felt dumb for the awkward lapse into caveman speak. I blamed the early hour of the day. To try and save face, I added, \quotation{Thanks.}

I fished out a sugar-dusted donut and bit into it. I could tell right away that it wasn't the kind of donut that was mass produced at some central factory and delivered overnight to the shops for baking in the morning. It was freshly made, probably right at the store a block away, sold right out of the oven.

\quotation{So good,} I said, sucking the sugar from my fingertips before reaching for one of the coffees. Seeing the logo, I looked over at the cafe and asked, \quotation{Don't coffees there cost, like, fifteen dollars a cup?}

Brian chuckled a little, \quotation{We can afford it, Taylor.}

It took me a second to process the idea, and as I made the connection, I felt like an idiot. These guys were raking in thousands of dollars on a given job, and they had given me two thousand dollars up front. I wasn't willing to spend the money, knowing where it came from, so it was just sitting in the cubbyhole I kept my costume in, nagging at me. I couldn't tell Brian that I wasn't spending it, either, without risking having to explain why.

\quotation{Yeah, I guess,} I said, eventually. I leaned my elbows on the wooden railing beside Brian and stared out over the water. There were a few diehard windsurfers just getting ready to start the day. I guess it made sense, since there would be the occasional boat going out on the water, later.

\quotation{How's your arm?} He asked.

I extended my arm, clenched my fist and relaxed it to demonstrate, \quotation{Only hurts when I flex it.} I didn't tell him that it had been hurting badly enough to cost me some sleep last night.

\quotation{We'll leave the stitches in for about a week, I think, before we take them out,} Brian said, \quotation{You can go to your doctor and have him do it, or drop by and I'll take care of it.}

I nodded. A turn of the salt-water and seaweed scented wind blew my hood back, and I took a second to push my hair out of my face and pull my hood back up.

\quotation{I'm sorry for Rachel and that whole incident last night} Brian said, \quotation{I wanted to apologize sooner, but I figured it would be a bad idea to bring it up while she was in earshot.}

\quotation{It's okay,} I said. I wasn't sure it was, but it wasn't really his fault. I tried to put my thoughts into words, \quotation{I think\ldots{} well, I guess I expected to have people attack me from the moment I put on a costume, so I shouldn't be surprised, right?}

Brian nodded, but didn't say anything, so I added, \quotation{It caught me a little off guard that it came from someone that's supposedly on my team, but I'm dealing.}

\quotation{Just so you know,} Brian told me, \quotation{Just from what I saw after you left last night and as people were waking up this morning, Rachel seems to have stopped protesting quite as loudly or often about the idea of having someone new join the team. She's still not happy about it, but I would be surprised if there was a repeat performance.}

I laughed, a little too abruptly and high pitched than I would have liked, \quotation{God, I hope not.}

\quotation{She's kind of a special case,} Brian said, \quotation{I think that growing up the way she did kind of messed her up. No family, too old and, uh, not really attractive enough to be a good candidate for adoption. I feel bad saying that, but that's the way those things work, you know?} He glanced over his shoulder at me.

I nodded.

\quotation{So she spent a good decade in foster care, no fixed place to live, fighting tooth and nail with the other foster kids for even the most basic luxuries and possessions. My guess? She was screwed up before she got her powers, and with things happening the way they did, her powers pushed her into the deepest end of the antisocial pool.}

\quotation{Makes sense,} I said, then I added, \quotation{I read her page on the wiki.}

\quotation{So you've got the gist of it,} Brian said, \quotation{She's a handful to deal with, even for me, and I think she actually considers me a friend\ldots{} or as much a friend as someone like her can have, anyways. But if you can at least tolerate her, you should see we've got a pretty good thing going with the team.}

\quotation{Sure,} I said, \quotation{We'll give it a shot, anyways.}

He smiled at me, and I dropped my gaze, embarrassed.

I spotted a crab scuttling across the beach almost directly below us. I reached out with my power and stopped it in its tracks. Though I didn't need to, I extended my finger and pointed at it, then waved my finger lazily as I made the crab follow where my my index finger was pointing. Since Brian and I were both leaning over the railing, and there was practically nobody on the Boardwalk that wasn't busy with work or getting their store opened for the day, I was pretty certain nobody else would figure out what I was doing.

Brian saw the crab dancing in circles and figure eights and smiled. Conspiratorially, he leaned closer to me and whispered, \quotation{You can control crabs, too?}

I nodded, feeling just a bit of a thrill at how we were huddled like this, sharing secrets while the people around us were totally in the dark. I told him, \quotation{I used to think I could control anything with an exoskeleton or shell. But I can control earthworms too, among other things, and they don't have shells. I think all it takes is that they have to have very simple brains.}

I made it run in circles and figure eights for a short while longer, then released it to go about its business.

\quotation{I should bring the others their morning coffee before they come looking for me. Want to come with?} Brian asked.

I shook my head, \quotation{I gotta get home and get ready for school.}

\quotation{Ah, right,} Brian said, \quotation{I forget about stuff like that.}

\quotation{You guys don't go?}

\quotation{I take courses online,} Brian said, \quotation{My folks think it's so I can hold a job to pay for my apartment\ldots{} which is kind of true. Alec dropped out, Rachel never went, and Lisa already applied for and tested for her G.E.D. Cheated using her power, but she has it.}

\quotation{Ah,} I said, my focus more or less dwelling on the idea that Brian had an apartment. Not the fact that Grue the successful supervillain had an apartment -- Lisa had mentioned that to me -- but that Brian the teenager with parents and schoolwork to focus on did. He kept changing my frame of reference for trying to figure him out.

\quotation{Here, a gift,} he said, as he reached into his pocket and then extended his hand.

I felt a moment of trepidation at the notion of accepting another gift. The two grand they had given me was a weight on my conscience already. Still, it would look bad if I didn't accept. I made myself put my hand under his, and he dropped a key with a short beaded chain looped through it into my palm.

\quotation{That's to our place,} he told me, \quotation{And I mean that. Ours as in yours too. You're free to come by any time, even if nobody is there. Kick back and watch TV, eat our food, track mud on our floor, yell at the others for tracking mud on the floor, whatever.}

\quotation{Thank you,} I said, surprising myself by actually meaning it.

\quotation{You going to come by after school, or should I meet you here again tomorrow morning?}

I thought on it for a second. Last night, not long before I'd left, Brian and I had gotten to talking about our training. When I had mentioned my morning runs, he had suggested meeting me regularly. The idea was to keep me up to date, since I wasn't living at the group's hideout like Lisa, Alec and Rachel were. It had made sense, and I'd agreed. It didn't hurt that I liked Brian the most of anyone in the group. He was easier to relate to, somehow. That wasn't to say I didn't like Lisa, but just being around her made me feel like I had the Sword of Damocles hanging over my head.

\quotation{I'll come by later,} I decided aloud, knowing I might chicken out if I didn't commit somehow. Before we could get caught in another thread of conversation, I gave him a quick wave and started my run back, the key to their place clenched in my hand.

Heading back home and preparing for school left me with a gradually increasing feeling of dread, like a weight sitting on my chest. I'd been trying not to think of Emma's taunting and my fleeing from the school with tears on my face. I had spent an hour or two tossing and turning in bed, the event replaying over my head while the throbbing of my wrist jarred me awake every time I started to drift off. Beyond that, I had been pretty successful in avoiding thinking about it. Now that the prospect of going back was looming, though, it was impossible not to dwell on the subject as I headed home, got ready and caught the bus.

I couldn't help but dwell on the coming day. I still had to face the consequences of missing two afternoons. That was a biggie, especially since I had missed the due date for handing in my art project. I realized my art project had been in my bag, and the last time I had seen my bag had been when Sophia was standing on it, smirking at me.

There was also the issue of going to Mr. Gladly's class. That usually sucked enough, what with Madison being in that class and my having to do group work with the likes of Sparky and Greg. Knowing that I had to sit there and listen to Mr. Gladly teach when I'd seen him blatantly turn his back to me when I was being bullied\ldots{} that sucked more.

This wasn't the first time I'd needed to psych myself up to going to school. Deceive myself into going and staying. The worst days had been back in my first year at high school, when the wounds of Emma's betrayal were still fresh and I wasn't yet experienced enough to anticipate the variety of things they could come up with. Back then, it had been terrifying, because I hadn't yet known what to expect, didn't know where, when or if they would draw the line. It had been hard, too, to go back in January. I'd spent a week in the hospital under psychiatric observation, and I'd known that everyone else had heard the story.

I stared out the window of the bus, watching the people and the cars. On days like this, after being publicly humiliated, getting myself to the point where I was willing to walk through the door was about making deals with myself and trying to look past the school day. I told myself that I would go to Mrs. Knott's computer class. None of the Trio would be there, it was usually pretty easygoing, and I could take the time to browse the web. From there, it was just a matter of convincing myself to walk down the hall to Mr. Gladly's class.

If I just made myself do that, I promised myself, I would give myself a treat. A lunch break spent reading one of the books I'd been saving, or a rare snack bought from the store after school. For the afternoon classes, I'd inevitably come up with something else to look forward to, like watching a TV show I liked or working on my costume. Or, I thought, maybe I could just look forward to hanging out with Lisa, Alec and Brian. Outside of the part where I nearly got mauled by Bitch's dogs, it had been a nice night. Thai food, five of us lounging on two couches, watching an action movie on a huge entertainment system with surround sound. I wasn't forgetting what they were, but I rationalized that I had no reason to feel bad about spending time with them when we were -- for all intents and purposes -- just a group of teenagers hanging out. Besides, it was for a good cause, if it meant they relaxed around me and maybe revealed secrets. Right?

As I got off the bus, a pair of old notebooks in one hand, I just kept all that in mind. I could relax in Mrs. Knott's class, and then I just had to sit through three 90 minute classes. Maybe, it occurred to me, I could try and find and talk to my art teacher over the lunch break. It would mean staying out of the trio's way, and I could maybe work something out as far as doing another project or at least not getting a zero. My marks were okay enough that I could probably manage a passing grade with a zero on the midterm project, but still, it would help. I wanted to do more than just pass, especially with all this crap I had to put up with.

Mrs. Knott arrived at the classroom around the same time I did, and unlocked the room to let us file in. As one of the last of fortyish students to arrive, I'd wound up at the back of the crowd. While I waited for enough space to open up at the door, I saw Sophia talking to three of the girls from the class. It looked like she had just come from her track practice. Sophia was dark skinned with black hair normally long enough to reach to the small of her back, though she currently had it in a ponytail. I couldn't help but resent the fact that even with her being sweaty, dusty, and a notorious bitch, pretty much every guy in the school would still pick her over me.

She said something, and all of the girls laughed. Even though I knew, rationally, that I probably wasn't on the list of their top five things to talk about and that they likely weren't talking about me, I felt my heart sink. I moved up towards the jam of students waiting to get into the door, to break the line of sight between myself and the girls. It didn't quite work. As a group of students entered the room, I saw Sophia looking at me. She made an exaggerated pouting expression, drawing one fingertip in a line from the corner of her eye down her cheek like a mock tear. One of the other girls noticed and chuckled, leaned closer to Sophia as Sophia whispered something in her ear, then they both laughed. My cheeks flushed with humiliation. Sophia gave me a final smirk and turned to saunter away while the other girls filed into the classroom.

Kicking myself even as I did it, I turned away and walked back down the hall towards the front doors of the school. I knew it would be that much harder to go back tomorrow. For one and three-quarter school years, I had been putting up with this shit. I'd been going against the current for a long time, and even though I was aware of the consequences I'd face if I kept missing school like this, it was so much easier to stop pushing so hard against the current and just step in the other direction.

My hands jammed into my pockets, already feeling an ambivalent sort of relief, I caught the bus back to the docks.


\chapter{Agitation 3.2}

I was pleasantly surprised to find that the bus line that ended at the old ferry put me only a fifteen or twenty minute walk away from the loft that Lisa, Alec and Bitch called home. I could be spending a fair bit of time there before I gathered enough information or earned enough trust from them to turn them in to the authorities, so the convenience was nice.

It was a nice day, if a bit windy. The air was crisp and cool, the sky was a brilliant and cloudless blue that was reflected in the ocean, and the sand of the beach sparkled in the light of the sun. Tourists were already crowding the railings or migrating to the beach, pinning down the corners of their beach blankets under picnic baskets and shopping bags. It was too cold to go in the water but the view was spectacular. I enjoyed it for a few moments before venturing into the crowd. I walked with my hands in my pockets, as much to protect the stuff in my pockets as keeping the worst of the chill out.

Living in Brockton Bay, you learned stuff like that. How to protect yourself, what to watch for. I knew that the Vietnamese teenagers who were leaning against the railing of the boardwalk were members of the ABB, even if they weren't wearing their gang colors, because the only Asian kids in Brockton Bay that had that much swagger were already part of Lung's gang. I knew the tattoo on the arm of the guy lifting boxes into the florist's van that read \quote{Erase, Extinguish, Eradicate} meant the guy was a white supremacist because it had the letter E repeated three times.

The man in the uniform who was talking to a shop owner wasn't a cop or security guard, but one of the enforcers the merchants of the Boardwalk hired to keep the undesirables from making trouble. They were why the Boardwalk didn't have beggars, addicts, or people wearing gang colors hanging around. If your presence offended or worried the tourists, they would step up to scare you off. If someone shoplifted or panhandled in the Boardwalk, they ran the risk that one or two enforcers would drag them behind one of the shops and teach them a lesson. Anything more serious than shoplifting or panhandling, well, there was always someone on duty in the floating base of the Protectorate Headquarters. Any of the store owners or employees could call the likes of Miss Militia, Armsmaster or Triumph in, given a minute. The tourism revenue the Boardwalk picked up earned a {\em lot} of goodwill from the government and government sponsored capes.

I headed off the boardwalk and into one of the alleys leading into the Docks. Glancing over my shoulder, I saw one of the uniformed enforcers staring at me. I wondered what he was thinking. Good kids didn't hang out in the Docks, and I doubted I looked the part of a guileless tourist.

The abandoned factories, warehouses and garages of the Docks all blended into one another very quickly. The colors of the building exteriors weren't different enough from one another to make buildings recognizable, and the people or piles of garbage that I had been unconsciously noting my previous visit had all shifted locations or been replaced. I found myself glad for the artistic graffiti and the row of weed-entangled power lines that I could use as landmarks. I did not want to get lost. Not here.

As I arrived at the foot of the huge factory with the Redmond Welding sign, I found myself wondering whether I should knock or just go on up. I didn't have to decide -- the door opened just a second after I'd come. It was Brian, and he looked as surprised to see me as I was to see him.

\quotation{Hey,} he said, \quotation{Lisa said you'd arrived. I thought you had school.}

It took me a few seconds to get my mental footing. Any demonstration or mention of Lisa's power kind of had a way of doing that to me, and that was on top of having a conversation sprung on me without a chance to prepare. \quotation{Changed my mind,} I said, lamely.

\quotation{Huh. Well, come on up.}

We headed upstairs. I saw Brian was wearing different clothes than what he had been wearing earlier in the morning. What he was wearing now bore a closer resemblance to his clothes from the day before -- a green sleeveless t-shirt and black slacks with a lightweight fabric, like yoga pants or something.

Alec was waiting, leaning against the back of a couch, as we entered the living room. He was wearing a t-shirt with some cartoon or video game character on it and basketball shorts. He stood straight as he noticed us.

\quotation{Alec and I were sparring,} Brian told me, \quotation{Lisa's on the phone in the kitchen. Rachel and her dogs are in her room. You can watch us, if you want, but no pressure. Feel free to use the TV, put on a DVD or play a video game.}

\quotation{Don't save over any of my files, dork,} Alec said. He'd started with the \quote{dork} thing last night. It wasn't exactly malicious, but it grated.

\quotation{My name is Taylor, not dork, and I wouldn't do that,} I told him. Turning to Brian, I said, \quotation{I'll watch, if it's cool.}

Brian smiled and nodded, while I moved to kneel on the couch and watch them over the back of it.

As it turned out, it was less of a \quote{sparring} session than an attempt on Brian's part to give a less than fully committed Alec some basic lessons on hand to hand fighting.

It was one-sided, and not just because Alec wasn't trying very hard. Alec was a very average fifteen year old guy in that he had little muscle worth speaking about. Brian, by contrast, was fit. He wasn't big in the sense of a bodybuilder or someone who exercised just to pack on muscle like you saw with some of the people just out of prison. It was a little more streamlined than that. You could see the raised line of a vein running down his bicep, and the definition of his chest showed through his shirt.

Besides the difference in raw physical power, there was also the age and height gap. Alec was two or three years younger and nearly a foot shorter. That meant Brian had more reach -- and I'm not just referring to the length of his arms. When he stepped forward or backward, he moved further. He covered more ground, which put Alec on the defensive, and since Brian was stronger, that put Alec in a bad position.

Brian stood without much of a fighting stance, hands at his sides, bouncing just a little where he stood. Twice in a row, I watched Alec swing a punch, only for Brian to lean out of the way. The second time Alec's arm flew by, Brian leaned in and jabbed Alec in the center of his chest. It didn't look like much of a punch, but Alec still sort of woofed out a breath and stepped back.

\quotation{I keep telling you,} Brian said, \quotation{You're throwing punches like you'd throw a baseball. Don't bring your arm so far back before you punch. You're just broadcasting what you're about to do and it doesn't add enough power to the hit to be worth that.}

\quotation{What am I supposed to do, then?}

\quotation{Look at how I'm standing. Arms up, bent, then I just extend my arm, wrist straight. Fast enough that whoever I'm hitting generally can't step out of the way, so they've got to either take it or block it.}

\quotation{But you weren't standing like that ten seconds ago when I was punching you,} Alec complained.

\quotation{I left an opening to see if you would take advantage of it,} Brian replied.

\quotation{And I didn't,} Alec noted with a sigh.

Brian shook his head.

\quotation{Well fuck this then,} Alec said, \quotation{If you're going to go easy on me and still kick my ass, I don't see the point.}

\quotation{You should learn how to fight,} Brian said.

\quotation{I'll do like I have been and bring my taser,} was Alec's response, \quotation{one poke and they're out cold. Better than any punch.}

\quotation{And if the taser breaks or you lose it?} Brian asked. He needn't have bothered. Alec was already sitting himself down in front of the TV, remote in one hand and game controller in the other. Brian's disappointment was palpable.

\quotation{Mind giving me a few quick and dirty pointers?} I asked.

Alec sniggered, Beavis and Butthead style.

\quotation{Grow up, Alec,} Brian said, \quotation{If you want to quit, fine, but don't be a dick.} He turned to me and flashed that boyish smile. Then we started.

I knew he was going easy on me, but he was still a damn tough teacher.

\quotation{Make two fists. No, don't wrap your fingers over your thumbs. You'll do more damage to your hands than you will to the person you're hitting, if you do that. That's better. Now jab at me, okay?}

I tried to emulate what he'd been describing to Alec. Arms up, bent, and extending my fist with a snap. He caught my right hand in his left.

\quotation{Okay, now you're going to do two things different. Step into the jab so you've got your body's momentum behind the hit, on top of your arm's power. Second, I want your left arm up as you're jabbing with your right, and vice versa. If I see the chance, I'm going to pop you one on the shoulder or ribs, so be ready to fend me off.}

I winced at the idea, but I played along. I jabbed, he stepped away, and he jabbed me in the shoulder. He didn't hit as hard as he could have -- I think he only hit as hard as it took to make it hurt and drive the lesson home, but I suddenly felt a stab of sympathy for Alec.

Things continued in that vein. Brian didn't stay on one topic for long. When I started struggling with something, he shifted gears to another area that complemented or built on what I was having problems with. When I failed for the fifth time to fend off his retaliatory jabs at my shoulders and ribs, he started talking about posture.

\quotation{Rest your weight on the balls of your feet.}

I tried it, then told him, \quotation{I feel like I'm going to tip over backwards if you hit me.}

He bent down to check, and I lifted my toes two or three inches off the ground to demonstrate how I had my weight balanced on my heels.

“No, Taylor. The balls of your feet. He raised his bare foot and pointed at the padded part between his toes and the bridge of his foot.

\quotation{How is that a ball?} I asked, raising my own foot to point at the vaguely spherical part of the foot where the ankle met the ground, \quotation{this is the only part that looks ball-like.}

\quotation{You guys are so lame,} Alec chimed in, without turning around. Brian swatted him in the back of the head.

We moved on from posture, Brian's recommendations on balancing did help, to self-defense again. From there, we changed topics to the mental side of things, both for me and my opponent.

\quotation{So I throw a punch like I'm aiming to put my fist through them?} I confirmed.

\quotation{Right,} Brian said, \quotation{Instead of just trying to make contact with the point where your hand meets their body.}

\quotation{What about when they're attacking me?}

\quotation{Best bet? Don't give them a chance. Stay aggressive and keep them on their heels. If neither of you have formal training, then that's going to give you the best odds. They won't be able to turn the tables on you unless you make a mistake or they can guess what you're going to do as you do it. Which is why you mix it up. Rights, lefts, punches, jabs, elbow, knee, kicks and if you're bigger and stronger than them, you can try tackling them to the ground. With all of that, you stay on them until they aren't in a position to fight back.}

\quotation{Are you formally trained in anything?} I asked. I suspected he was, since the only other way for him to know as much as he was demonstrating was to have actually been in a good number of fights, and I wasn't thinking that he seemed the type to fight without reason.

\quotation{Ehhh,} he hedged, “Some. My dad was a boxer when he was in the service, and he taught me some when I was little. I moved on to other stuff on my own -- Karate, Tae Kwon Do, Krav Maga -- but nothing really held my interest. I only took a few weeks or a month of classes for each. I know enough and keep in shape, which is enough to hold my own against anyone who isn't a black belt in whatever, which is the important thing, I think. Keeping up with the more serious martial artists is a full time job, and you're {\em still} going to run into people who are better than you, so I don't see the point in stressing too much over it.”

I nodded.

We moved on to key areas to attack.

Brian pointed to the body parts in question as he explained, \quotation{Eyes, nose, temple, chin and throat are the areas above the shoulder. Teeth or ears if you can hit hard enough. I can, you can't.}

\quotation{Sure,} I said. I wasn't offended by his bluntness. He was stronger than me, so he had more options. Tip toeing around it didn't do either of us any favors.

\quotation{Below the shoulders, diaphragm, kidney, groin, knee, bridge of the foot, toes. Elbow is a good one if you can do anything with it,} he took my wrist in his left hand and my shoulder in his right, extending my arm straight as he brought his knee up to gently tap the outside of my elbow. I could see how he would have screwed up or broken my arm if he'd done it full strength. He went on, \quotation{But in my experience, it doesn't come up often enough to worry about.}

It was a little disquieting to hear Brian methodically describing how to break a human being. I saw him as a nice guy, if I ignored his career choice.

Not entirely by accident, I changed the subject, \quotation{I was thinking about investing in a weapon for hand to hand. When I was fighting Lung, fists were no good and I found myself really wanting a knife or a baton or something. Don't know if they would have been any good against his armor, but you know\ldots{}} I trailed off.

Brian nodded, \quotation{Makes sense. You don't have a lot in the way of upper body strength, no offense.}

\quotation{None taken. I tried to get something like a push-up routine going, but I got sick of it fast. At least with running, there's that sense of going places, you get the scenery.}

\quotation{Push-ups get repetitive, yeah. Well, the boss is good about supplying us with gear. Lisa's the one who talks to him, she's talking to him right now, in fact. Put in a word with her if you want something like that. It's untraceable too, so the good guys aren't going to be tracing any serial numbers or whatever from your weapon back to your purchase.}

The fact that Lisa was talking to their boss made me very curious, all of a sudden. That said, I couldn't really traipse in to eavesdrop without being suspicious. Instead, since Lisa was out of earshot, I thought I'd seize the opportunity to ask, \quotation{So who is this boss of ours?}

Brian and Alec exchanged a look. When they didn't immediately say anything, I wondered if I'd pushed it too far. Had I been too nosy?

\quotation{Figured you'd ask,} Brian said, \quotation{Thing is, we don't know.}

\quotation{What?} I asked, \quotation{We have an anonymous sponsor?}

\quotation{It's really fucking weird, yeah,} Alec said, then he hammered a button on the game controller, \quotation{Boom! Triple headshot!}

\quotation{Alec, stay focused,} Brian sighed the words, with a tone suggesting he didn't expect to be listened to.

Alec bobbed his head in a nod, his eyes not leaving the television, before adding, \quotation{It's weird but it's basically free money, a good team, contacts, access to everything we need for stuff, and pretty much no drawbacks.}

“{\em Lisa} knows, I think,” Brian grumbled, \quotation{But she says that when she joined the Undersiders, she made a deal that she was going to keep quiet on the subject. I'm not sure if that means she knows who he is or if it's just to keep her mouth shut if her power tells her.}

\quotation{So let me get this straight,} I said, \quotation{This guy gathers you all together, offers you a salary and what? Doesn't ask for anything in return?}

Brian shrugged, \quotation{He asks us to do jobs, but most of the time it's stuff we'd do anyways, and if we say no, he doesn't make an issue of it.}

\quotation{What kind of jobs does he ask us to do?} I asked.

Lisa's voice just behind me startled me, \quotation{This. Pull up your socks, boys and girl, because we're robbing a bank.}


\chapter{Agitation 3.3}

\quotation{No,} Brian intoned, \quotation{Such a bad idea.}

Lisa still had the phone in her hand. Bitch had arrived just behind her, and stood in stark contrast to Lisa's jeans, sweater and tight ponytail, with an army jacket, and virtually no attention paid to her hair. The littlest of the dogs, the one-eyed, one eared terrier, trailed after her.

\quotation{Come on,} Lisa wheedled, \quotation{It's a rite of passage for dastardly criminals like us.}

\quotation{Robbing a bank is moronic. We've been over this,} Brian closed his eyes and pinched the bridge of his nose, \quotation{You know what the average haul is for hitting a bank?}

Lisa paused, \quotation{Twenty thou?}

“Exactly. It's not millions like you see people getting away with in the movies. Banks don't keep a lot of loose cash on hand, so we'd be pulling in {\em less} than we would for most other jobs. Account for cost and the fact that this is Brockton fucking Bay, where banks have a little more reason to keep the amount of cash in their vaults to a minimum, and we'd be bringing in twelve to sixteen thou. Divide five ways and it's what, two or three thousand bucks each?”

\quotation{I could do with an extra three thousand dollars to spend,} Alec said, putting down his game controller and shifting his position on the couch to follow the conversation better.

\quotation{On what?} Brian asked. When Alec shrugged, Brian sighed and explained, \quotation{It's a horrible payoff for the amount of risk involved. There's three big superhero teams in this city. Figure there's another dozen heroes that fly solo, and we're almost guaranteed to get into a fight.}

\quotation{So?} Bitch spoke for the first time, “We win fights. We won before we had {\em her}.” She raised her chin in my direction as she said that last word.

\quotation{We won because we picked our battles. We wouldn't have that option if we were cooped up in the bank and waiting for them to come to us, letting them decide how and where the fight happened.}

Lisa nodded and smiled as he spoke. I thought for a second that she was going to say something, but she didn't.

Brian continued, getting pretty passionate as he ranted, \quotation{We won't be able to slip away like we have when things got a little out of control in the past. Can't avoid the fight if we want to get away with anything worth taking. The bank is going to have layers of protection. Iron bars, vault doors, whatever. Even with your power, Lise, there's a limit to how fast we can get through those. Add the time we have to spend managing hostages and making a safe exit, and I pretty much guarantee that there will be time for a cape to get wind of the robbery and slow us down even more.}

Alec said, \quotation{I kind of want to do it anyways. Hitting a bank gets you on the front page. It's huge for our rep.}

\quotation{The runt is right,} Bitch said.

Brian grumbled, \quotation{Not fucking up is better for our reputation in the long run.} His deeper voice was really good for grumbling.

Alec looked at me, \quotation{What do you think?}

I'd almost forgotten I was a part of the discussion. The last thing I wanted was to rob a bank. Hostages could get hurt. The fact that it would potentially put me on the front page of the paper wasn't a high point, either, if I ever wanted to drop the supervillain ruse and become a hero in good standing. I ventured, \quotation{I think Brian makes a good case. It seems reckless.}

Bitch snorted. I think I saw Alec roll his eyes.

Lisa leaned forward, \quotation{He does make good points, but I have better ones. Hear me out?} The rest of us turned our attention to her, though Brian had a frown that made it seem like it would take a lot to convince him.

“Ok, so Brian said similar stuff before, before we hit that casino a few weeks ago. So I was kind of expecting this. But it's not as bad as it sounds. The boss wants us to do a job at a very specific time. I got the sense he was willing to offer a fair bit if we went the extra mile, and I negotiated a pretty good deal.

\quotation{The bank robbery was my idea, and he liked it. According to him, the Protectorate is busy with an event on Thursday, just outside of town. That's part of the reason the timing is so important. If we act then, there's almost no chance we'll have to deal with them. If we hit the Bay Central, downtown-}

\quotation{That's the biggest bank in Brockton Bay,} I interrupted her, half-disbelieving.

\quotation{So everything I said about them having security and being careful is doubly true,} Brian added.

“{\em If we hit the Bay Central, downtown},” Lisa repeated herself, ignoring us, \quotation{Then we're hitting a location just a mile away from Arcadia High, where most of the Wards go to school. Given jurisdictions, New Wave won't be able to jump on us without stepping on the Wards' toes, which pretty much guarantees we go up against the team of junior superheroes. With me so far?}

We all nodded or murmured agreement.

“Figure that's happening in the middle of the school day, and they won't {\em all} be able to slip away to stop a robbery without drawing attention. People know the Wards are attending Arcadia, they just don't know who they are. So everyone's constantly watching for that. Since they can't have all six or seven of the same kids disappear from class every time the Wards go off to foil a crime without giving away the show, chances are good that we'd go up against a couple of their strongest members, or one of the strongest with a group of the ones with less amazing powers. We can beat them.”

\quotation{Okay,} Brian begrudged, \quotation{I'll accept that we'd probably do alright in those circumstances, but-}

Lisa interrupted him, “I also got the boss to agree to match us two for one on the haul. We bring in fifteen grand, he pays us thirty. {\em Or} he gives us enough money to bring our total up to twenty five, whichever is more in the end. So we could walk away with two thousand dollars and he'd pay us twenty three thou. So as long as we don't wind up in jail, we're guaranteed five thousand dollars apiece, bare minimum.”

Brian's eyes widened, \quotation{That's insane. Why would he do that?}

“{\em And},” Lisa grinned, \quotation{He'll cover all our costs, just this once. Equipment, information, bribes if we want 'em.}

\quotation{Why?} I echoed Brian's earlier question, disbelieving. Lisa was throwing around sums of money that I couldn't even wrap my head around. I had never even had more than five hundred dollars in my bank account.

\quotation{Because he's sponsoring us and it stands to reason he doesn't want to fund a team of nobodies. We manage this, we won't be nobodies. That, and he really wants us to do a job at that particular time.}

There was a few moments of silence as everyone considered the deal. I was frantically trying to think of a way to try to convince these guys it was a bad idea. A bank robbery could get me arrested. Worse, it could lead to me or a bystander getting hurt or killed.

Brian beat me to it, \quotation{The risk to reward still isn't great. Five grand each for hitting what may well be the most fortified location in Brockton Bay and an almost guaranteed confrontation with the Wards?}

\quotation{Second most fortified location,} Lisa countered, \quotation{The Protectorate Headquarters is the first.}

\quotation{Fair point,} Brian said, \quotation{But my argument stands.}

\quotation{It'll be more than five grand for each of us, I guarantee you,} Lisa told him, \quotation{It's the biggest bank in Brockton Bay. It's also the hub of cash distribution for the entire county. Said cash gets transferred in and out by armored cars on a regular schedule-}

\quotation{So why don't we hit one of the cars?} Alec asked.

“They have ride-alongs or aerial cover from various members of the Wards and the Protectorate, so we'd be caught in a fight with another cape from minute one. Same problems that Brian's talking about, as far as getting caught up in a fight, difficulty accessing the money before shit goes down, yadda yadda. {\em Anyways}, the Brockton Bay Central has cars coming in twice a week, and leaving four times a week. We hit on a Thursday just after noon, and it should be the best day and time for the sheer size of the take. Only way we're getting away with less than thirty thousand is if we fuck up. With what the boss is offering, that's ninety thou.”

She folded her arms.

Brian sighed, long and loud, \quotation{Well, you got me, I guess. It sounds good.}

Lisa turned to Alec. There wasn't any resistance to be found there. He just said, \quotation{Fuck yeah, I'm in.}

Bitch didn't need convincing any more than Alec had. She nodded once and then turned her attention to the scarred little dog.

Then everyone looked at me.

\quotation{What would I be doing?} I asked, nervously, hoping to stall or find holes in the plan that I could use to argue against it.

So Lisa outlined a general plan. Brian made suggestions, good ones, and the plan was adjusted accordingly. I realized with a growing disappointment and a knot of anxiety in my gut that it was almost inevitably going to happen.

Arguing against the bank robbery at this point would hurt my undercover operation more than it helped anyone. With that in mind, I began offering suggestions that -- I hoped -- would minimize the possibility of disaster. The way I saw it, if I helped things go smoothly, it would help my scheme to get info on the Undersiders and their boss. It would minimize the chance that someone would panic or be reckless and get a civilian hurt. I think I would feel worse if that happened than I would about going to jail.

The discussion went on for a while. At one point, Lisa got her laptop, and we debated entrance and exit strategies while she sketched out a map of the bank layout. It was uncanny, seeing her power at work. She copied a satellite image of the bank from a web search into a paint program, then drew over it with thick bold lines to show how the rooms were laid out. With another search and a single picture of the bank manager standing in front of his desk, she was able to mark out where the manager's desk was. That wouldn't have been too amazing, but without pausing, she then went on to mark where the tellers were, as well as the vaults, the vault doors and the enclosed room that held the safe deposit boxes. She noted where the fuse box and air conditioning vents were, but we decided we wouldn't mess with either of those. That stuff was cool in the movies, but it didn't do much good in real life. Besides, this was a robbery, not a heist.

While we worked, Alec got restless and went to make an early lunch. Of the four of us, I got the impression he had the least to contribute, at least strategically, and that he knew it. I wasn't sure if he just didn't have a very tactical mindset or if he just didn't care that much about the planning stage of things. My assumptions led to the latter, as he seemed more willing to go with the flow than Brian or Lisa.

He brought us a plate of pizza pockets along with assorted sodas, and we ate as we wrapped up the plan.

\quotation{Alright,} Brian said, as Lisa shut her laptop, \quotation{I think we have a general idea of what we're doing. We know how we get in, we know who does what when we're inside, and we know how we want to get out. Keeping in mind that no plan survives contact with the enemy, I think the odds are still pretty good.}

\quotation{So, the enemy,} I said, resisting the urge to wince at the realization that I would be up against good guys, \quotation{My only experience fighting in costume\ldots{} or even just fighting, is against Lung, and that didn't go well.}

\quotation{Don't sell yourself short,} Brian said, \quotation{You did better than most.}

\quotation{I'll rephrase,} I said, “It could have gone {\em better.} We're going up against the Wards and they aren't pushovers.”

Brian nodded, \quotation{True. Let's talk strategy and weaknesses. You know who the Wards are?}

I shrugged, \quotation{I've researched them. I've seen them on TV. That doesn't mean I know the important stuff.}

\quotation{Sure,} he said, \quotation{So let's go down the list. Team leader: Aegis. You'd think he has the standard Alexandria package, flight, super strength, invincibility, but that isn't exactly right. He does fly, but the other two powers work differently than you'd expect. See, he isn't invincible\ldots{} he just doesn't have any weak points. His entire biology is filled with so many redundancies and reinforcements that you just can't put him down. Throw sand in his eyes and he can still see by sensing the light on his skin. Cut his throat and it doesn't bleed any more than the back of his hand would. The guy's had an arm cut off and it was attached and working fine the next day. Stab him through the heart and another organ takes over the necessary functions.}

\quotation{Not that we're stabbing anyone through the heart?} I made it a hopeful half-question, half-statement.

\quotation{No. Well, it wouldn't be a bad idea to stab Aegis through the heart just to slow him down. If you did it with something big enough. The guy's like a zombie, he gets back up within seconds of you beating him down, keeps coming at you until you're too tired to fight back or you make a mistake.}

\quotation{And he's super strong?} I asked.

Brian shook his head, \quotation{Lisa, want to field this one?}

She did. \quotation{Aegis isn't strong, but he can abuse his body in ways that makes it seem like he is. He can throw punches hard enough that they'd break his hand, mangle his joints and tear his muscles, and his body just takes it. He has no reason to hold back, and he doesn't need to waste any time protecting himself from you. He can also draw on adrenaline\ldots{} you've heard stories like how little old grandmothers lifted cars off the ground to save their grandkids?}

I nodded.

\quotation{That's adrenaline at work, and Aegis can do that for hours at a stretch. His body doesn't run out of steam, he doesn't get tired, he doesn't exhaust his reserves of adrenaline. He just keeps going.}

\quotation{So how do you stop him?} I asked.

\quotation{You don't, really,} Brian said, \quotation{Best bet is to keep him occupied, keep him sufficiently distracted or stick him somewhere he can't escape. Trap him in a dumpster and throw it in the river, you can get a few minutes of relief. Which is all harder than it sounds. He's the team captain, and he isn't stupid. Rachel? Sic your dogs on him. A two ton canine or two should keep him out of our hair until we're ready to run.}

\quotation{I don't need to hold back?} Bitch asked, her eyebrow quirked.

\quotation{For once, no. Go nuts. Just, you know, don't kill him. Alec? You're the backup there. Keep an eye on Aegis, see if you can't use your power to throw him off. Buy enough time for a dog to get its jaws on him and he's probably out of action.}

\quotation{Sure,} Alec said.

Brian extended two fingers and tapped the second, \quotation{Number two. Clockblocker. Let it be known, I fucking hate people who mess with time.}

\quotation{He stops time, if I remember right?} I inquired, as much to stay in the conversation as to get the clarification.

\quotation{More specific than that,} Brian said, \quotation{He can stop time for whatever he touches. The person or object he touches is basically put on \quote{pause} for anywhere from thirty seconds to ten minutes. Only good thing is that he doesn't control or know how long it's going to last. But if he gets his hands on you, you're out of action. He'll either stand next to you and wait until you start moving, then touch you again, or he'll just tie you up in chains and handcuffs so that when his power wears off, you're already in custody.}

\quotation{Long story short, he touches you, you're boned,} Alec said.

\quotation{The upside is that whoever he touches is also untouchable. Can't be hurt, can't be moved. Period. He uses that defensively, and he can do stuff like throw paper or cloth in the air and freeze it in time, making an unbreakable shield. You don't want to run into something that's frozen. A car that drove into the side of a piece of paper that Clockblocker had touched would be cut in two before it budged the paper.}

\quotation{Noted,} I said.

Brian continued, “The third heavy hitter on the Wards is Vista. You know that myth about how the capes that get their powers young are exponentially more powerful? Vista's one of the kids who keeps the myth alive. Clockblocker is sort of a one trick pony, his trick involves screwing with one of the key forces of our universe, but it's just one thing. Vista {\em also} messes with physics on a fundamental level, but she's versatile.

\quotation{Twelve years old, and she has the power to reshape space. She can stretch a building like taffy, so it's twice as tall, or squeeze two sidewalks closer together so she can cross the street with a single step.}

\quotation{Her weakness,} Lisa added, \quotation{Is the Manton effect.} She turned her full attention to me, \quotation{You know what that is?}

\quotation{I've heard it mentioned, but I don't know the details.}

“Wherever our powers come from, they also came with some limitations. For most of us, there's a restriction about using our powers on living things. The reach of powers generally stops at the outside of a person or animal's body. There's exceptions for the people with powers that {\em only} work on living things, like you, Alec and Rachel. But the long and short of it is that the Manton effect is why most telekinetics can't just reach into your chest and crush your heart. Most people who can create forcefields can't create one through the middle of your body and cut you in two.”

\quotation{Narwhal can,} Alec cut in.

\quotation{I said most,} Lisa said, \quotation{Why these restrictions exist is a question nearly as big as where we got our powers in the first place. The capes that can get around the Manton effect are among the strongest of us.}

I nodded, slowly. I wondered if that had something to do with why Lung didn't burn himself, but I didn't want to get further off topic, \quotation{And Vista?}

“Vista can stretch and compress space. She can also do funny things with gravity. Thing is, the Manton effect keeps her from stretching or compressing {\em you}. It also makes altering an area a lot harder for her if there's more people in that space. So if all of us are in one room, chances are she won't be able to affect the whole room.”

“{\em But},” Brian added, wiping a string of cheese from the corner of his lip, “Every time we've run into her, she's been faster and overall more powerful with her power, and she's had new tricks. Every second she's on the battlefield is a second things become harder for us. We take her down sooner than later. Aegis, Clockblocker, Vista. Those are the ones we're most likely to run into, and whoever else winds up coming, they're the ones we have to deal with, or we're fucked.

\quotation{Let's quickly go through the rest. Kid Win.}

\quotation{Tinker,} Lisa said, \quotation{Flying skateboard, laser pistols, high tech visor are staples for him. Expect something new, depending on what he's come up in his workshop. He's mobile but not that threatening.}

\quotation{Triumph?} Brian said.

\quotation{He turned eighteen and graduated to the Protectorate. Don't have to worry about him,} Lisa said.

\quotation{Gallant.}

\quotation{Glory Girl's on and off boyfriend, he pretends to be a Tinker in the same vein as Kid Win, but I think he just runs around in secondhand armor with a fresh paint job. His thing is these blasts of light. Getting hit by one feels like a punch in the gut, but the blasts also mess with your feelings. Make you sad, make you scared, ashamed, giddy, whatever. Not that bad unless you get hit by a bunch in a row. Don't.}

\quotation{That just leaves Shadow Stalker. Bloodthirsty bitch,} Brian scowled.

Alec explained to me, \quotation{She's got it in her head that Brian is her nemesis. You know, her number one enemy, her dark opposite. She's been going after him every chance she gets.}

\quotation{She was a solo hero,} Tattletale said, \quotation{Vigilante of the night, until she went too far and nearly killed someone, nailing him to a wall with one of her crossbows. The local heroes were called in, she got arrested, and made some sort of deal. Now she's a probationary member of the Wards, with the condition that she uses tranquilizer bolts and nonlethal ammo for her crossbow.}

\quotation{Which she isn't,} Brian growled, \quotation{At least, not when she comes after me. That arrow she shot through my side had a fucking arrowhead on it.}

Tattletale shook her head, \quotation{Her powers and Brian's sort of have a weird interaction with one another. Shadow Stalker can sort of transform. She becomes extremely lightweight, can pass through glass and thin walls and she's nearly invisible. Only thing is, while she and the stuff she carries are all wispy in her transformed state, the stuff she shoots with her crossbow only stays that way for a half second. Then the effect wears off and it's a regular arrow flying towards you. So she can leap between rooftops, almost impossible to see, hard to even touch, and all the while she's shooting very real arrows at you.}

\quotation{So what do you do?} I asked.

\quotation{Her power doesn't work well while she's inside Brian's darkness, for whatever reason. She isn't as fast or agile, he can see her better, and she can't see him in the darkness,} Tattletale told me, \quotation{So it becomes something of a very intense game of tag, with one very fast person that's essentially blind and deaf but carrying lethal weapons, while Brian, the other, is trying to take her out without getting shot.}

\quotation{Let's avoid that,} Brian said, “It's too time consuming and she may want to use that kind of scenario to delay us. Just don't get shot, and if you see her or see the opportunity, inform the team and do your best to take her down without losing sight of a priority target.

\quotation{So that's the plan, then?} I said, \quotation{So many maybes.}

\quotation{That's the way these things go, Taylor,} Brian said, his tone a bit terse, \quotation{I think we've done a pretty good job of covering all the bases.}

\quotation{Oh, I didn't mean to sound like I was criticizing your plan-} I said.

\quotation{Our plan,} Brian interrupted.

I didn't want to think of it that way. Instead, I said, \quotation{I'm a touch nervous, is all.}

\quotation{You don't have to come,} Bitch said, her tone a touch too casual.

\quotation{In all seriousness,} Brian told me, \quotation{If you're having second thoughts\ldots{}}

\quotation{I am,} I admitted, \quotation{as well as third thoughts, fourth thoughts, and so on. But I'm not going to let that stop me. I'm coming with.}

\quotation{Good,} Brian replied, \quotation{Then we've got the rest of today and tomorrow to prepare. Taylor? You can meet me on your run first thing. I'll have a cell phone for you. You can text Lisa with anything you think you'll need, like those weapons you were talking about. Look up models and brands ahead of time if you want something specific.}

\quotation{What's her number?} I asked.

\quotation{I'll put it in the phone before I give it to you. Lisa? You confirm the job with the boss, talk to him about the other stuff.}

\quotation{Got it.}

\quotation{So unless there's anything else, I think we just planned a bank robbery before noon,} Lisa said with a grin. I looked at the digital clock displayed under the TV. Sure enough, it was half past eleven.

I couldn't help but wonder if that was a good thing.


\chapter{Agitation 3.4}

\quotation{I'll be there. Yes-} I saw a light in the living room window and put my hand over the lower half of my cell phone while I briefly investigated. Damn, my dad was home. I put the phone to my ear, \quotation{I'm sorry, I've got to run. No. No. Look-}

As I heard the front door open, I snapped the phone shut and jammed it into my pocket. I'd apologize for hanging up later. I definitely didn't want my dad to see the phone. I didn't think he would stop me from owning one, but ever since my mom's death, cell phones had carried strong negative connotations. That, and I'd have to explain where I got it and how I'd paid for it.

Brian had given me three identical cell phones -- all disposables -- first thing in the morning, and I'd decided to go with him to the loft rather than head to school. The way I figured it, I didn't have much of a chance of focusing on classes with Thursday's bank robbery occupying my attention on top of the stress of just being there and waiting for the other shoe to drop as far as my skipped classes. Besides, I rationalized, it didn't make a lot of sense to go if I knew I would be skipping again to go rob the bank. I'd promised myself I would go the day after tomorrow. Face the music.

I'd spent the day with the group. Rachel had been out of the apartment, the others weren't specific on why and I wasn't interested enough to risk looking too curious by asking. So it had just been me, Brian, Alec and Lisa. We'd hammered out the fine details of the robbery and I had decided what weapons I wanted Lisa to ask the boss for. I had elected for both a combat knife and a telescoping police baton. The knife would serve for emergencies and those people who were just too tough to hurt with the baton. The baton, twenty one inches long when fully extended, was for more general use, offering more clout than I'd otherwise get with my fists. Lisa had promised I would have them for tomorrow.

After that, we kind of avoided the subject of the robbery, by some unspoken agreement. It wouldn't do to overthink it or risk getting too nervous. Either way, I had felt a need to burn some nervous energy, so I had helped clear out the storage closet around lunchtime, with Lisa and Brian's help. We'd sorted out the stuff, found a place for it all, and set up the room with odds and ends they had lying around. The stuff included an extendable clothes rack, a dresser, an inflatable mattress and a bedside table with a lamp attached. It was enough space for me to keep some toiletries, a spare change of clothes or two, my costume and my equipment. Lisa spent a lot of time talking about what I could do to make the space my own, what I could buy, how I could decorate, but I was happy enough with what we had there. I kind of liked that it was a bit spartan, because it sort of fit with how I didn't plan to be around that long while still feeling weirdly appreciative at being accepted as a part of the group.

Having tired ourselves out, we'd all collapsed on the couches and watched some of Alec's movies from Earth-Aleph, the alternate Earth that our Earth had been communicating with since Professor Haywire tore a hole between realities. Media was one of the few things that could be traded back and forth through the hole. Long story short, you could get books, movies and DVDs of TV shows from the other world, if you were willing to accept the price tag. The benefit? I got to spend the afternoon seeing how the other universe had handled episodes one and two of the Star Wars films.

Fact: they were still pretty disappointing.

By the time my dad got in, I had pork chops defrosted, dusted with lemon and pepper and sitting in a frying pan, with vegetables in the microwave. Cooking was sort of something you started doing when you had only one parent, unless you really, really liked takeout.

\quotation{Heya,} my dad greeted me, \quotation{Smells good.}

\quotation{I started dinner a bit early because I have somewhere I want to be, tonight, If that's cool?}

He tried to hide it, but I could see a bit of disappointment. \quotation{Of course,} he said, \quotation{Your new friends?}

I nodded.

\quotation{Let me get changed and then I'll ask you all about them,} he promised as he headed upstairs.

Great. I hadn't had to answer these questions last night because my dad had been working late. My mind started racing to anticipate questions and come up with plausible details. Should I use their real names? Or at least, the names they had given me? I wasn't sure if that would be a breach of trust. I decided to use their real names for much the same reason I'd decided to use my own with them. It just prevented disasters if my dad ever happened to meet them, which was a terrifying thought, or if they called for me.

I didn't need to worry about my dad hearing about four kids being arrested, all of whom had the same name as my \quote{friends}, since most or all of them were minors and their names would be kept from the media under the law. I was also under the impression that the courts didn't always unmask capes when they arrested them. I wasn't entirely sure what was up with that. It seemed like something to ask Lisa about.

By the time my dad had come back downstairs, I'd resolved to try and keep my lies as close to the truth as possible. It would be easiest to keep everything straight that way. That, and I hated lying to my dad.

My dad had changed out of his dress shirt and khakis, into a t-shirt and jeans. He mussed up my hair and then took over the last bit of the cooking. I sat down at the table so I could talk to him.

\quotation{So what's going on?} he asked.

I shrugged. I hated feeling this tense around my dad. He'd never bugged me about the bullying, so I'd always been able to come home and sort of let my guard drop. I couldn't do that now, because I was just waiting for the other shoe to drop as far as my skipped classes, and my new \quote{friends} brought a whole mess of secrets and lies into the mix as well. I felt like I was on the verge of a terminal breach of trust. One mistake or a single concerned phone call from the school, and my dad would probably flip, and things wouldn't be the same between us for a long time.

\quotation{Are you going to tell me their names?} he asked. He set the food on plates and brought it to the table.

\quotation{Brian, Lisa, Alec, Rachel,} I confessed, \quotation{They're alright. Get along with most of them.}

\quotation{Where did you meet them? School?}

I shook my head, \quotation{I wanted to get away from school for a bit, so I caught a bus downtown to catch a bit of a break. I ran into them at the library.} Partial truths. You couldn't really catch a bus downtown and back during the lunch break -- I'd tried, when I was avoiding the trio -- but I doubted my dad would research that. I did sort of cross paths with the Undersiders at the library, though.

\quotation{They go to the library at lunch? What are they like?}

\quotation{Brian's pretty cool. He's the one I've talked to the most.}

\quotation{A boy, eh?} My dad wiggled his eyebrows at me.

\quotation{Dad, stop! It's not like that,} I protested. I doubted Brian had the slightest interest in me, not least because I was two or three years younger than him. Besides, well, I was me. I opted not to mention the age difference to my dad.

Changing the subject, I said, \quotation{Lisa's alright too. Really smart, though I haven't talked to her all that much. It's nice being able to hang out with another girl again, even if she's pretty different from me.}

\quotation{If she's smart, she can't be that much different from you.}

I could've kicked myself. I couldn't explain she was a bad guy, while I was an aspiring superhero, or exactly how she was \quote{smart}. I'd talked myself into a minor corner where I didn't have an answer ready, and I needed to avoid doing that. Fumbling for an answer, I said, \quotation{She's only a year older than me, and she's graduated high school already.} That was the truth. She cheated, but she did technically graduate.

My dad smiled, \quotation{Impressive. Tell me they're all excellent students that can serve as good role models for you.}

I could have choked. Good role models? Them? I kept my composure and limited myself to a little smile and a shake of the head, \quotation{Sorry.}

\quotation{Alas. What about the others?}

\quotation{Alec is the youngest, I think. Kinda hard to connect with. He's an amazing artist, from what I've seen, but I don't really see him draw. It seems kind of hard to get him interested or involved in anything. He always looks bored.} As I said the words aloud, I realized they weren't exactly true. The two times I'd seen Alec react to anything had been when he'd played his little prank on Brian, tripping him, and after Bitch and I had been fighting. A streak of schadenfreude to his personality, maybe.

\quotation{And the last one? Rita? Rachel?}

\quotation{Yeah, Rachel. I don't get along with her. I don't like her.}

My dad nodded, but didn't say anything. I was halfway expecting the typical parental line of \quote{maybe if you try to show interest in things she likes} or some other inane advice. My dad didn't pull that on me, he just took another bite of pork chop.

I elaborated a bit, to fill the silence, \quotation{She wants things her way, and when she doesn't get that, she gets mean. I dunno. I get enough of that at school, you know?}

\quotation{I know,} my dad said. It was a good lead-in for him to question me about what was going on at school, but he didn't take it. He stayed quiet.

I felt immensely grateful, right then. My dad was respecting the boundaries I'd set, not pushing, not digging for more. It made this conversation so much easier that it might otherwise have been, and I knew it couldn't be that easy for him.

I felt like I owed him something for that. Sighing, I admitted, \quotation{Like, at school. The, uh, the people who're giving me a hard time? They sort of ganged up on me on Monday. Just, you know, taking turns insulting me. It's why I needed to get away and went downtown.} I felt embarrassed, saying it, because it was humiliating enough to live through without having to recap it, and because it felt so disconnected from the rest of the conversation. But if I didn't say it right then, I don't think I would've been able to.

My dad sort of went still. I could see him compose himself and choose his words before he asked, \quotation{Not to diminish how much it sucks to get put down like that, but they didn't do anything else?}

I raised my eyebrows in question as I chewed. They had, kind of, but I couldn't really say \quote{They used Mom's death to fuck with my head} without having to explain the Emma thing.

\quotation{Anything like what happened in January?} he asked.

I lowered my eyes to my plate, then shook my head. After a few moments I said, \quotation{No. January was a one time thing. They've pulled smaller \quote{pranks} since then, hassled me, but no repeat performances on that front.} I made air quotes with my fingers as I said \quote{pranks}.

\quotation{Okay,} my dad said, quietly, \quotation{That's a relief to know.}

I didn't feel like sharing any more. You'd think I would feel better, after opening up, but I didn't. I felt frustrated, angry, awkward. It was a reminder that I couldn't have a real conversation with my dad like I used to be able to. More than anything, I felt guilty. Part of the guilt was because I'd apparently let my dad think that every time I was bullied, it was like it had been that day, nearly four months ago, when things had been at their worst. I stabbed at a bit of fat with my fork.

\quotation{When were you going out?} My dad asked. I glanced at the digital clock on the stove and noted the time.

I was glad for the excuse to escape, \quotation{Now? Is that okay? I won't be long.}

\quotation{Meeting your friends?} he asked.

\quotation{Just going to meet Lisa for coffee and conversation, away from the rest of the group,} I told him as I stood up and moved my plate to the sink. The lie was heavier on my conscience after the open disclosure I'd just had with him.

\quotation{Here, wait,} he said. He stood up and fished in his pocket for his wallet. He handed me a ten, \quotation{For the coffee. Sorry I don't have more. Have fun?}

I hugged him, feeling painfully guilty, then headed to the back door to pull my shoes on. I was just opening the door when I barely heard him say, \quotation{Thank you.}

\quotation{Love you, Dad.}

\quotation{I love you too. Be safe.}

I shut the door, grabbed the gym bag I'd stashed under the back steps and headed around the house at a light jog. I held the gym bag low so my dad wouldn't see me carrying it.

I took the same general route I took on my morning runs, heading east, towards the Bay. This time, though, instead of turning up towards the Boardwalk, I headed south.

Back in its heyday, every inch of the city had been a bustling metropolis. Ships were coming and going at all hours, trains were coming through to deliver goods to be shipped overseas and the city teemed with people. The northern end of the bay -- especially the area close to the water -- was all about the industry. Ships, warehouses, factories, railroad and the homes for everyone who worked those jobs. You also had the ferry running across the bay itself.

The ferry was my dad's pet project. Apparently, it had been one of the first things to go when the import/export dried up. With the ferry gone, the Docks had sort of been cut off from the rest of the city, unless you were willing to drive for an extra half hour to an hour. My dad held the opinion that the lack of that transportation to the rest of the city was why the Docks had become what they were today. He believed that if the ferry were to start running again, jobs would be created, the people in the low income neighborhoods would have more access to the rest of the city, and the low-class, high-class, no-middle-class dynamic of Brockton Bay would smooth out.

So when I'd been trying to think of a place that was fairly private but easy to find, I thought of the ferry. I could probably thank my dad for the idea.

I approached the station and found a disused restroom to change into my costume.

The building and the ferry itself were well kept, at least on the outside, which was one of the reasons my dad felt it would take so little effort to get things going again. Still, that wasn't the city's issue. They didn't want to provide the addicts and the gangbangers easy access to to the rest of the city, all the while paying to provide the service, for mere hopes of {\em maybe} getting improvements for the future. So the city kept the station and the ferry looking pretty for any tourists that wandered far enough south from the Boardwalk and maintained eternal \quote{temporarily out of service} and \quote{coming soon} signs up around the building and in the brochures. Aside from the regular replacements to keep them looking new, the signs hadn't been taken down in nearly a decade.

I ignored the doors to the station's interior, and instead headed up the stairs to the outdoor patio that overlooked the bay. There were some large panes of glass to break the wind, and stone tables and benches for those wanting to sit to eat. It was one of the best vantage points for seeing the PHQ in all its splendor. The headquarters was a series of arches and spires mounted on a retrofitted oil rig. Even the platform it was built on was beautiful, though, with hard edges and sweeping lines. The entire thing was lit up by tinted spotlights and set against a faint corona of shifting colors, like the aurora borealis trapped in the shape of a soap bubble. A forcefield, forever on, shielding the people who watched over Brockton Bay.

\quotation{Wasn't sure if you would show up,} a male voice broke the silence.

I turned to face Armsmaster, \quotation{I'm sorry. I had to hang up on your receptionist. Real life called.}

He looked somehow different than the first time I'd met him. His lips were set in a hard line, his feet set further apart. His arms were folded across his chest with his Halberd in one hand, the pole resting against his shoulder. It conveyed such a different attitude that I momentarily wondered if he was the same person under the suit.

\quotation{I need to call in a favor.}



