\define[1]\paragraphnumber{¶#1.}

\showgrid
%\tracingmacros1
%\tracingoutput1
%\tracingpages}
%\tracingparagraphs}
%\tracingrestores}
%\tracinggroups}
%\tracingifs}
%\tracingnesting}
%\tracingassigns}
%\tracingmath}
%changing \tracingmath=0}
%into \tracingmath=1}
%\tracingalignments}
%changing \tracingalignments=0}
%into \tracingalignments=2}
%\showboxbreadth}
%reassigning \showboxbreadth=2147483647}
%\showboxdepth}
%reassigning \showboxdepth=2147483647}
%\newlinepar}

\paragraphnumber{72} Modern society is in certain respects extremely permissive. In matters that are irrelevant to the functioning of the system we can generally do what we please. We can believe in any religion we like (as long as it does not encourage behavior that is dangerous to the system). We can go to bed with anyone we like (as long as we practice “safe sex”). We can do anything we like as long as it is {\em unimportant}. But in all {\em important} matters
\smash{\vrule height 0pt width 10pt depth 15pt}
\smash{\vrule height 0pt depth 10pt}
\smash{\vrule height 0pt depth 10pt} the system tends increasingly to regulate our behavior.

\paragraphnumber{73} Behavior is regulated not only through explicit rules and not only by the government. Control is often exercised through indirect coercion or through psychological pressure or manipulation, and by organizations other than the government, or by the system as a whole. Most large organizations use some form of propaganda [14] to manipulate public attitudes or behavior. Propaganda is not limited to “commercials” and advertisements, and sometimes it is not even consciously intended as propaganda by the people who make it. For instance, the content of entertainment programming is a powerful form of propaganda. An example of indirect coercion: There is no law that says we have to go to work every day and follow our employer’s orders. Legally there is nothing to prevent us from going to live in the wild like primitive people or from going into business for ourselves. But in practice there is very little wild country left, and there is room in the economy for only a limited number of small business owners. Hence most of us can survive only as someone else’s employee.
