\environment assignments-environment

\starttext

\nameanddate{Megan Andrews}{January}{29}{2022}

%\chapter[ownnumber=3,title={Anatomy I: Basics}]
\vskip1.3em
\startalignment[middle]
\dontleavehmode
\raise0.35em\hbox{\rotate[rotation=180]{\scale[xscale=1260]{\symbol[zapfswash]}}}%
\hskip0.6em%
Topic 3%
\hskip0.6em
\raise0.35em\hbox{\rotate[rotation=180]{\scale[xscale=1260]{\mirror{\symbol[zapfswash]}}}}
\vskip-\lineheight
\vskip1ex
Anatomy I: Basics
\vskip0.5ex
\dontleavehmode
\scale[xscale=1885]{\mirror{\symbol[zapfswash]}}
\kern2.5pt
\scale[xscale=1885]{\symbol[zapfswash]}
\stopalignment
\vskip2.1ex

\question[ownnumber={3.1}, title={Briefly describe each of these tissues and its function: Muscle, Ligament, Tendon, Cartilage.}] 

{\sc Muscle}
\startnarrower[left]
\noindent A muscle is tissue consisting of bundles of parallel fibers wrapped in fascia.

%\start
%\setupalign[nothyphenated]
%\startcolumns[n=2,separator=rule,balance=no]
%\startalignment[middle]
%Properties
%\stopalignment
%\startalignment[right]
%\startitemize[symbol=mybullet1]
%\item Excitable – Reacts to stimuli via innervation
%\item Contractible – May contract upon stimulation in 3 different ways
%    \startitemize[symbol=mybullet2]
%    \item Concentrically – Muscle shortens
%    \item Eccentrically – Muscle lengthens
%    \item Isometrically  – Muscle neither shortens or lengthens during contraction
%    \stopitemize
%\item Extensible - May be stretched
%\item Elastic - Able to recoil
%\stopitemize
%\stopalignment
%\column
%\startalignment[middle]
%Roles
%\stopalignment
%\startitemize[symbol=mybullet1]
%\item Agonist – Prime mover muscle
%\item Synergist – Supporting muscle to help the agonist in movement
%\item Antagonist – Opposes prime mover to slow and monitor movement
%\item Protects joint
%\item Stabilizer – A muscle that pairs with an opposing muscle to keep a bone in place
%\stopitemize
%\stopcolumns
%\stop

\blank[halfline]
\noindent Properties:
\startalignment[right]
\startitemize[symbol=mybullet1]
\item Excitable – Reacts to stimuli via innervation
\item Contractible – May contract upon stimulation in 3 different ways
    \startitemize[symbol=mybullet2]
    \item Concentrically - Muscle shortens
    \item Eccentrically - Muscle lengthens
    \item Isometrically  - Muscle neither shortens or lengthens during contraction
    \stopitemize
\item Extensible - May be stretched
\item Elastic - Able to recoil
\stopitemize
\stopalignment

\blank[halfline]
\noindent Roles:
\startitemize[symbol=mybullet1]
\item Agonist - Prime mover muscle
\item Synergist - Supporting muscle to help the agonist in movement
\item Antagonist - Opposes prime mover to slow and monitor movement
\item Protects joint
\item Stabilizer - A muscle that pairs with an opposing muscle to keep a bone in place
\stopitemize
\stopnarrower
\blank[halfline]

\noindent{\sc Ligament}
\startnarrower[left]
\noindent A ligament is a dense piece of connective tissue that connects one bone to another bone.
\blank[halfline]
\startitemize[mybullet1]
\item Protects joints
\item Avascular - no blood supply
\item Barely elastic \symbol[zapfarrow]\ difficult to heal
\stopitemize
\stopnarrower
\blank[halfline]
\page[yes]

\noindent{\sc Tendon}
\startnarrower[left]
\noindent A tendon is a tough, fibrous continuation of muscle fascia that attaches muscle to bone.
\blank[halfline]
\startitemize[mybullet1]
\item Stabilizes joint
\item Transmits and supports muscle force
\item Woven into the periosteum/outer shell of the bone
\stopitemize
\stopnarrower
\blank[halfline]

\noindent{\sc Cartilage}
\startnarrower[left]
\noindent Cartilage is supplementary tissue to muscle providing strength, rigidity, and a little elasticity.
\blank[halfline]
\startitemize[mybullet1]
\item Fibrocartilage
    \startitemize[symbol=mybullet2]
    \item Cushioning, pillowy 
    \item Shock-absorbing
    \item Avascular - no blood supply
    \stopitemize
\item Hyaline Cartilage
    \startitemize[symbol=mybullet2]
    \item Smooth, glassy
    \item Reinforcing coat where bones rub against one another
    \item Avascular - no blood supply
    \stopitemize
\stopitemize
\stopnarrower
\vskip0.5\lineheight

\question[ownnumber={3.2},title={The elbow is a hinge joint and the hip is a ball-and-socket joint.
What are the movements of a hinge joint? Name the plane.}]

The movements of a hinge joint are flexion and extension, and the plane is the Sagittal Plane.

\question[ownnumber={3.3},title={What are the movements of a ball-and-socket joint? In how many planes does a ball-and-socket joint move and what are the actions in each of the planes?}]

The movements of a ball-and-socket joint and their corresponding planes are:
\startitemize[n]
\item Rotation (internal and external) in the Transverse Plane
\item Flexion & Extension in the Sagittal Plane
\item Adduction & Abduction in Coronal Plane
\stopitemize
\startitemize[symbol=bulletstar]
\item Circumduction involves all movements and planes in combination, except rotation in the Transverse plane
\stopitemize

\question[ownnumber={3.4},title={What is the job of an agonist? What is the job of a synergist?}]

An agonist is the prime mover muscle that does the bulk of the work in a movement. The synergist is a supporting muscle that assists the agonist in doing the movement or compensates if the agonist is injured/dysfunctional.

\stoptext

