\environment assignments-environment

\starttext

\nameanddate{Megan Andrews}{January}{29}{2022}

\chapter[ownnumber=3,title={Anatomy I: Basics}]

\question{Briefly describe each of these tissues and its function: Muscle, Ligament, Tendon and Cartilage} 

{\sc A muscle} is tissue consisting of bundles of parallel fibers wrapped in fascia.
\blank[line]

\definesymbol[dotmath][\dotmath]

\startcolumns[n=2,separator=rule,balance=no]
\startalignment[middle]
Properties
\stopalignment
\startitemize[symbol=mybullet1]
\item Excitable – Reacts to stimuli via innervation
\item Contractible – May contract upon stimulation in 3 different ways
    \startitemize[symbol=mybullet2]
    \item Concentrically – Muscle shortens
    \item Eccentrically – Muscle lengthens
    \item Isometrically  – Muscle neither shortens or lengthens during contraction
    \stopitemize
\item Extensible - May be stretched
\item Elastic - Able to recoil
\stopitemize
\column
\startalignment[middle]
Roles
\stopalignment
\startitemize[symbol=mybullet1]
\item Agonist – Prime mover muscle
\item Synergist – Supporting muscle to help the agonist in movement
\item Antagonist – Opposes prime mover to slow and monitor movement
\item Protects joint
\item Stabilizer – A muscle that pairs with an opposing muscle to keep a bone in place
\stopitemize
\stopcolumns






%LIGAMENT: 
A {\sc ligament} is a dense piece of connective tissue that connects one bone to another bone.
Protects joints
Avascular - no blood supply
Barely elastic → difficult to heal

%TENDON: 
A {\sc tendon} is a tough, fibrous continuation of muscle fascia that attaches muscle to bone.
Stabilizes joint
Transmits and supports muscle force
Woven into the periosteum/outer shell of the bone


CARTILAGE: Cartilage is supplementary tissue to muscle providing strength, rigidity, and a little elasticity.

Fibrocartilage
Hyaline Cartilage
Cushioning, pillowy 
Shock-absorbing
Avascular - no blood supply
Smooth, glassy
Reinforcing coat where bones rub against one another
Avascular - no blood supply



\question{The elbow is a hinge joint and the hip is a ball-and-socket joint.
What are the movements of a hinge joint? Name the plane.}

The movements of a hinge joint are flexion and extension, and the plane is the Sagittal Plane.

\question{What are the movements of a ball-and-socket joint? In how many planes does a ball-and-socket joint move and what are the actions in each of the planes?}

The movements of a ball-and-socket joint and their corresponding planes are:
1. Rotation (internal and external) in the Transverse Plane
2. Flexion & Extension in the Sagittal Plane
3. Adduction & Abduction in Coronal Plane
* Circumduction involves all movements and planes in combination

\question{What is the job of an agonist? What is the job of a synergist?}

An agonist is the prime mover muscle that does the bulk of the work in a movement. The synergist is a supporting muscle that assists the agonist in doing the movement or compensates if the agonist is injured/dysfunctional.

\stoptext

